### 2020-03-10 draft_analysis.Rmd
first commit

---
title: "COVID19 - Forecast analysis"
author: "PG"
date: "3/11/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
###import italian dataset updated 10 March 2020
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
dat_csv$t<-1:dim(dat_csv)[1]
```

## The COVID dataset

The present analysis used the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19.
We used a SEIR model to estimate the prediction of this epidemic in Italy.
We estimated the R0 parameter by means of a simple linear regression ad in https://kingaa.github.io/clim-dis/parest/parest.html

fit1 <- lm(log(totale_casi)~t,data=dat_csv)

```{r cars}
show(dat_csv)
fit1 <- lm(log(totale_casi)~t,data=dat_csv)
summary(fit1)
```

Estimates

You can also embed plots, for example:

```{r model plot, echo=TRUE}
plot(dat_csv$t,log(dat_csv$totale_casi),ylab="log cases",xlab="time")
abline(coef(summary(fit1))[,1])
```
The slope coefficient estimated in the linear regression model can be used to estimate R0, the number of 
```{r r0 plot, echo=TRUE}
slope <-coef(summary(fit1))[2,1]; slope
slope.se <- coef(summary(fit1))[2,2]; slope.se
### R0 estimates and 95%IC 
R_0=slope*14+1;R_0
(slope+c(-1,1)*1.96*slope.se)*14+1
```
We want to make a short term forecast (14 days) with 3 scenario.
We fix a series of initial parameters.
```{r initation plot, echo=TRUE}
# initial number of infectus
I0<-max(dat_csv$totale_casi); I0
# initial number of recovered
R0<-max(dat_csv$dimessi_guariti); R0
#beta 
beta0<-R_0/(14)
# italian poulation
N=60480000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics)
```{r first scenario plot, echo=TRUE}
library(EpiDynamics)
parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
# average number of single connections of an infected person
# less contacts, less probability of new infections
# we keep constant the other parameters
f1<-10
initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
seir1 <- SEIR(pars = parameters, init = initials, time = 0:14)

f2<-5
initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
seir2 <- SEIR(pars = parameters, init = initials, time = 0:14)

f3<-3
initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
seir3 <- SEIR(pars = parameters, init = initials, time = 0:14)

plot(c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
lines(c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
lines(c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```

### 2020-03-11 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

---
title: "COVID19 - Forecast analysis"
author: "PG"
date: "3/11/2020"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
rm(list=ls())
###import italian dataset updated 10 March 2020
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",header=T)
dat_csv<-dat_csv[dat_csv$codice_regione==5,]
dat_csv$t<-1:dim(dat_csv)[1]
```

fit1 <- lm(log(totale_casi)~t,data=dat_csv)

```{r cars}
show(dat_csv)
fit1 <- lm(log(totale_casi)~t,data=dat_csv)
summary(fit1)
```

Estimates

```{r model plot, echo=TRUE}
plot(dat_csv$t,log(dat_csv$totale_casi),ylab="log cases",xlab="time")
abline(coef(summary(fit1))[,1])
```
The slope coefficient estimated in the linear regression model can be used to estimate R0, the number of 
```{r r0 plot, echo=TRUE}
slope <-coef(summary(fit1))[2,1]; slope
slope.se <- coef(summary(fit1))[2,2]; slope.se
### R0 estimates and 95%IC 
### I have used 14 for infection time, but it has a bimodal distribution (tested vs non tested)
R_0=slope*14+1;R_0
(slope+c(-1,1)*1.96*slope.se)*14+1
```
We want to make a short term forecast (14 days) with 3 scenario.
We fix a series of initial parameters.
```{r initation plot, echo=TRUE}
# initial number of infectus
I0<-max(dat_csv$totale_casi); I0
# initial number of recovered
R0<-max(dat_csv$dimessi_guariti); R0
#beta 
beta0<-R_0/(14)
# Veneto poulation
N=4800000
# duration of COVID19 
duration<-14
#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
sigma0<-0.05
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```
We use the library(EpiDynamics)
```{r first scenario plot, echo=TRUE}
library(EpiDynamics)
parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
# average number of single connections of an infected person
# less contacts, less probability of new infections
# we keep constant the other parameters
f1<-10
initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
seir1 <- SEIR(pars = parameters, init = initials, time = 0:14)

f2<-5
initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
seir2 <- SEIR(pars = parameters, init = initials, time = 0:14)

f3<-3
initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
seir3 <- SEIR(pars = parameters, init = initials, time = 0:14)

plot(c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
lines(c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
lines(c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
```

### 2020-03-11 Previsioni_COVID_2019.R

@@ -1,10 +1,10 @@
 rm(list=ls())
-###italia
+###import italian dataset updated 10 March 2020
 dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
 dat_csv$t<-1:dim(dat_csv)[1]
-##### SEIR model semplice
+##### SEIR model
 
-v_c<-NULL
+#stimo r0
 #vedi https://kingaa.github.io/clim-dis/parest/parest.html
 
 fit1 <- lm(log(totale_casi)~t,data=dat_csv)
@@ -33,13 +33,22 @@ duratac<-14
 mu0<-1/(82*365.25) # 1/durata della vita
 parameters <- c(mu = mu0, beta = 1/14, sigma = 0.05, gamma = 1/duratac)
 # numero medio di persone esposte per ogni infetto
-f1<-
-  initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
-#Solve and plot.
-seir <- SEIR(pars = parameters, init = initials, time = 0:60) 
-c(dat_csv$totale_casi,seir$results$I[-1]*60480000)
+f1<-10
+initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
+seir1 <- SEIR(pars = parameters, init = initials, time = 0:14)
 
-plot(c(dat_csv$totale_casi,seir$results$I[-1]*60480000))
+f2<-5
+initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
+seir2 <- SEIR(pars = parameters, init = initials, time = 0:14)
 
-PlotMods(seir)
+f3<-3
+initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
+seir3 <- SEIR(pars = parameters, init = initials, time = 0:14)
+
+
+
+plot(c(dat_csv$totale_casi,seir1$results$I[-1]*60480000),type="l",ylab="Number of infectus",xlab="time")
+lines(c(dat_csv$totale_casi,seir2$results$I[-1]*60480000),col=2)
+lines(c(dat_csv$totale_casi,seir3$results$I[-1]*60480000),col=3)
+legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
 

### 2020-03-11 draft_analysis_Italy.Rmd
Add home page, navigation and theme

--- 2020-03-10 draft_analysis.Rmd
@@
 date: "3/11/2020"
-output: html_document
+output: github_document
 ---
@@
 
+```{r  plot data, echo=TRUE}
+plot(dat_csv$data,dat_csv$totale_casi,ylab="Total Covid cases",xlab="Date")
+mean(diff(log(dat_csv$totale_casi)))
+```
+The grow is exponential (rate of increase of about 25\%). 
+We estimate the R0 parameter by means of a linear model. R0 indicates how contagious an infectious disease is. It's also referred to as the reproduction number.
+
+- <img src="https://latex.codecogs.com/gif.latex?log(Y_t) = a + b*t + e_t" />
+
+where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.
+
+The slope b indicate the rate of exponetial increase, used for the R0 calculation.   
+R code:  
 fit1 <- lm(log(totale_casi)~t,data=dat_csv)
 
-```{r cars}
-show(dat_csv)
+```{r data}
+head(dat_csv)
 fit1 <- lm(log(totale_casi)~t,data=dat_csv)
@@
 ```
-
-Estimates
-
-You can also embed plots, for example:
+The prediction is  good ($R^2$ near 1).
+However the tendency is for a reduction of the slope.
 
@@
 ```
-The slope coefficient estimated in the linear regression model can be used to estimate R0, the number of 
+The slope coefficient estimated in the linear regression model can be used to estimate R0.
+
+R0=1+b*incubation period.
+
+The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
+Please se https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
+In the calculation we considered an incubation period of 14 days for two reasons: 1) the majority are asymptomatic, contagiousness is greater than 5, maybe 14.
+A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to begin.
 ```{r r0 plot, echo=TRUE}
@@
 slope.se <- coef(summary(fit1))[2,2]; slope.se
-### R0 estimates and 95%IC 
+### R0 estimates and 95% IC 
 R_0=slope*14+1;R_0
@@
 ```
-We want to make a short term forecast (14 days) with 3 scenario.
-We fix a series of initial parameters.
+We want to make a short term forecast (14 days) with 3 scenario:  
+-Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
+-Scenario 2: 5 exposed people for each COVID-19 case and beta reduced of 50% (-50% both exposed people and -50% COVID19 contagious power)
+-Scenario 3: 3 exposed people for each COVID-19 case and beta reduced of 50%  (-70% both exposed people and -50% COVID19 contagious power)
+
+We fix a series of initial parameters:
+-I0: initial number of COVID-19 cases  
+- R0: initial number of recovered  
+- beta: the quantity connected to R0  
+- N: Italian population   
+- duration: infection duration of COVID-19   
+- sigma0: the coronavirus transmission rate (half of flu epidemic)  
+- mu0: the overall mortality rate
+
 ```{r initation plot, echo=TRUE}
@@
 library(EpiDynamics)
-parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
+
 # average number of single connections of an infected person
@@
 # we keep constant the other parameters
+forecast<-14
+parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
 f1<-10
 initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
-seir1 <- SEIR(pars = parameters, init = initials, time = 0:14)
-
+seir1 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
+parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
 f2<-5
 initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
-seir2 <- SEIR(pars = parameters, init = initials, time = 0:14)
-
+seir2 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
+parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
 f3<-3
 initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
-seir3 <- SEIR(pars = parameters, init = initials, time = 0:14)
+seir3 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
 
-plot(c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
-lines(c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
-lines(c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
+date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
+plot(date,c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
+lines(date,c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
+lines(date,c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
+lines(date[1:dim(dat_csv)[1]],dat_csv$totale_casi,lwd=2)
 legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)

### 2020-03-11 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

@@
 ```{r cars}
-show(dat_csv)
+head(dat_csv)
 fit1 <- lm(log(totale_casi)~t,data=dat_csv)
@@
 
-f3<-3
+f3<-2
 initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
@@
 lines(c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
-legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
+legend("topleft",c("1° scenario 10exp for case","2° scenario 5exp fo case","3° scenario 2exp for case"),lty=1,col=1:3)
 ```

### 2020-03-12 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 ```{r  plot data, echo=TRUE}
-plot(dat_csv$data,dat_csv$totale_casi,ylab="Total Covid cases",xlab="Date")
-mean(diff(log(dat_csv$totale_casi)))
+plot(dat_csv$data,dat_csv$totale_attualmente_positivi,ylab="Total Covid cases",xlab="Date")
+mean(diff(log(dat_csv$totale_attualmente_positivi)))
 ```
@@
 R code:  
-fit1 <- lm(log(totale_casi)~t,data=dat_csv)
+fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
 
@@
 head(dat_csv)
-fit1 <- lm(log(totale_casi)~t,data=dat_csv)
+fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
 summary(fit1)
@@
 ```{r model plot, echo=TRUE}
-plot(dat_csv$t,log(dat_csv$totale_casi),ylab="log cases",xlab="time")
+plot(dat_csv$t,log(dat_csv$totale_attualmente_positivi),ylab="log cases",xlab="time")
 abline(coef(summary(fit1))[,1])
@@
 -Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
--Scenario 2: 5 exposed people for each COVID-19 case and beta reduced of 50% (-50% both exposed people and -50% COVID19 contagious power)
--Scenario 3: 3 exposed people for each COVID-19 case and beta reduced of 50%  (-70% both exposed people and -50% COVID19 contagious power)
+-Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% COVID19 contagious power)
+-Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)
 
@@
 # initial number of infectus
-I0<-max(dat_csv$totale_casi); I0
+I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I0
 # initial number of recovered
-R0<-max(dat_csv$dimessi_guariti); R0
+R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
 #beta 
@@
 We use the library(EpiDynamics)
-```{r first scenario plot, echo=TRUE}
+```{r  scenario plot, echo=TRUE}
 library(EpiDynamics)
@@
 parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
-f2<-5
+f2<-10
 initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
@@
 parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
-f3<-3
+f3<-5
 initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
@@
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-plot(date,c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
-lines(date,c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
-lines(date,c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
-lines(date[1:dim(dat_csv)[1]],dat_csv$totale_casi,lwd=2)
+plot(date,c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N),type="l",ylab="Cases",xlab="time",main="Infected")
+lines(date,c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N),col=2)
+lines(date,c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N),col=3)
+lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
 legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
 ```
+The 3 scenarios show how measures of restriction can help to reduce the number of infected.  
+At the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$results$I*N)[length(seir2$results$I)]`).
+In the next plot the cumulative number of infected.
+```{r  cumulative plot, echo=TRUE}
+
+plot(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N)),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
+lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N)),col=2)
+lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N)),col=3)
+lines(date[1:dim(dat_csv)[1]],cumsum(dat_csv$totale_attualmente_positivi),lwd=2)
+legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
+```

### 2020-03-12 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

--- 2020-03-11 draft_analysis_Veneto.Rmd
@@
 dat_csv$t<-1:dim(dat_csv)[1]
+plot(dat_csv$data,dat_csv$totale_attualmente_positivi,ylab="Infected",xlab="Date")
 ```
 
-fit1 <- lm(log(totale_casi)~t,data=dat_csv)
+fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
 
 ```{r cars}
-show(dat_csv)
-fit1 <- lm(log(totale_casi)~t,data=dat_csv)
+head(dat_csv)
+fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
 summary(fit1)
@@
 ```
-We want to make a short term forecast (14 days) with 3 scenario.
-We fix a series of initial parameters.
+We want to make a short term forecast (14 days) with 3 scenario:  
+-Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
+-Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% COVID19 contagious power)
+-Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)
+
+We fix a series of initial parameters:
+-I0: initial number of COVID-19 cases  
+- R0: initial number of recovered  
+- beta: the quantity connected to R0  
+- N: Italian population   
+- duration: infection duration of COVID-19   
+- sigma0: the coronavirus transmission rate (half of flu epidemic)  
+- mu0: the overall mortality rate  
 ```{r initation plot, echo=TRUE}
 # initial number of infectus
-I0<-max(dat_csv$totale_casi); I0
+I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I0
 # initial number of recovered
-R0<-max(dat_csv$dimessi_guariti); R0
+R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
 #beta 
 beta0<-R_0/(14)
-# Veneto poulation
-N=4800000
+# Veneto population
+N=4905000
 # duration of COVID19 
@@
 library(EpiDynamics)
+forecast<-14
 parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
-# average number of single connections of an infected person
-# less contacts, less probability of new infections
-# we keep constant the other parameters
 f1<-10
 initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
-seir1 <- SEIR(pars = parameters, init = initials, time = 0:14)
+seir1 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
+parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
+f2<-10
+initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
+seir2 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
+parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
+f3<-5
+initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
+seir3 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
 
-f2<-5
-initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
-seir2 <- SEIR(pars = parameters, init = initials, time = 0:14)
-
-f3<-3
-initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
-seir3 <- SEIR(pars = parameters, init = initials, time = 0:14)
-
-plot(c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
-lines(c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
-lines(c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
+date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
+plot(date,c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N),type="l",ylab="Cases",xlab="time",main="Infected")
+lines(date,c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N),col=2)
+lines(date,c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N),col=3)
+lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
 legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
 ```
+The 3 scenarios show how measures of restriction can help to reduce the number of infected.  
+At the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$results$I*N)[length(seir2$results$I)]`).
+In the next plot the cumulative number of infected.
+```{r  cumulative plot, echo=TRUE}
+
+plot(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N)),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
+lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N)),col=2)
+lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N)),col=3)
+lines(date[1:dim(dat_csv)[1]],cumsum(dat_csv$totale_attualmente_positivi),lwd=2)
+legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
+```

### 2020-03-12 Previsioni_COVID_2019.R

@@ -1,28 +1,29 @@
 rm(list=ls())
-###import italian dataset updated 10 March 2020
+#base R code for
+###import italian dataset updated 11 March 2020
 dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
 dat_csv$t<-1:dim(dat_csv)[1]
 ##### SEIR model
 
-#stimo r0
+#estimate r0
 #vedi https://kingaa.github.io/clim-dis/parest/parest.html
 
-fit1 <- lm(log(totale_casi)~t,data=dat_csv)
+fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
 
 # stimo 
 
-plot(dat_csv$t,log(dat_csv$totale_casi))
+plot(dat_csv$t,log(dat_csv$totale_attualmente_positivi))
 abline(coef(summary(fit1))[,1])
 summary(fit1)
 # R0
 slope <-coef(summary(fit1))[2,1]; slope
 slope.se <- coef(summary(fit1))[2,2]; slope.se
-###R0 stima + o meno co durata contagio 14 giorni
+###R0 incubation time of 14 days
 R_0=slope*14+1;R_0
 (slope+c(-1,1)*1.96*slope.se)*14+1
 
-I0<-max(dat_csv$totale_casi)
-R0<-max(dat_csv$dimessi_guariti)
+I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]
+R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]] 
 N<-60480000
 library(EpiDynamics)
 #beta
@@ -31,24 +32,30 @@ beta0<-R_0/(14)
 duratac<-14
 #ipotiziamo un sigma (coronavirus transmission rate) del 5%  (la metà dell'influenza)
 mu0<-1/(82*365.25) # 1/durata della vita
-parameters <- c(mu = mu0, beta = 1/14, sigma = 0.05, gamma = 1/duratac)
+parameters <- c(mu = mu0, beta = beta0, sigma = 0.05, gamma = 1/duratac)
 # numero medio di persone esposte per ogni infetto
 f1<-10
 initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
 seir1 <- SEIR(pars = parameters, init = initials, time = 0:14)
-
[…]

### 2020-03-12 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
-Please se https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
-In the calculation we considered an incubation period of 14 days for two reasons: 1) the majority are asymptomatic, contagiousness is greater than 5, maybe 14.
-A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to begin.
+Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
+In the calculation we considered an incubation period of 14 days for two reasons: 1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14.
+A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider.
 ```{r r0 plot, echo=TRUE}

### 2020-03-12 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
-dat_csv$t<-1:dim(dat_csv)[1]
+days<-dim(dat_csv)[1]
+dat_csv$t<-1:days
+
 ```
@@
 
+#estimate r0
+#see https://kingaa.github.io/clim-dis/parest/parest.html
+#calculate r0 based in the last XX observation 
+```{r R0 trend}
+beta_vec<-NULL
+sd_vec<-NULL
+for (i in 1:(days-6)){
+fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
+beta_vec<-c(beta_vec,coef(fit)[2])
+sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
+}
+
+label <-days:7
+mean  <- (beta_vec*14+1)
+lower <- ((beta_vec-1.96*sd_vec)*14+1)
+upper <- ((beta_vec+1.96*sd_vec)*14+1)
+
+df <- data.frame(label, mean, lower, upper)
+df$label <- factor(df$label, levels=rev(df$label))
+
+library(ggplot2)
+fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
+  geom_pointrange() + 
+  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
+  coord_flip() +  # flip coordinates (puts labels on y axis)
+  xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
+  theme_bw()  # use a white background
+print(fp)
+```
+
 - <img src="https://latex.codecogs.com/gif.latex?log(Y_t) = a + b*t + e_t" />
@@
 where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.  
+We calculate the R0 values based on a different number of days before the last. 
+The R0 shows a decreasing trend, but it is stable in the last 12 days.
 
@@
 -Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
--Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% COVID19 contagious power)
+-Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% exposed people)
 -Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)
@@
 seir1 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
-parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
-f2<-10
+parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
+f2<-5
 initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
@@
 
-plot(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N)),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
-lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N)),col=2)
-lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N)),col=3)
-lines(date[1:dim(dat_csv)[1]],cumsum(dat_csv$totale_attualmente_positivi),lwd=2)
+plot(date,c(dat_csv$totale_casi,(seir1$results$I[-1]+seir1$results$R[-1])*N),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
+lines(date,c(dat_csv$totale_casi,(seir2$results$I[-1]+seir2$results$R[-1])*N),col=2)
+lines(date,c(dat_csv$totale_casi,(seir3$results$I[-1]+seir3$results$R[-1])*N),col=3)
+lines(date[1:dim(dat_csv)[1]],(dat_csv$totale_casi),lwd=2)
 legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
+PlotMods(seir2)
 ```

### 2020-03-12 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

--- 2020-03-11 draft_analysis_Veneto.Rmd
@@
 dat_csv$t<-1:dim(dat_csv)[1]
+plot(dat_csv$data,dat_csv$totale_attualmente_positivi,ylab="Infected",xlab="Date")
 ```
 
-fit1 <- lm(log(totale_casi)~t,data=dat_csv)
+fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
+The actual situation is 
+```{r plots}
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+df <- dat_csv %>%
+  select(data, ricoverati_con_sintomi, terapia_intensiva, 
+         totale_ospedalizzati, isolamento_domiciliare, 
+         totale_attualmente_positivi, nuovi_attualmente_positivi, 
+         dimessi_guariti, deceduti, totale_casi) %>%
+  gather(key = "variable", value = "value", -data)
+head(df, 3)
+
+ggplot(df, aes(x = as.Date(data), y = value)) + 
+  geom_line(aes(color = variable), size = 1) 
+  ```
 
 ```{r cars}
-show(dat_csv)
-fit1 <- lm(log(totale_casi)~t,data=dat_csv)
+head(dat_csv)
+fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
 summary(fit1)
@@
 ```
-We want to make a short term forecast (14 days) with 3 scenario.
-We fix a series of initial parameters.
+We want to make a short term forecast (14 days) with 3 scenario:  
+-Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)  
+-Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% exposed people)  
+-Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)
+
+We fix a series of initial parameters:
+-I0: initial number of COVID-19 cases  
+- R0: initial number of recovered  
+- beta: the quantity connected to R0  
+- N: Italian population   
+- duration: infection duration of COVID-19   
+- sigma0: the coronavirus transmission rate (half of flu epidemic)  
+- mu0: the overall mortality rate  
 ```{r initation plot, echo=TRUE}
 # initial number of infectus
-I0<-max(dat_csv$totale_casi); I0
+I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I0
 # initial number of recovered
-R0<-max(dat_csv$dimessi_guariti); R0
+R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
 #beta 
 beta0<-R_0/(14)
-# Veneto poulation
-N=4800000
+# Veneto population
+N=4905000
 # duration of COVID19 
@@
 library(EpiDynamics)
+forecast<-14
 parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
-# average number of single connections of an infected person
-# less contacts, less probability of new infections
-# we keep constant the other parameters
 f1<-10
 initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
-seir1 <- SEIR(pars = parameters, init = initials, time = 0:14)
-
+seir1 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
+parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
 f2<-5
 initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
-seir2 <- SEIR(pars = parameters, init = initials, time = 0:14)
+seir2 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
+parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
+f3<-5
+initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
+seir3 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
 
-f3<-3
-initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
-seir3 <- SEIR(pars = parameters, init = initials, time = 0:14)
-
-plot(c(dat_csv$totale_casi,seir1$results$I[-1]*N),type="l",ylab="Number of infectus",xlab="time")
-lines(c(dat_csv$totale_casi,seir2$results$I[-1]*N),col=2)
-lines(c(dat_csv$totale_casi,seir3$results$I[-1]*N),col=3)
+date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
+plot(date,c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N),type="l",ylab="Cases",xlab="time",main="Infected")
+lines(date,c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N),col=2)
+lines(date,c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N),col=3)
+lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
 legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
 ```
+
+The 3 scenarios show how measures of restriction can help to reduce the number of infected.    
+At the end of the 2 weeks (`r date[length(date)]`) the number of infected is estimated to be  (`r (seir2$results$I*N)[length(seir2$results$I)]`).  
+In the next plot the cumulative number of infected.  
+```{r  cumulative plot, echo=TRUE}
+
+plot(date,c(dat_csv$totale_casi,(seir1$results$I[-1]+seir1$results$R[-1])*N),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
+lines(date,c(dat_csv$totale_casi,(seir2$results$I[-1]+seir2$results$R[-1])*N),col=2)
+lines(date,c(dat_csv$totale_casi,(seir3$results$I[-1]+seir3$results$R[-1])*N),col=3)
+lines(date[1:dim(dat_csv)[1]],dat_csv$totale_casi,lwd=2)
+legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
+PlotMods(seir2)
+
+```

### 2020-03-12 Previsioni_COVID_2019.R

@@ -2,11 +2,43 @@ rm(list=ls())
 #base R code for
 ###import italian dataset updated 11 March 2020
 dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
-dat_csv$t<-1:dim(dat_csv)[1]
+days<-dim(dat_csv)[1]
+dat_csv$t<-1:days
+
 ##### SEIR model
 
 #estimate r0
-#vedi https://kingaa.github.io/clim-dis/parest/parest.html
+#see https://kingaa.github.io/clim-dis/parest/parest.html
+#calculate r0 based in the last XX observation 
+beta_vec<-NULL
+sd_vec<-NULL
+for (i in 1:(days-6)){
+fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
+beta_vec<-c(beta_vec,coef(fit)[2])
+sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
+}
+
+label <-days:7
+mean  <- (beta_vec*14+1)
+lower <- ((beta_vec-1.96*sd_vec)*14+1)
+upper <- ((beta_vec+1.96*sd_vec)*14+1)
+
+df <- data.frame(label, mean, lower, upper)
+
+# reverses the factor level ordering for labels after coord_flip()
+df$label <- factor(df$label, levels=rev(df$label))
+
+library(ggplot2)
+fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
+  geom_pointrange() + 
+  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
+  coord_flip() +  # flip coordinates (puts labels on y axis)
+  xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
+  theme_bw()  # use a white background
+print(fp)
+
+
+
 
 fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
 
@@ -37,12 +69,14 @@ parameters <- c(mu = mu0, beta = beta0, sigma = 0.05, gamma = 1/duratac)
 f1<-10
 initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
[…]

### 2020-03-13 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 author: "PG"
-date: "3/11/2020"
+date: "3/13/2020"
 output: github_document
@@
 The present analysis used the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19.
-We used a SEIR model to estimate the prediction of this epidemic in Italy.
-We estimated the R0 parameter by means of a simple linear regression ad in https://kingaa.github.io/clim-dis/parest/parest.html
+We used a SEIR model to predict the trend of the actual COVID19 epidemic in Italy.
+We estimated the R0 parameter by means of a linear regression model as reported in https://kingaa.github.io/clim-dis/parest/parest.html
 
@@
 ```
-The grow is exponential (rate of increase of about 25\%). 
-We estimate the R0 parameter by means of a linear model. R0 indicates how contagious an infectious disease is. It's also referred to as the reproduction number.
+The plot shows an exponential gros.
+We estimate the R0 parameter by means of a linear model.
+R0 indicates how contagious an infectious disease is.
+It's also referred to as "the reproduction number" of COVID19.
 
@@
 where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.  
-We calculate the R0 values based on a different number of days before the last. 
-The R0 shows a decreasing trend, but it is stable in the last 12 days.
-
-The slope b indicate the rate of exponetial increase, used for the R0 calculation.   
+We calculate several R0 values, each one based on a different number of days before the last day. 
+The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponetial increase.  
+However we considere the value of the last 10 days, because is more stable.
 R code:  
-fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
-
 ```{r data}
 head(dat_csv)
-fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
+fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(days-10):days,])
 summary(fit1)
 ```
-The prediction is  good ($R^2$ near 1).
-However the tendency is for a reduction of the slope.
 
-```{r model plot, echo=TRUE}
-plot(dat_csv$t,log(dat_csv$totale_attualmente_positivi),ylab="log cases",xlab="time")
-abline(coef(summary(fit1))[,1])
-```
 The slope coefficient estimated in the linear regression model can be used to estimate R0.
@@
 In the calculation we considered an incubation period of 14 days for two reasons: 1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14.
-A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider.
+A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider in this phase.
 ```{r r0 plot, echo=TRUE}
@@
 We fix a series of initial parameters:
--I0: initial number of COVID-19 cases  
+- I0: initial number of COVID-19 cases  
 - R0: initial number of recovered  
@@
 ```
-We use the library(EpiDynamics)
+We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:
+
+<img src="http://www.public.asu.edu/~hnesse/classes/seireqn.png"/>
+
 ```{r  scenario plot, echo=TRUE}
@@
 ```
-The 3 scenarios show how measures of restriction can help to reduce the number of infected.  
-At the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$results$I*N)[length(seir2$results$I)]`).
+
+The 3 scenarios show different numbers. 
+If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$results$I*N)[length(seir2$results$I)]`).
+
 In the next plot the cumulative number of infected.  
+At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r ((seir2$results$I+seir3$results$R)*N)[length(seir2$results$I)]`).
+
 ```{r  cumulative plot, echo=TRUE}
@@
 legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
-PlotMods(seir2)
 ```
+
+We make an effort to estimate the effect of restriction measures took by the Italian Government (11 March 2020). 
+After 14 days  (`r date[length(date)]`) we have the current status in the second scenario:  
+- I14=(`r (seir2$results$I*N)[length(seir2$results$I)]`);  
+- R14=(`r (seir2$results$R*N)[length(seir2$results$R)]`);  
+- S14=(`r (seir2$results$S*N)[length(seir2$results$S)]`);  
+- E14=(`r (seir2$results$E*N)[length(seir2$results$E)]`);
+
+We reduce R0 to 1.10 and the number of exposed to the number of infected.
+We estimate the trend of the number infected, recovered and the total cases for other 14 days forward.
+
+```{r  scenario plot2, echo=TRUE}
+I14=seir2$results$I[length(seir2$results$I)]
+R14=seir2$results$R[length(seir2$results$I)]
+S14=seir2$results$S[length(seir2$results$I)]
+E14=seir2$results$E[length(seir2$results$I)]
+forecast2<-14
+beta14<-1.1/14
+parameters <- c(mu = mu0, beta = beta14, sigma = sigma0, gamma = 1/duration)
+initials <- c(S = 1-I14*2-R14, E = I14, I = I14, R = R14)
+seir2_2 <- SEIR(pars = parameters, init = initials, time = 0:forecast2)
+
+date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast2+forecast-1+dim(dat_csv)[1],1)
+plot(date,c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N,seir2_2$results$I[-1]*N),type="l",ylab="Cases",xlab="time",ylim=c(0,((seir2_2$results$I+seir2_2$results$R[-1])*N)[forecast2]),col=2)
+lines(date,c(dat_csv$dimessi_guariti,seir2$results$R[-1]*N,seir2_2$results$R[-1]*N),col=3)
+lines(date,c(dat_csv$totale_casi,(seir2$results$I[-1]+seir2$results$R[-1])*N,c(seir2_2$results$I[-1]+seir2_2$results$R[-1])*N),col=1)
+
+legend("topleft",c("All cases","Infected","Recovered"),lty=1,col=1:3)
+```

### 2020-03-13 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

@@
 author: "PG"
-date: "3/11/2020"
+date: "3/13/2020"
 output: rmarkdown::github_document
@@
 rm(list=ls())
-###import italian dataset updated 10 March 2020
+###import italian dataset updated 13 March 2020
 dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",header=T)
@@
 - duration: infection duration of COVID-19   
-- sigma0: the coronavirus transmission rate (half of flu epidemic)  
+- sigma0: the coronavirus transmission rate (0.05 is the half of flu epidemic)  
 - mu0: the overall mortality rate  
@@
 # Veneto population
-N=4905000
+N=4905854
 # duration of COVID19 

### 2020-03-13 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 
-```{r  plot data, echo=TRUE}
-plot(dat_csv$data,dat_csv$totale_attualmente_positivi,ylab="Total Covid cases",xlab="Date")
-mean(diff(log(dat_csv$totale_attualmente_positivi)))
+The actual status is
+
+```{r plots}
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+df <- dat_csv %>%
+  select(data, ricoverati_con_sintomi, terapia_intensiva, 
+         totale_ospedalizzati, isolamento_domiciliare, 
+         totale_attualmente_positivi, nuovi_attualmente_positivi, 
+         dimessi_guariti, deceduti, totale_casi) %>%
+  gather(key = "variable", value = "value", -data)
+head(df, 3)
+
+ggplot(df, aes(x = as.Date(data), y = value)) + 
+  geom_line(aes(color = variable), size = 1) 
   ```
-The plot shows an exponential gros.
+
+The plot shows an exponential grow.
 We estimate the R0 parameter by means of a linear model.
 R0 indicates how contagious an infectious disease is.
-It's also referred to as "the reproduction number" of COVID19.
+It is also referred to as "the reproduction number" of COVID19.
 
@@
 However we considere the value of the last 10 days, because is more stable.
+
 R code:  
@@
 We want to make a short term forecast (14 days) with 3 scenario:
+
 -Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
+
 -Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% exposed people)
+
 -Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)

### 2020-03-14 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 ---
-title: "COVID19 - Forecast analysis"
+title: "COVID19 - Forecast and predictions using a time dependent SEIR model" 
 author: "PG"
@@
 sd_vec<-NULL
-for (i in 1:(days-6)){
+for (i in 1:(days-4)){
 fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
@@
 
-label <-days:7
-mean  <- (beta_vec*14+1)
-lower <- ((beta_vec-1.96*sd_vec)*14+1)
-upper <- ((beta_vec+1.96*sd_vec)*14+1)
+label <-days:5
+mean  <- (beta_vec*14)
+lower <- ((beta_vec-1.96*sd_vec)*14)
+upper <- ((beta_vec+1.96*sd_vec)*14)
 
@@
 print(fp)
+
 ```
 
-- <img src="https://latex.codecogs.com/gif.latex?log(Y_t) = a + b*t + e_t" />
+Y_t= a + beta * t +e_t
 
 where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.  
+The slope coefficient estimated in the linear regression model can be used to estimate R0.
+
+R0=slope*incubation period.
+
+The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
+Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/, but the incubation period is used for epidemic diseases that causes the immediate isolation of the infected subject.
+
+In the calculation we considered an "incubation period" of 14 days for two reasons: 1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14.
+A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider in this phase.
+
 We calculate several R0 values, each one based on a different number of days before the last day. 
 The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponetial increase.  
-However we considere the value of the last 10 days, because is more stable.
+However,R0 is going to decrease in the next days. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
+We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing is a value greater than 0. 
+```{r R0 forecast}
+library(splines)
+time<-1:(days-4)
 
-R code:  
-```{r data}
-head(dat_csv)
-fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(days-10):days,])
-summary(fit1)
+beta.model<-glm(beta_vec~bs(time,3),weights = 1/sd_vec,family=Gamma)
+forecast=14
+# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
+pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
+date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
+beta.predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
+
+r0.predict<-beta.predict
+r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14
+# plot the points (actual observations), regression line, and confidence interval
+p <- ggplot(r0.predict, aes(date,beta_vec))
+p <- p + geom_point() +labs(x="time from t0",y="R0 value") 
+p <- p + geom_line(aes(date,fit))
+p <- p + geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)
+p
 ```
 
-The slope coefficient estimated in the linear regression model can be used to estimate R0.
+R0 passes from values greater than 4 in the initial phase to values <1 at the ending of the 14-days forecast.
 
-R0=1+b*incubation period.
-
-The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
-Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
-In the calculation we considered an incubation period of 14 days for two reasons: 1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14.
-A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider in this phase.
-```{r r0 plot, echo=TRUE}
-slope <-coef(summary(fit1))[2,1]; slope
-slope.se <- coef(summary(fit1))[2,2]; slope.se
-### R0 estimates and 95% IC 
-R_0=slope*14+1;R_0
-(slope+c(-1,1)*1.96*slope.se)*14+1
-```
 We want to make a short term forecast (14 days) with 3 scenario:
 
--Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
+-Scenario 1: 10 exposed people for each COVID-19 case (no restrictions made or even no effects)
 
--Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% exposed people)
+-Scenario 2: 5 exposed people for each COVID-19 case (-50% exposed people)
 
--Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)
+-Scenario 3: 3 exposed people for each COVID-19 case  (-70% exposed people)
 
-We fix a series of initial parameters:
+We use a SEIR model fixing a series of initial parameters:
+
 - I0: initial number of COVID-19 cases  
 - R0: initial number of recovered  
-- beta: the quantity connected to R0  
+- beta: the quantity connected to R0 is considered to vary according the previous estimation
 - N: Italian population   
@@
 R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
-#beta 
-beta0<-R_0/(14)
 # italian poulation
@@
 <img src="http://www.public.asu.edu/~hnesse/classes/seireqn.png"/>  
+<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>
 
+where the parameter beta here is time dependent, as estimated before.
 ```{r  scenario plot, echo=TRUE}
@@
 forecast<-14
-parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
+seir1<-seir2<-seir3<-NULL
+for(i in 1:forecast){
+parameters <- c(mu = mu0, beta = beta.predict$fit[days+i], sigma = sigma0, gamma = 1/duration)
 f1<-10
-initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
-seir1 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
-parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
+if( i==1) initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
+if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
+seir1_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
+seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
 f2<-5
-initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
-seir2 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
-parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
-f3<-5
-initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
-seir3 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
+if( i==1) initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
+if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
+seir2_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
+seir2 <- rbind(seir2,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
+f3<-3
+if( i==1) initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
+if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])
+seir3_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
+seir3 <- rbind(seir3,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
+}
 
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-plot(date,c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N),type="l",ylab="Cases",xlab="time",main="Infected")
-lines(date,c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N),col=2)
-lines(date,c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N),col=3)
+plot(date,c(dat_csv$totale_attualmente_positivi,seir1$I*N),type="l",ylab="Cases",xlab="time",main="Infected")
+lines(date,c(dat_csv$totale_attualmente_positivi,seir2$I*N),col=2)
+lines(date,c(dat_csv$totale_attualmente_positivi,seir3$I*N),col=3)
 lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
@@
 The 3 scenarios show different numbers. 
-If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$results$I*N)[length(seir2$results$I)]`).
+If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$I*N)[length(seir2$I)]`).
 
 In the next plot the cumulative number of infected.  
-At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r ((seir2$results$I+seir3$results$R)*N)[length(seir2$results$I)]`).
+At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r ((seir2$I+seir3$R)*N)[length(seir2$I)]`).
 
@@
 
-plot(date,c(dat_csv$totale_casi,(seir1$results$I[-1]+seir1$results$R[-1])*N),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
-lines(date,c(dat_csv$totale_casi,(seir2$results$I[-1]+seir2$results$R[-1])*N),col=2)
-lines(date,c(dat_csv$totale_casi,(seir3$results$I[-1]+seir3$results$R[-1])*N),col=3)
+plot(date,c(dat_csv$totale_casi,(seir1$I+seir1$R)*N),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
+lines(date,c(dat_csv$totale_casi,(seir2$I+seir2$R)*N),col=2)
+lines(date,c(dat_csv$totale_casi,(seir3$I+seir3$R)*N),col=3)
 lines(date[1:dim(dat_csv)[1]],(dat_csv$totale_casi),lwd=2)
@@
 ```
-
-We make an effort to estimate the effect of restriction measures took by the Italian Government (11 March 2020). 
-After 14 days  (`r date[length(date)]`) we have the current status in the second scenario:  
-- I14=(`r (seir2$results$I*N)[length(seir2$results$I)]`);  
-- R14=(`r (seir2$results$R*N)[length(seir2$results$R)]`);  
-- S14=(`r (seir2$results$S*N)[length(seir2$results$S)]`);  
-- E14=(`r (seir2$results$E*N)[length(seir2$results$E)]`);
-
-We reduce R0 to 1.10 and the number of exposed to the number of infected.
-We estimate the trend of the number infected, recovered and the total cases for other 14 days forward.
-
-```{r  scenario plot2, echo=TRUE}
-I14=seir2$results$I[length(seir2$results$I)]
-R14=seir2$results$R[length(seir2$results$I)]
-S14=seir2$results$S[length(seir2$results$I)]
-E14=seir2$results$E[length(seir2$results$I)]
-forecast2<-14
-beta14<-1.1/14
-parameters <- c(mu = mu0, beta = beta14, sigma = sigma0, gamma = 1/duration)
-initials <- c(S = 1-I14*2-R14, E = I14, I = I14, R = R14)
-seir2_2 <- SEIR(pars = parameters, init = initials, time = 0:forecast2)
-
-date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast2+forecast-1+dim(dat_csv)[1],1)
-plot(date,c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N,seir2_2$results$I[-1]*N),type="l",ylab="Cases",xlab="time",ylim=c(0,((seir2_2$results$I+seir2_2$results$R[-1])*N)[forecast2]),col=2)
-lines(date,c(dat_csv$dimessi_guariti,seir2$results$R[-1]*N,seir2_2$results$R[-1]*N),col=3)
-lines(date,c(dat_csv$totale_casi,(seir2$results$I[-1]+seir2$results$R[-1])*N,c(seir2_2$results$I[-1]+seir2_2$results$R[-1])*N),col=1)
-
-legend("topleft",c("All cases","Infected","Recovered"),lty=1,col=1:3)
-```

### 2020-03-14 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

--- 2020-03-12 draft_analysis_Veneto.Rmd
@@
 ---
-title: "COVID19 - Forecast analysis"
+title: "COVID19 - Forecast analysis - time dependent SEIR model - Veneto"
 author: "PG"
-date: "3/11/2020"
+date: "3/13/2020"
 output: rmarkdown::github_document
@@
 rm(list=ls())
-###import italian dataset updated 10 March 2020
+###import italian dataset updated 13 March 2020 - 
 dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",header=T)
+#we restricted the analysis to the Veneto Region
 dat_csv<-dat_csv[dat_csv$codice_regione==5,]
@@
 plot(dat_csv$data,dat_csv$totale_attualmente_positivi,ylab="Infected",xlab="Date")
+days<-dim(dat_csv)[1]
 ```
 
-fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
+## The COVID dataset
 
-```{r cars}
-head(dat_csv)
-fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
-summary(fit1)
+The present analysis used the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19.
+We used a SEIR model to predict the trend of the actual COVID19 epidemic in Italy.
+We estimated the R0 parameter by means of a linear regression model as reported in https://kingaa.github.io/clim-dis/parest/parest.html
+
+The actual status is
+
+```{r plots}
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+df <- dat_csv %>%
+  select(data, ricoverati_con_sintomi, terapia_intensiva, 
+         totale_ospedalizzati, isolamento_domiciliare, 
+         totale_attualmente_positivi, nuovi_attualmente_positivi, 
+         dimessi_guariti, deceduti, totale_casi) %>%
+  gather(key = "variable", value = "value", -data)
+head(df, 3)
+
+ggplot(df, aes(x = as.Date(data), y = value)) + 
+  geom_line(aes(color = variable), size = 1) 
   ```
 
-Estimates
+The plot shows an exponential grow.
+We estimate the R0 parameter by means of a linear model.
+R0 indicates how contagious an infectious disease is.
+It is also referred to as "the reproduction number" of COVID19.
 
-```{r model plot, echo=TRUE}
-plot(dat_csv$t,log(dat_csv$totale_attualmente_positivi),ylab="log cases",xlab="time")
-abline(coef(summary(fit1))[,1])
+#estimate r0
+#see https://kingaa.github.io/clim-dis/parest/parest.html
+#calculate r0 based in the last XX observation 
+```{r R0 trend}
+beta_vec<-NULL
+sd_vec<-NULL
+for (i in 1:(days-4)){
+fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
+beta_vec<-c(beta_vec,coef(fit)[2])
+sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
+}
+
+label <-days:5
+mean  <- (beta_vec*14)
+lower <- ((beta_vec-1.96*sd_vec)*14)
+upper <- ((beta_vec+1.96*sd_vec)*14)
+
+df <- data.frame(label, mean, lower, upper)
+df$label <- factor(df$label, levels=rev(df$label))
+
+library(ggplot2)
+fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
+  geom_pointrange() + 
+  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
+  coord_flip() +  # flip coordinates (puts labels on y axis)
+  xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
+  theme_bw()  # use a white background
+print(fp)
+
 ```
-The slope coefficient estimated in the linear regression model can be used to estimate R0, the number of 
-```{r r0 plot, echo=TRUE}
-slope <-coef(summary(fit1))[2,1]; slope
-slope.se <- coef(summary(fit1))[2,2]; slope.se
-### R0 estimates and 95%IC 
-### I have used 14 for infection time, but it has a bimodal distribution (tested vs non tested)
-R_0=slope*14+1;R_0
-(slope+c(-1,1)*1.96*slope.se)*14+1
+
+$$Y_t= a + \beta  t +e_t  $$
+
+where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.  
+The slope coefficient estimated in the linear regression model can be used to estimate R0.
+
+R0=slope*incubation period.
+
+The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
+Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/, but the incubation period is used for epidemic diseases that causes the immediate isolation of the infected subject.
+
+In the calculation we considered an "incubation period" of 14 days for two reasons: 1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14.
+A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider in this phase.
+
+We calculate several R0 values, each one based on a different number of days before the last day. 
+The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponetial increase.  
+However,R0 is going to decrease in the next days. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
+We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value greater than 0. 
+```{r R0 forecast}
+library(splines)
+time<-1:(days-4)
+
+beta.model<-glm(beta_vec~bs(time,3),weights = 1/sd_vec,family=Gamma)
+forecast=14
+# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
+pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
+date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
+beta.predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
+
+r0.predict<-beta.predict
+r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14
+# plot the points (actual observations), regression line, and confidence interval
+p <- ggplot(r0.predict, aes(date,beta_vec))
+p <- p + geom_point() +labs(x="time from t0",y="R0 value") 
+p <- p + geom_line(aes(date,fit))
+p <- p + geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)
+p
 ```
+
+R0 passes from values around to 3 in the initial phase to values <1 at the ending of the 14-days forecast.
+
 We want to make a short term forecast (14 days) with 3 scenario:
--Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
--Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% COVID19 contagious power)
--Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)
 
-We fix a series of initial parameters:
--I0: initial number of COVID-19 cases  
+-Scenario 1: 10 exposed people for each COVID-19 case (no restrictions made or even no effects)
+
+-Scenario 2: 5 exposed people for each COVID-19 case (-50% exposed people)
+
+-Scenario 3: 3 exposed people for each COVID-19 case  (-70% exposed people)
+
+We use a SEIR model fixing a series of initial parameters:
+
+- I0: initial number of COVID-19 cases  
 - R0: initial number of recovered  
-- beta: the quantity connected to R0  
-- N: Italian population   
+- beta: the quantity connected to R0 is considered to vary according the previous estimation
+- N:  population   of the Veneto Region
 - duration: infection duration of COVID-19   
@@
 - mu0: the overall mortality rate
+
 ```{r initation plot, echo=TRUE}
@@
 R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
-#beta 
-beta0<-R_0/(14)
-# Veneto population
-N=4905000
+# italian poulation
+N=4905854
 # duration of COVID19 
@@
 ```
-We use the library(EpiDynamics)
-```{r first scenario plot, echo=TRUE}
+We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:
+
+<img src="http://www.public.asu.edu/~hnesse/classes/seireqn.png"/>  
+<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>
+
+where the parameter beta here is time dependent, as estimated before.
+```{r  scenario plot, echo=TRUE}
 library(EpiDynamics)
+
+# average number of single connections of an infected person
+# less contacts, less probability of new infections
+# we keep constant the other parameters
 forecast<-14
-parameters <- c(mu = mu0, beta = beta0, sigma = sigma0, gamma = 1/duration)
+seir1<-seir2<-seir3<-NULL
+for(i in 1:forecast){
+parameters <- c(mu = mu0, beta = beta.predict$fit[days+i], sigma = sigma0, gamma = 1/duration)
 f1<-10
-initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
-seir1 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
-parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
-f2<-10
-initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
-seir2 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
-parameters <- c(mu = mu0, beta = beta0*1/2, sigma = sigma0, gamma = 1/duration)
-f3<-5
-initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
-seir3 <- SEIR(pars = parameters, init = initials, time = 0:forecast)
+if( i==1) initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
+if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
+seir1_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
+seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
+f2<-5
+if( i==1) initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
+if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
+seir2_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
+seir2 <- rbind(seir2,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
+f3<-3
+if( i==1) initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
+if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])
+seir3_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
+seir3 <- rbind(seir3,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
+}
 
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-plot(date,c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N),type="l",ylab="Cases",xlab="time",main="Infected")
-lines(date,c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N),col=2)
-lines(date,c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N),col=3)
+plot(date,c(dat_csv$totale_attualmente_positivi,seir1$I*N),type="l",ylab="Cases",xlab="time",main="Infected")
+lines(date,c(dat_csv$totale_attualmente_positivi,seir2$I*N),col=2)
+lines(date,c(dat_csv$totale_attualmente_positivi,seir3$I*N),col=3)
 lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
@@
 ```
-The 3 scenarios show how measures of restriction can help to reduce the number of infected.  
-At the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$results$I*N)[length(seir2$results$I)]`).
+
+The 3 scenarios show different numbers. 
+If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$I*N)[length(seir2$I)]`).
+
 In the next plot the cumulative number of infected.  
+At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r ((seir2$I+seir3$R)*N)[length(seir2$I)]`).
+
 ```{r  cumulative plot, echo=TRUE}
 
-plot(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir1$results$I[-1]*N)),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
-lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir2$results$I[-1]*N)),col=2)
-lines(date,cumsum(c(dat_csv$totale_attualmente_positivi,seir3$results$I[-1]*N)),col=3)
-lines(date[1:dim(dat_csv)[1]],cumsum(dat_csv$totale_attualmente_positivi),lwd=2)
+plot(date,c(dat_csv$totale_casi,(seir1$I+seir1$R)*N),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
+lines(date,c(dat_csv$totale_casi,(seir2$I+seir2$R)*N),col=2)
+lines(date,c(dat_csv$totale_casi,(seir3$I+seir3$R)*N),col=3)
+lines(date[1:dim(dat_csv)[1]],(dat_csv$totale_casi),lwd=2)
 legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)

### 2020-03-14 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 rm(list=ls())
-###import italian dataset updated 10 March 2020
+###import italian dataset updated 13 March 2020
 dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
@@
 dat_csv$t<-1:days
-
+names(dat_csv)
 ```
@@
 p <- ggplot(r0.predict, aes(date,beta_vec))
-p <- p + geom_point() +labs(x="time from t0",y="R0 value") 
+p <- p + geom_point() +labs(x="Date",y="R0 value") 
 p <- p + geom_line(aes(date,fit))

### 2020-03-14 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 ---
+## The COVID dataset
 
-```{r setup, include=FALSE}
+The present analysis started from the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19, database provided by the Italian Civil Protection.  
+```{r setup}
 rm(list=ls())
@@
 dat_csv$t<-1:days
+#database contains  the following variables
 names(dat_csv)
+# The total number of epidemic day is
+days
+
 ```
 
-## The COVID dataset
+With the aim to predict the future number of COVID19 cases on the basis of the actual data, we used a SEIR model applied to the COVID19 epidemic in Italy.
+A first parameter to estimate is R0, the reproduction number (https://en.wikipedia.org/wiki/Basic_reproduction_number).
+R0 indicates how contagious an infectious disease is.
+It is also referred to as "the reproduction number" of COVID19.
 
-The present analysis used the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19.
-We used a SEIR model to predict the trend of the actual COVID19 epidemic in Italy.
-We estimated the R0 parameter by means of a linear regression model as reported in https://kingaa.github.io/clim-dis/parest/parest.html
+The R0 value has been estimated by means of a linear regression model as reported in https://kingaa.github.io/clim-dis/parest/parest.html.
 
-The actual status is
+The following plot presented the actual status.
 
@@
 
-The plot shows an exponential grow.
+The plot shows an exponential grow of cases.\\
+
 We estimate the R0 parameter by means of a linear model.
-R0 indicates how contagious an infectious disease is.
-It is also referred to as "the reproduction number" of COVID19.
 
-#estimate r0
-#see https://kingaa.github.io/clim-dis/parest/parest.html
+Y_t= a + beta * t +e_t
+
+where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.
+
+The slope coefficient is used to estimate R0 as in the following formula:
+
+R0=beta*incubation period.
+
+The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
+Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
+However, the incubation period is used for epidemic diseases that causes the immediate home isolation of infected subjects.
+
+In the calculation we considered an "incubation period" of 14 days for two reasons:  
+1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 
+2) 14 days is the worst scenario because in this period low impact COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).
+
+We calculate several R0 values, each one based on a different number of days before the last day, in order to assess if the R0 trend is decreasing (how is expected to be).
+
+```{r R0 trend}
 #calculate r0 based in the last XX observation 
-```{r R0 trend}
+#vector for beta and standard deviation
 beta_vec<-NULL
 sd_vec<-NULL
+#for cycle for all days to only the last 5.
 for (i in 1:(days-4)){
@@
 label <-days:5
+
 mean  <- (beta_vec*14)
@@
 
-df <- data.frame(label, mean, lower, upper)
+df <- data.frame(label, mean, lower, upper)[order(label),]
 df$label <- factor(df$label, levels=rev(df$label))
@@
   geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
-  coord_flip() +  # flip coordinates (puts labels on y axis)
   xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
-  theme_bw()  # use a white background
+  theme_bw() 
 print(fp)
@@
 
-Y_t= a + beta * t +e_t
-
-where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.  
-The slope coefficient estimated in the linear regression model can be used to estimate R0.
-
-R0=slope*incubation period.
-
-The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
-Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/, but the incubation period is used for epidemic diseases that causes the immediate isolation of the infected subject.
-
-In the calculation we considered an "incubation period" of 14 days for two reasons: 1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14.
-A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider in this phase.
-
-We calculate several R0 values, each one based on a different number of days before the last day. 
-The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponetial increase.  
-However,R0 is going to decrease in the next days. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
-We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing is a value greater than 0. 
+The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponential increase.  
+However,  R0 is going to decrease in the next days. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
+We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0.
+The trend was not monotonic, we use a simple splines to increase the fitting of the model to the data.
 ```{r R0 forecast}
@@
 time<-1:(days-4)
-
+# splines with only 3 basis, weighted for standard deviation of estimates.
 beta.model<-glm(beta_vec~bs(time,3),weights = 1/sd_vec,family=Gamma)
@@
 
-R0 passes from values greater than 4 in the initial phase to values <1 at the ending of the 14-days forecast.
+R0 passes from a value of `r round(r0.predict[1,1],2)`   in the initial phase to an  estimated value of `r round(r0.predict[days+forecast,3],2)` at the ending of the 14-days forecast.
 
@@
 
-We use a SEIR model fixing a series of initial parameters:
-
+We made a forecast by means of a SEIR model fixing a series of initial status:  
+- S0=N, Italian population   
+- E= f * I0 (with f a fixed factor of the previous scenario)  
 - I0: initial number of COVID-19 cases  
 - R0: initial number of recovered
+
+and parameters:  
 - beta: the quantity connected to R0 is considered to vary according the previous estimation  
-- N: Italian population   
-- duration: infection duration of COVID-19   
+- gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
 - sigma0: the coronavirus transmission rate (half of flu epidemic)   
@@
 
-where the parameter beta here is time dependent, as estimated before.
+where the parameter beta here is time dependent, as estimated before by the gamma regression model.
 ```{r  scenario plot, echo=TRUE}
@@
 ```
+
+```{r  forecast plot, echo=TRUE}
+
+plot(date[(days+1):(days+forecast)],seir2$E*N,type="l",ylab="Cases",xlab="Date",main="14 days - Forecast Scenario 2",ylim=c(0,max(seir2$E*N,seir2$I*N,seir2$R*N )))
+lines(date[(days+1):(days+forecast)],seir2$I*N,col=2)
+lines(date[(days+1):(days+forecast)],seir2$R*N,col=3)
+legend("topleft",c("Exposed","Infected","Recovered"),lty=1,col=1:4)
+```

### 2020-03-14 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

@@
 
-```{r setup, include=FALSE}
+## The COVID dataset
+
+The present analysis used the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19.  
+We used a SEIR model to predict the trend of the actual COVID19 epidemic in the Veneto Region.
+
+```{r setup, include=TRUE}
 rm(list=ls())
@@
 ```
-
-## The COVID dataset
-
-The present analysis used the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19.
-We used a SEIR model to predict the trend of the actual COVID19 epidemic in Italy.
-We estimated the R0 parameter by means of a linear regression model as reported in https://kingaa.github.io/clim-dis/parest/parest.html
 
@@
 
-The plot shows an exponential grow.
+The plot shows an exponential grow of cases.\\
+
 We estimate the R0 parameter by means of a linear model.
-R0 indicates how contagious an infectious disease is.
-It is also referred to as "the reproduction number" of COVID19.
 
-#estimate r0
-#see https://kingaa.github.io/clim-dis/parest/parest.html
-#calculate r0 based in the last XX observation 
+Y_t= a + beta * t +e_t
+
+where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.
+
+The slope coefficient is used to estimate R0 as in the following formula:
+
+R0=beta*incubation period.
+
+The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
+Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
+However, the incubation period is used for epidemic diseases that causes the immediate home isolation of infected subjects.
+
+In the calculation we considered an "incubation period" of 14 days for two reasons:  
+1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 
+2) 14 days is the worst scenario because in this period low impact COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).
+
+We calculate several R0 values, each one based on a different number of days before the last day, in order to assess if the R0 trend is decreasing (how is expected to be).
+
 ```{r R0 trend}
@@
   geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
-  coord_flip() +  # flip coordinates (puts labels on y axis)
   xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
@@
 
-$$Y_t= a + \beta  t +e_t  $$
-
-where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.  
-The slope coefficient estimated in the linear regression model can be used to estimate R0.
-
-R0=slope*incubation period.
-
-The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
-Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/, but the incubation period is used for epidemic diseases that causes the immediate isolation of the infected subject.
-
-In the calculation we considered an "incubation period" of 14 days for two reasons: 1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14.
-A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider in this phase.
-
-We calculate several R0 values, each one based on a different number of days before the last day. 
-The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponetial increase.  
-However,R0 is going to decrease in the next days. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
-We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value greater than 0. 
+The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponential increase.  
+However,  R0 is going to decrease in the next days. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
+We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0.
+The trend was not monotonic, we use a simple splines to increase the fitting of the model to the data.
 ```{r R0 forecast}
@@
 
-R0 passes from values around to 3 in the initial phase to values <1 at the ending of the 14-days forecast.
+R0 passes from a value of `r round(r0.predict[1,1],2)`   in the initial phase to an  estimated value of `r round(r0.predict[days+forecast,3],2)` at the ending of the 14-days forecast.
 
@@
 
-We use a SEIR model fixing a series of initial parameters:
-
+We made a forecast by means of a SEIR model fixing a series of initial status:  
+- S0=N, Italian population   
+- E= f * I0 (with f a fixed factor of the previous scenario)  
 - I0: initial number of COVID-19 cases  
 - R0: initial number of recovered
+
+and parameters:  
 - beta: the quantity connected to R0 is considered to vary according the previous estimation  
-- N:  population   of the Veneto Region
-- duration: infection duration of COVID-19   
+- gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
 - sigma0: the coronavirus transmission rate (half of flu epidemic)   

### 2020-03-14 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 days
-
+plot(as.Date(substr(dat_csv$data,1,10)),dat_csv$totale_attualmente_positivi,ylab=" Infected",xlab="Date",type="l")
+points(as.Date(substr(dat_csv$data,1,10)),dat_csv$totale_attualmente_positivi)
 ```
 
-With the aim to predict the future number of COVID19 cases on the basis of the actual data, we used a SEIR model applied to the COVID19 epidemic in Italy.
+With the aim of predicting the future number of COVID19 cases on the basis of the actual data, we used a SEIR model applied to the COVID19 epidemic in Italy.
 A first parameter to estimate is R0, the reproduction number (https://en.wikipedia.org/wiki/Basic_reproduction_number).
@@
 
-where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.
+where $`Y_t`$ is the  number of infected at the time t, while b is beta, the slope of the regression line.
 
@@
 
-The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponential increase.  
-However,  R0 is going to decrease in the next days. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
+The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
 We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0.
@@
 
-We want to make a short term forecast (14 days) with 3 scenario:
+We want to make a short term forecast (14 days) with 3 scenario, based on the number of exposed people:
 
--Scenario 1: 10 exposed people for each COVID-19 case (no restrictions made or even no effects)
+-Scenario 1: 10 exposed people for each COVID-19 case (no home restrictions made or even no effects)
 
@@
 We made a forecast by means of a SEIR model fixing a series of initial status:  
-- S0=N, Italian population   
+- S0=N, the size of Italian population   
 - E= f * I0 (with f a fixed factor of the previous scenario)  
@@
 lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
-legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
+legend("topleft",c("first scenario - Exp=10*I","second scenario Exp=5*I","third scenario Exp=3*I"),lty=1,col=1:3)
 ```
@@
 lines(date[1:dim(dat_csv)[1]],(dat_csv$totale_casi),lwd=2)
-legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
+legend("topleft",c("Scenario 1 -Exp=10*I","Scenario 2 - Exp=5*I","Scenario 3- Exp=3*I"),lty=1,col=1:3)
 ```

### 2020-03-14 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

@@
 
-The present analysis used the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19.  
+The present analysis is based on the COVID19 dataset updated in https://github.com/pcm-dpc/COVID-19.  
+This dataset contained the time series of infected, recovered, deaths, etc...for each day after 24/02/2020.
 We used a SEIR model to predict the trend of the actual COVID19 epidemic in the Veneto Region.
@@
 dat_csv$t<-1:dim(dat_csv)[1]
-plot(dat_csv$data,dat_csv$totale_attualmente_positivi,ylab="Infected",xlab="Date")
+plot(as.Date(substr(dat_csv$data,1,10)),dat_csv$totale_attualmente_positivi,ylab=" Infected",xlab="Date",type="l")
+points(as.Date(substr(dat_csv$data,1,10)),dat_csv$totale_attualmente_positivi)
 days<-dim(dat_csv)[1]
@@
 
-The actual status is
+The actual time series are
 
@@
 
-where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.
+where $`Y_t`$ is the number of infected at the time t, while b is beta, the slope of the regression line.
 
@@
 
--Scenario 1: 10 exposed people for each COVID-19 case (no restrictions made or even no effects)
+-Scenario 1: 10 exposed people for each COVID-19 case (no home restrictions made or even no effects)
 
@@
 We made a forecast by means of a SEIR model fixing a series of initial status:  
-- S0=N, Italian population   
+- S0=N, the size of Veneto Region population 
 - E= f * I0 (with f a fixed factor of the previous scenario)  
@@
 R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
-# italian poulation
+# Veneto Region population
 N=4905854
-# duration of COVID19 
+# duration of COVID19 diseases before get recovered
 duration<-14
@@
 lines(date[1:dim(dat_csv)[1]],(dat_csv$totale_casi),lwd=2)
-legend("topleft",c("first scenario","second scenario","third scenario"),lty=1,col=1:3)
+legend("topleft",c("Scenario 1 -Exp=10*I","Scenario 2- Exp=5*I","Scenario 3- Exp=3*I"),lty=1,col=1:3)
 ```

### 2020-03-14 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

@@
 lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
-legend("topleft",c("first scenario - Exp=10*I","second scenario Exp=5*I","third scenario Exp=3*I"),lty=1,col=1:3)
+legend("topleft",c("Scenario 1 - Exp=10*I","Scenario - Exp=5*I","Scenario - Exp=3*I"),lty=1,col=1:3)
 ```
@@
 lines(date[1:dim(dat_csv)[1]],(dat_csv$totale_casi),lwd=2)
-legend("topleft",c("Scenario 1 -Exp=10*I","Scenario 2- Exp=5*I","Scenario 3- Exp=3*I"),lty=1,col=1:3)
+legend("topleft",c("Scenario 1 - Exp=10*I","Scenario 2 - Exp=5*I","Scenario 3 - Exp=3*I"),lty=1,col=1:3)
 ```

### 2020-03-15 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
 days
-plot(as.Date(substr(dat_csv$data,1,10)),dat_csv$totale_attualmente_positivi,ylab=" Infected",xlab="Date",type="l")
-points(as.Date(substr(dat_csv$data,1,10)),dat_csv$totale_attualmente_positivi)
 ```
@@
 
-We calculate several R0 values, each one based on a different number of days before the last day, in order to assess if the R0 trend is decreasing (how is expected to be).
+We calculate several R0 values, each one based on a mobile window of 5 days, that can be sufficient to estimate a local trend, in order to assess if the R0 trend is decreasing (how is expected to be).
 
 ```{r R0 trend}
-#calculate r0 based in the last XX observation 
+#calculate r0 based with a mobile window of 5 days
 #vector for beta and standard deviation
@@
 sd_vec<-NULL
-#for cycle for all days to only the last 5.
-for (i in 1:(days-4)){
-fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
+#for cycle for R0 estimates from days-2 to days+2
+for (i in 3:(days-2)){
+fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(i-2):(i+2),])
 beta_vec<-c(beta_vec,coef(fit)[2])
@@
 
-label <-days:5
+label<-as.Date(substr(dat_csv$data,1,10))[3:(days-2)]
 
@@
 
-df <- data.frame(label, mean, lower, upper)[order(label),]
-df$label <- factor(df$label, levels=rev(df$label))
+df <- data.frame(label, mean, lower, upper)
 
@@
   geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
-  xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
+  xlab("Date") + ylab("R0 Mean (95% CI)") +
   theme_bw() 
@@
 library(splines)
-time<-1:(days-4)
-# splines with only 3 basis, weighted for standard deviation of estimates.
-beta.model<-glm(beta_vec~bs(time,3),weights = 1/sd_vec,family=Gamma)
+time<-3:(days-2)
+beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=Gamma)
 forecast=14

### 2020-03-15 draft_analysis_Veneto.Rmd
Add home page, navigation and theme

@@
 
-We calculate several R0 values, each one based on a different number of days before the last day, in order to assess if the R0 trend is decreasing (how is expected to be).
+We calculate several R0 values, each one based on a mobile window of 5 days, that can be sufficient to estimate a local trend, in order to assess if the R0 trend is decreasing (how is expected to be).
 
 ```{r R0 trend}
+#calculate r0 based with a mobile window of 5 days
+#vector for beta and standard deviation
 beta_vec<-NULL
 sd_vec<-NULL
-for (i in 1:(days-4)){
-fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
+#for cycle for R0 estimates from days-2 to days+2
+for (i in 3:(days-2)){
+fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(i-2):(i+2),])
 beta_vec<-c(beta_vec,coef(fit)[2])
@@
 
-label <-days:5
+label<-as.Date(substr(dat_csv$data,1,10))[3:(days-2)]
+
 mean  <- (beta_vec*14)
@@
 upper <- ((beta_vec+1.96*sd_vec)*14)
-
-df <- data.frame(label, mean, lower, upper)[order(label),]
-df$label <- factor(df$label, levels=rev(df$label))
+df <- data.frame(label, mean, lower, upper)
 
@@
   geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
-  xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
-  theme_bw()  # use a white background
+  xlab("Date") + ylab("R0 Mean (95% CI)") +
+  theme_bw() 
 print(fp)
@@
 ```{r R0 forecast}
-library(splines)
-time<-1:(days-4)
-
-beta.model<-glm(beta_vec~bs(time,3),weights = 1/sd_vec,family=Gamma)
+time<-3:(days-2)
+beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=Gamma)
 forecast=14

### 2020-03-15 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
+```{r echo=FALSE}
+library(knitr)
+knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_knit$set(root.dir= normalizePath('..'))
+knitr::opts_chunk$set(error = FALSE)
+```
+
 ---
 title: "COVID19 - Forecast and predictions using a time dependent SEIR model"
-author: "PG"
-date: "3/13/2020"
-output: github_document
+author: "Paolo Girardi"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output:
+  html_document:
+  df_print: paged
+pdf_document: 
+  toc: yes
 ---
+
+<a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.
+
+# Disclaimer
+
+* We want to investigate the evolution of the coronavirus pandemic in Italy from a statistical perspective using aggregated data.
+
+* Our point of view is that of surveillance with the goal of detecting important changes in the underlying (random) process as soon as possible after it has occured.
+
+* We use data provided by  Italian Civil Protection Department
+
+* This document is in a draft mode, and it is continuously updated.
+
+* The layout of the draft must definitely be improved.
+
 ## The COVID dataset
 
-The present analysis started from the dataset on COVID19 updated in https://github.com/pcm-dpc/COVID-19, database provided by the Italian Civil Protection.  
-```{r setup}
+The present analysis started from the dataset on COVID19 updated in
+<https://github.com/pcm-dpc/COVID-19>, database provided by the Italian
+Civil Protection.
+
+# Software
+
+Install  packages `dygraphs`,  `xts` and `EpiDynamics` if not available
+
+```{r echo=TRUE}
+checkpackage <- function(package) {
+  if (!package %in% installed.packages()) install.packages(package)
+}
+checkpackage("dygraphs")
+checkpackage("xts")
+checkpackage("EpiDynamics")
+```
+and load them.
+```{r, warning=FALSE}
+library(dygraphs)
+library(xts)
+library(EpiDynamics)
+```
+
+# Source of the data
+
+Download the data from
+
+https://github.com/pcm-dpc/COVID-19/
+
+# Results
+
+## Load dataset
+```{r}
+
 rm(list=ls())
@@
 dat_csv$t<-1:days
-#database contains  the following variables
-names(dat_csv)
 # The total number of epidemic day is
 days
+
+```
+Several outcomes can be potentially monitored, that is
+
+```{r}
+names(dat_csv[,-c(1:2,13)])
+
 ```
 
-With the aim of predicting the future number of COVID19 cases on the basis of the actual data, we used a SEIR model applied to the COVID19 epidemic in Italy.
-A first parameter to estimate is R0, the reproduction number (https://en.wikipedia.org/wiki/Basic_reproduction_number).
-R0 indicates how contagious an infectious disease is.
-It is also referred to as "the reproduction number" of COVID19.
+It is worth noting that some outcomes present negative counts in some regions.
+It looks like some of these negative counts are redesignations.
+Outcomes presenting negative values cannot be analyzed using the proposed model.
 
-The R0 value has been estimated by means of a linear regression model as reported in https://kingaa.github.io/clim-dis/parest/parest.html.
+Then we extract the timeseries.
 
-The following plot presented the actual status.
-
-```{r plots}
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-df <- dat_csv %>%
-  select(data, ricoverati_con_sintomi, terapia_intensiva, 
-         totale_ospedalizzati, isolamento_domiciliare, 
-         totale_attualmente_positivi, nuovi_attualmente_positivi, 
-         dimessi_guariti, deceduti, totale_casi) %>%
-  gather(key = "variable", value = "value", -data)
-head(df, 3)
-
-ggplot(df, aes(x = as.Date(data), y = value)) + 
-  geom_line(aes(color = variable), size = 1) 
+```{r}
+myDateTimeStr1 <- dat_csv$data
+myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
+days_dy<-as.Date(myPOSIXct1)
+dat_csv_dy<-xts(dat_csv[,-(1:2)], order.by = days_dy, frequency = 7)
 ```
 
-The plot shows an exponential grow of cases.\\
+```{r}
+p <- dygraph(dat_csv_dy,main=paste("Italy",sep =""),xlab="Day",height=400,width=800) 
+p
+```
 
-We estimate the R0 parameter by means of a linear model.
+### The S(E)IR model (to be revised)
 
-Y_t= a + beta * t +e_t
+With the aim of predicting the future number of COVID19 cases on the
+basis of the actual data, we used a SEIR model applied to the COVID19
+epidemic in Italy.
 
-where $`Y_t`$ is the  number of infected at the time t, while b is beta, the slope of the regression line.
+We will consider  the classical [SIR model](https://en.wikipedia.org/wiki/Compartmental_models_in_epidemiology) [@Kermack1927].
 
-The slope coefficient is used to estimate R0 as in the following formula:
+The model divides a population of hosts into three classes:
+susceptible, infected, recovered.
+The model describes how the portion of the population in each of these classes changes with time.
+Births are modeled as flows from "elsewhere" into the susceptible class;
+deaths are modeled as flows from the $S$, $I$, or $R$ compartment into "elsewhere".
+If $S$, $I$, and $R$ refer to the numbers of individuals in each compartment, then these **state variables** change according to the following system of differential equations:
+$$\begin{aligned}
+\frac{d}{dt}S(t) &= B(t)-\lambda\,S(t)-\mu\,S(t)\\
+\frac{d}{dt}I(t) &= \lambda\,S(t)-\gamma\,I(t)-\mu\,I(t)\\
+\frac{d}{dt}R(t) &= \gamma\,I(t)-\mu\,R(t).\\
+\end{aligned}$$
+Here, $B$ is the crude birth rate (births per unit time), $\mu$ is the death rate and $\gamma$ is the recovery rate.
+We'll assume that the force of infection, $\lambda$, for a constant population $N$
+$$\lambda = \beta\,\frac{I}{N},$$
+so that the risk of infection a susceptible faces is proportional to the *prevalence* (the fraction of the population that is infected).
+This is known as the assumption of frequency-dependent transmission.
 
-R0=beta*incubation period.
+# The reproduction number of COVID19.
 
-The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
+The number of infected individuals $I$ at time $t$ is approximately 
+$$I(t)\;\approx\;I_0\,e^{R_0\,(\gamma+\mu)\,t}$$
+where $I_0$ is the (small) number of infectives at time $0$, $\frac{1}{\gamma}$ is the infectious period, and $\frac{1}{\mu}$ is the host lifespan.
+
+$R_0$ is  the reproduction
+number (<https://en.wikipedia.org/wiki/Basic_reproduction_number>) and
+indicates how contagious an infectious disease is.
+
+Taking logs of both sides, we get
+
+$$\log{I}(t)\;\approx\;\log{I_0}+(R_0)\,(\gamma+\mu)\,t,$$
+which implies that a semi-log plot of $I$ vs $t$ should be approximately linear with a slope proportional to $R_0$ and the recovery rate.
+
+```{r}
+dat_csv_dy$log_totale_attualmente_positivi<-log(dat_csv_dy$totale_attualmente_positivi)
+p <- dygraph(dat_csv_dy$log_totale_attualmente_positivi,main=paste("Italy"),ylab="Log Infected case",xlab="Day",height=400,width=800) 
+p
+```
+
+We estimate the $R_0$ parameter in the linear model.
+
+$$
+\log(I(t))= \alpha + \beta  t +e_t
+$$
+
+The estimated slope coefficient $\hat\beta$ is used to estimate $R_0$ as in the following formula:
+
+$$\widehat\beta=(\widehat{R_0})\,(\gamma+\mu)$$
+The parameter $\mu$<<$\gamma$ and it can not be considered. As consequence, R0 can be estimated as follows
+$$\hat{R_0}=\frac{\hat{\beta}}{\gamma}
+$$
+
+The incubation period $1/ \gamma$ for the coronavirus is in mean 5.1 days with a range from 2-14 days.
 Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.

### 2020-03-15 draft_analysis_Italy.Rmd
Add home page, navigation and theme

@@
+---
+title: "COVID19 - Forecast and predictions using a time dependent SEIR model"
+author: "Paolo Girardi"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: rmarkdown::github_document
+always_allow_html: yes
+---
 ```{r echo=FALSE}
@@
 ```
-
----
-title: "COVID19 - Forecast and predictions using a time dependent SEIR model"
-author: "Paolo Girardi"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:
-  html_document:
-  df_print: paged
-pdf_document: 
-  toc: yes
----
-
 <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.
@@
 checkpackage("EpiDynamics")
+checkpackage("webshot")
 ```
@@
 library(EpiDynamics)
+library(webshot)
+install_phantomjs()
 ```
@@
 The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
-We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0.
-The trend was not monotonic, we use a simple splines to increase the fitting of the model to the data.
+We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Log-normal distribution for the beta (the slope) forcing its value to be greater than 0.
 ```{r R0 forecast}
-library(splines)
+
 time<-3:(days-2)
-beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=Gamma)
+beta.model<-glm(log(beta_vec)~time,weights = 1/sd_vec,family=gaussian)
 forecast=14
@@
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-beta.predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
-
+predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
+beta.predict<-predict 
+beta.predict[,3:5]<-exp(predict[,3:5])
 r0.predict<-beta.predict

### 2020-03-15 draft_analysis_Italy_html.Rmd
Add home page, navigation and theme

--- 2020-03-15 draft_analysis_Italy.Rmd
@@
----
-title: "COVID19 - Forecast and predictions using a time dependent SEIR model"
-author: "Paolo Girardi"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: rmarkdown::github_document
-always_allow_html: yes
----
 ```{r echo=FALSE}
@@
 ```
+---
+title: "COVID19 - Forecast and predictions using a time dependent SEIR model"
+author: "Paolo Girardi"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output:html_document:
+  df_print: paged
+pdf_document: 
+  toc: yes
+---
+
 <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.
@@
 checkpackage("webshot")
+checkpackage("bsts")
 ```
@@
 library(webshot)
-install_phantomjs()
+library(bsts)
 ```
@@
 rm(list=ls())
-###import italian dataset updated 13 March 2020
+###import italian updated dataset 
   dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
@@
 days_dy<-as.Date(myPOSIXct1)
-dat_csv_dy<-xts(dat_csv[,-(1:2)], order.by = days_dy, frequency = 7)
+dat_csv_dy<-xts(dat_csv[,-c(1:2,12:13)], order.by = days_dy, frequency = 7)
 ```
@@
 
-where the parameter beta here is time dependent, as estimated before by the gamma regression model.
-```{r  scenario plot, echo=TRUE}
+We perfom simulation on the trend of beta coefficient by means of a Bayesian Structural Time Series using the library bsts of R.   
+We made a forecast forward of 14 days.
+
+```{r bsts model fit and prediction, echo=TRUE}
+# Bayesian Structural Time Series
+ss <- AddLocalLinearTrend(list(), log(beta_vec))
+model1 <- bsts(log(beta_vec),
+               state.specification = ss,
+               niter = 200)
+par(mfrow = c(1,1))
+plot(model1, "components", ylab="Log beta coefficient", xlab="days")
+#previsioni
+pred1 <- predict(model1, horizon = 16, burn = 100)
+par(mfrow = c(1,1))
+plot(pred1 , ylab="Log beta coefficient",main="Data and predictions")
+# matrix of beta coefficients
+coef<-(exp(pred1$distribution)[,3:16])
+par(mfrow = c(1,1))
+```
+For each vectors of simulated beta coefficients we perform a SEIR model (one for each scenario).  
+We save the results and plot a credible interval at 90% (5-95% percentile).
+
+```{r  scenario plot , echo=TRUE}
 library(EpiDynamics)
-
+seir1_sim<-seir2_sim<-seir3_sim<-NULL
+for(s in 1:dim(coef)[1]){
   # average number of single connections of an infected person
@@
   for(i in 1:forecast){
-parameters <- c(mu = mu0, beta = beta.predict$fit[days+i], sigma = sigma0, gamma = 1/duration)
+    parameters <- c(mu = mu0, beta = matrix(coef[s,i]), sigma = sigma0, gamma = 1/duration)
     f1<-10
@@
   }
+  seir1_sim<-rbind(seir1_sim,cbind(rep(s,forecast),seir1))
+  seir2_sim<-rbind(seir2_sim,cbind(rep(s,forecast),seir2))
+  seir3_sim<-rbind(seir3_sim,cbind(rep(s,forecast),seir3))
+}
+seir1_sim[,2]<-seir2_sim[,2]<-seir3_sim[,2]<-rep(1:forecast,dim(coef)[1])
+colnames(seir1_sim)[1]<-colnames(seir2_sim)[1]<-colnames(seir3_sim)[1]<-"sim"
 
-date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-plot(date,c(dat_csv$totale_attualmente_positivi,seir1$I*N),type="l",ylab="Cases",xlab="time",main="Infected")
-lines(date,c(dat_csv$totale_attualmente_positivi,seir2$I*N),col=2)
-lines(date,c(dat_csv$totale_attualmente_positivi,seir3$I*N),col=3)
-lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
-legend("topleft",c("first scenario - Exp=10*I","second scenario Exp=5*I","third scenario Exp=3*I"),lty=1,col=1:3)
+### confidence limits
+I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
+I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.05)
+I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.95)
+
+days.before<-date[1:days]
+days.ahead<-date[(days+1):(days+forecast)]
+step.ahead<-forecast+1
+mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+counts<-mu.med
+mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
+p <- dygraph(mu,main=paste("Italy: Scenario 1 Credible (confidence ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
+p
+
+### confidence limits - scenario 2
+I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
+I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.05)
+I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.95)
+
+mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+counts<-mu.med
+mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
+p <- dygraph(mu,main=paste("Italy: Scenario 2 Credible (confidence ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
+p
+
+### confidence limits - scenario 3
+I_seir_med<-tapply(seir3_sim$I,seir3_sim$time,median)
+I_seir_lwr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=0.05)
+I_seir_upr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=0.95)
+
+mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+counts<-mu.med
+mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
+p <- dygraph(mu,main=paste("Italy: Scenario 3 Credible (confidence ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
+p
 ```
@@
 The 3 scenarios show different numbers. 
-If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (seir2$I*N)[length(seir2$I)]`).
+If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N)[forecast]`).
 
 In the next plot the cumulative number of infected.  
-At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r ((seir2$I+seir3$R)*N)[length(seir2$I)]`).
-
-```{r  cumulative plot, echo=TRUE}
-
-plot(date,c(dat_csv$totale_casi,(seir1$I+seir1$R)*N),type="l",ylab="Cases",xlab="time",main="Cumulative Infected")
-lines(date,c(dat_csv$totale_casi,(seir2$I+seir2$R)*N),col=2)
-lines(date,c(dat_csv$totale_casi,(seir3$I+seir3$R)*N),col=3)
-lines(date[1:dim(dat_csv)[1]],(dat_csv$totale_casi),lwd=2)
-legend("topleft",c("Scenario 1 -Exp=10*I","Scenario 2 - Exp=5*I","Scenario 3- Exp=3*I"),lty=1,col=1:3)
-```
-
-```{r  forecast plot, echo=TRUE}
-
-plot(date[(days+1):(days+forecast)],seir2$E*N,type="l",ylab="Cases",xlab="Date",main="14 days - Forecast Scenario 2",ylim=c(0,max(seir2$E*N,seir2$I*N,seir2$R*N )))
-lines(date[(days+1):(days+forecast)],seir2$I*N,col=2)
-lines(date[(days+1):(days+forecast)],seir2$R*N,col=3)
-legend("topleft",c("Exposed","Infected","Recovered"),lty=1,col=1:4)
-```
+At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N+tapply(seir2_sim$R,seir2_sim$time,median)*N)[forecast]`).

### 2020-03-15 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

--- 2020-03-15 draft_analysis_Italy.Rmd
@@
 
+NB:  the file draft_analysis_italy_html.Rmd performs the same analysis enabling Javascript Pictures.
 ## The COVID dataset
@@
 checkpackage("webshot")
+checkpackage("bsts")
+checkpackage("ggplot2")
 ```
@@
 library(webshot)
-install_phantomjs()
+library(bsts)
+library(EpiDynamics)
+library(ggplot2)
 ```
@@
 We calculate several R0 values, each one based on a mobile window of 5 days, that can be sufficient to estimate a local trend, in order to assess if the R0 trend is decreasing (how is expected to be).
+In this way, the R0 for the first and the last two days of observation is impossibile to estimate.
 
@@
 
-library(ggplot2)
 fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
@@
 The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
-We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Log-normal distribution for the beta (the slope) forcing its value to be greater than 0.
+We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Log-normal distribution for the beta (the slope) and forcing its value to be greater than 0.
 ```{r R0 forecast}
@@
 <img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>  
+Respect to the previous formulation, the compartment of exposed people (E) was inserted between the susceptible and infected compartments.  
+(https://en.wikipedia.org/wiki/Bayesian_structural_time_series)
 
-where the parameter beta here is time dependent, as estimated before by the gamma regression model.
-```{r  scenario plot, echo=TRUE}
-library(EpiDynamics)
+We perfom simulation on the trend of beta coefficient by means of a Bayesian Structural Time Series using the library bsts of R.    
+We estimate a BSTS model specifing a local linear trend ans 1000 simulation.
 
+We made a forecast forward of 14 days dropping the first 100 simulation as burn-in.
+
+```{r bsts model fit and prediction, echo=TRUE}
+# Bayesian Structural Time Series
+set.seed(123)
+ss <- AddLocalLinearTrend(list(), log(beta_vec))
+model1 <- bsts(log(beta_vec),
+               state.specification = ss,
+               niter = 1000)
+par(mfrow = c(1,1))
+plot(model1, "components", ylab="Log beta coefficient", xlab="days")
+#previsioni
+pred1 <- predict(model1, horizon = 16, burn = 100)
+par(mfrow = c(1,1))
+plot(pred1 , ylab="Log beta coefficient",main="Data and predictions")
+# matrix of beta coefficients
+coef<-(exp(pred1$distribution)[,3:16])
+par(mfrow = c(1,1))
+```
+For each vectors of simulated beta coefficients we perform a SEIR model (one for each scenario).  
+We save the results and plot a credible interval at 90% (5-95% percentile).
+
+```{r  scenario plot , echo=TRUE}
+
+seir1_sim<-seir2_sim<-seir3_sim<-NULL
+for(s in 1:dim(coef)[1]){
   # average number of single connections of an infected person
@@
   }
+  seir1_sim<-rbind(seir1_sim,cbind(rep(s,forecast),seir1))
+  seir2_sim<-rbind(seir2_sim,cbind(rep(s,forecast),seir2))
+  seir3_sim<-rbind(seir3_sim,cbind(rep(s,forecast),seir3))
+}
+seir1_sim[,2]<-seir2_sim[,2]<-seir3_sim[,2]<-rep(1:forecast,dim(coef)[1])
+colnames(seir1_sim)[1]<-colnames(seir2_sim)[1]<-colnames(seir3_sim)[1]<-"sim"
 
-date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-plot(date,c(dat_csv$totale_attualmente_positivi,seir1$I*N),type="l",ylab="Cases",xlab="time",main="Infected")
-lines(date,c(dat_csv$totale_attualmente_positivi,seir2$I*N),col=2)
-lines(date,c(dat_csv$totale_attualmente_positivi,seir3$I*N),col=3)
-lines(date[1:dim(dat_csv)[1]],dat_csv$totale_attualmente_positivi,lwd=2)
-legend("topleft",c("first scenario - Exp=10*I","second scenario Exp=5*I","third scenario Exp=3*I"),lty=1,col=1:3)
+### confidence limits
+I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
+limits<-c(0.05,0.95)
+confidence<-diff(limits)
+I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=limits[1])
+I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=limits[2])
+
+days.before<-date[1:days]
+days.ahead<-date[(days+1):(days+forecast)]
+step.ahead<-forecast+1
+mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+counts<-mu.med
+mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
+p <- dygraph(mu,main=paste("Italy: Scenario 1 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
+p
+
+### confidence limits - scenario 2
+I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
+I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=limits[1])
+I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=limits[2])
+
+mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+counts<-mu.med
+mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
+p <- dygraph(mu,main=paste("Italy: Scenario 2 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
+p
+
+### confidence limits - scenario 3
+I_seir_med<-tapply(seir3_sim$I,seir3_sim$time,median)
+I_seir_lwr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=limits[1])
+I_seir_upr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=limits[2])
+
+mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+counts<-mu.med
+mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
+p <- dygraph(mu,main=paste("Italy: Scenario 3 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
+p
 ```

### 2020-03-15 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

--- 2020-03-15 draft_analysis_Italy_new.Rmd
@@
 ---
-title: "COVID19 - Forecast and predictions using a time dependent SEIR model"
+title: "COVID19 - Forecast and predictions using a time dependent SEIR model - Veneto Region"
 author: "Paolo Girardi"
@@
 
-* We want to investigate the evolution of the coronavirus pandemic in Italy from a statistical perspective using aggregated data.
+* We want to investigate the evolution of the coronavirus pandemic in the Veneto Region from a statistical perspective using aggregated data.
 
@@
 
-* We use data provided by  Italian Civil Protection Department
+* We use data provided by  Italian Civil Protection Department and the analysis was restricted to the Veneto Rego
 
@@
 
-NB:  the file draft_analysis_italy_html.Rmd performs the same analysis enabling Javascript Pictures.
+*NB: set the file output format to
+
+#output:html_document:  
+  df_print: paged  
+pdf_document:   
+  toc: yes
+
+which  performs the same analysis enabling Javascript Pictures.
+
 ## The COVID dataset
@@
 ```{r}
-
 rm(list=ls())
 ###import italian updated dataset 
-  dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
+dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",header=T)
+#we restricted the analysis to the Veneto Region
+dat_csv<-dat_csv[dat_csv$codice_regione==5,]
 days<-dim(dat_csv)[1]
@@
 days
-
 ```
@@
 days_dy<-as.Date(myPOSIXct1)
-dat_csv_dy<-xts(dat_csv[,-c(1:2,12:13)], order.by = days_dy, frequency = 7)
+dat_csv_dy<-xts(dat_csv[,-c(1:7,16:17)], order.by = days_dy, frequency = 7)
 ```
@@
 ```{r}
-p <- dygraph(dat_csv_dy,main=paste("Italy",sep =""),xlab="Day",height=400,width=800) 
+p <- dygraph(dat_csv_dy,main=paste("Veneto Region",sep =""),xlab="Day",height=400,width=800) 
 p
@@
 basis of the actual data, we used a SEIR model applied to the COVID19
-epidemic in Italy.
+epidemic to the Veneto Region in Italy
 
@@
 dat_csv_dy$log_totale_attualmente_positivi<-log(dat_csv_dy$totale_attualmente_positivi)
-p <- dygraph(dat_csv_dy$log_totale_attualmente_positivi,main=paste("Italy"),ylab="Log Infected case",xlab="Day",height=400,width=800) 
+p <- dygraph(dat_csv_dy$log_totale_attualmente_positivi,main=paste("Veneto Region"),ylab="Log Infected case",xlab="Day",height=400,width=800) 
 p
@@
 We made a forecast by means of a SEIR model fixing a series of initial status:  
-- S0=N, the size of Italian population   
+- S0=N, the size of Veneto Region population   
 - E= f * I0 (with f a fixed factor of the previous scenario)  
@@
 R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
-# italian poulation
-N=60480000
+# Veneto Region population
+N=4905854
 # duration of COVID19 
@@
 # Bayesian Structural Time Series
-set.seed(123)
 ss <- AddLocalLinearTrend(list(), log(beta_vec))
@@
 For each vectors of simulated beta coefficients we perform a SEIR model (one for each scenario).  
-We save the results and plot a credible interval at 90% (5-95% percentile).
+We save the results and plot a credible interval at 50% (25-75% percentile).
 
@@
 I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
-limits<-c(0.05,0.95)
-confidence<-diff(limits)
-I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=limits[1])
-I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=limits[2])
+I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.25)
+I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.75)
 
@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Italy: Scenario 1 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
@@
 I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
-I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=limits[1])
-I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=limits[2])
+I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.25)
+I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.75)
 
@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Italy: Scenario 2 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 2  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
@@
 I_seir_med<-tapply(seir3_sim$I,seir3_sim$time,median)
-I_seir_lwr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=limits[1])
-I_seir_upr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=limits[2])
+I_seir_lwr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=0.25)
+I_seir_upr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=0.75)
 
@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Italy: Scenario 3 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 3  (Credible Interval ",100*50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")

### 2020-03-17 BYM_model/INLA_def.Rmd
Add home page, navigation and theme

---
title: "COVID19 - Forecast and predictions using a BYM model - Veneto Region"
author: "Paolo Girardi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::github_document
always_allow_html: yes
---
```{r echo=FALSE}
#Load libraries
library(spdep)
library(maptools)
library(INLA)
library(rgdal)
library(foreign)
library(dygraphs)
library(xts)
```
################################ IMPORT DATASETS
# ISTAT Province dataset
```{r}
db_istat<-read.csv("Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
pop_provincia<-data.frame(codice_provincia=as.numeric(names(table(db_istat[,3]))),pop=tapply(db_istat[,20],db_istat[,3],sum))
###import international dataset
dat_int<-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv",header=T)
hubei<-as.numeric(dat_int[dat_int$Province.State=="Hubei",5:dim(dat_int)[2]])

###import updated italian dataset  for each province
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv",header=T)
dat_csv<-merge(dat_csv,pop_provincia,by=c("codice_provincia"),all.y=TRUE)
dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
#select a Region, in this example "Lombardia" Code region 3
Region<-3
dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
dat_csv_n$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
#### number of province
nprov<-length(table(dat_csv$denominazione_provincia)[table(dat_csv$denominazione_provincia)>0])

#order the obsercation for date and space, useful for INLA later....
dat_csv<-dat_csv[order(dat_csv$data),]
dat_csv<-dat_csv[order(dat_csv$codice_provincia),]
###space ID
dat_csv$ID<-as.factor(dat_csv$codice_provincia)
levels(dat_csv$ID)<-1:length(levels(dat_csv$ID))
dat_csv$ID<-as.numeric(dat_csv$ID)
#calculate new_cases
dat_csv$cumulative_cases<-dat_csv$totale_casi
dat_csv$totale_casi[dat_csv$data!="2020-02-24 18:00:00"]<-unlist(tapply(dat_csv$totale_casi,dat_csv$ID,diff))
dat_csv$totale_casi[dat_csv$totale_casi<0]<-0

#import dataset shapefile Italian province
nc.province <- readShapePoly(paste(wd,"/ProvCM01012019_g/ProvCM01012019_g_WGS84.shp",sep=""))
nc.province<-nc.province[nc.province$COD_REG==Region,]
nc.province<-nc.province[order(nc.province$COD_PROV),]
nc.province$ID<-as.factor(nc.province$COD_PROV)
levels(nc.province$ID)<-1:length(levels(nc.province$ID))
nc.province$ID<-as.numeric(nc.province$ID)
#Create adjacency matrix
nc.nb <- poly2nb(nc.province,snap=0.01)
nb2INLA(paste(wd,"nc.adj",sep=""), nc.nb)
file.adj <- paste(wd,"nc.adj",sep="")
################################
```

################################
#prior distribution for iid and besag components
```{r}
prior.iid = c(1,0.01)
prior.besag = c(1,0.001)
initial.iid = 4
initial.besag = 3
```
####COVID19 MAPS BYM with INLA
##########################indexing, for 
#### number of days
```{r}
dat_csv$t<-as.numeric(dat_csv$data)
days<-max(dat_csv$t)
dat_csv$t3<-dat_csv$t2<-dat_csv$t
#### province
dat_csv$ID2<-dat_csv$ID
##########################
formula.bym= totale_casi ~ 1+f(ID, model = "bym", graph = file.adj, param = c(prior.iid, prior.besag), initial = c(initial.iid, initial.besag))
fit_1<-inla(formula.bym, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_1)
##incidence rate ratio for each province
IRR_mean <- exp(fit_1$summary.random$ID$mean[1:nprov])
nc.province$IRR_mean<-IRR_mean
spplot(nc.province, c( "IRR_mean"))
```
#Pure temporal model (RW2)
RW2 temporal model
```{r}
Date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days-1,1)
formula_t = totale_casi ~ 1+f(t,model="rw2", constr = FALSE)+f(t2,model="iid")
fit_2<-inla(formula_t, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_2)
plot(Date,fit_2$summary.random$t$mean,ylab="Temp Coef",xlab="Date")
#estimating a linear model without the first obs
lm_fit<-lm(fit_2$summary.random$t$mean~I(1:days),subset=-1)
coef(lm_fit)[2]
plot(1:days,fit_2$summary.random$t$mean,ylab="Temp Coef",xlab="Date")
abline(lm_fit)
##Incidence rate ratios, the rate of increase od cases respect to time 1
plot(Date,exp(fit_2$summary.random$t$mean),ylab="Indicence of cases",xlab="Date")
```

#interaction model type 4 (complete time trend interaction to estimate specific curves for each province)
```{r}
####specification
formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
  f(t,model="rw2",constr = FALSE) +
  f(t2,model="iid") +
  f(ID2,model="besag", graph=file.adj,
    group=t3,
    control.group=list(model="rw2"))
#estimation
fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_st4)
```
#### components
```{r}
t_1<-fit_st4$summary.random$t$mean #overall trend
t_2<-fit_st4$summary.random$t2$mean #iid trend
s_1<-fit_st4$summary.random$ID$mean[1:nprov] # spatial CAR component
s_2<-fit_st4$summary.random$ID$mean[(nprov+1):(nprov*2)] # spatial idd
st<-fit_st4$summary.random$ID2$mean # spatio-temporal model with car and ar1 specification ```
## Mean IRR for each province
```{r}
nc.province$IRR_st<-exp(s_1+apply(t(matrix(st,nrow=nprov)),2,mean))
spplot(nc.province, c( "IRR_st"),main="IRR")
```
## Trends for province
```{r}
trends<-t_1+t(matrix(st,nrow=nprov))
# trends+ specific
matplot(trends)
#specific for each province
matplot(t(matrix(st,nrow=nprov)))
```

##################predictions 3 days forward of new CODID 19 cases
```{r}
Forecast=3
dat_csv2<-dat_csv 
dat_csv2$t<-rep(1:days+days,nprov)
dat_csv2$totale_casi<-NA
dat_csv2<-dat_csv2[dat_csv2$t<=(days+Forecast),]
dat_csv_n<-rbind(dat_csv,dat_csv2)
dat_csv_n$t3<-dat_csv_n$t2<-dat_csv_n$t
#introducting hubei starting on 22/01, quarantine started on 25/01
# applying a smoothing , data has errors
hubei_new_cases<-filter(diff(hubei), rep(1, 5))

plot(hubei/58.5,type="l",ylab="COVID19 cases for milion of inhabitant",ylim=c(0,1530),xlab="Days since over 100 cases")
lines(c(rep(NA,2),cumsum(tapply(dat_csv_n$totale_casi,dat_csv_n$t,sum))/10),col=2)
abline(v=2,col=1,lty=2)
abline(v=4,col=2,lty=2)
abline(v=10,col=3,lty=3)
legend("bottomright",c("Hubei","Lombardy"),lty=1,col=1:2)
text(10,1500,"Hubei quarantine")
text(12,1350,"Lodi home lock",col=2)
text(20,1200,"Lombardy home lock",col=3)
#province of lodi restrictive measure started on 24/02 lag 28 days, at the beginning of the time series
#other provinces on 11/03 lagged started their restrictions
hubei_lodi<-hubei_new_cases[1:(days+Forecast)]
hubei_others<-c(rep(NA,8),hubei_new_cases[1:(days+Forecast-8)])
dat_csv_n<-dat_csv_n[order(dat_csv_n$codice_provincia,dat_csv_n$t),]
dat_csv_n$hubei<-rep(hubei_others,12)
dat_csv_n$hubei[dat_csv_n$denominazione_provincia=="Lodi"]<-hubei_lodi
#overdispersion
dat_csv_n$over<-1:dim(dat_csv_n)[1]
```
We model the number of cases.

In the rw2 I set constr = FALSE is set to FALSE and that, for this reason, the intercept is not included in the linear predictor. 
```{r}
## include overdispersion parameter
formula.intIVn<- totale_casi ~ log(hubei+1)+log(pop)+
  f(t,model="rw2", constr = FALSE) +
  f(t2,model="iid") +f(over,model="iid") +
f(ID,model="bym",graph=file.adj) +
  f(ID2,model="besag", graph=file.adj,
    group=t3,
    control.group=list(model="rw2"))+ f(over,model="iid")

fit_st4n<-inla(formula.intIVn, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = 1))
summary(fit_st4n)
```

```{r}
#### components
t_1<-fit_st4n$summary.random$t$mean #overall trend
t_2<-fit_st4n$summary.random$t2$mean #iid trend
s_1<-fit_st4n$summary.random$ID$mean[1:nprov] # spatial CAR component
s_2<-fit_st4n$summary.random$ID$mean[(nprov+1):(nprov*2)] # spatial idd
st<-fit_st4n$summary.random$ID2$mean # spatio-temporal coefficients 
## Trends for province
Date_n<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days+Forecast-1,1)
trends<-t_1+t(matrix(st,nrow=nprov))
plot(Date_n,t_1,ylab="RW2 Time coefficient",xlab="Date")
matplot(t(matrix(st,nrow=nprov)),ylab="RW2 Time coefficient for Province",xlab="Date")
#extract number of cases for lodi
est_lodi<-fit_st4n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Lodi",]

days.before<-Date_n[1:days]
days.ahead<-Date_n[(days+1):(days+Forecast)]
mu.lower<-est_lodi$mean-1.96*est_lodi$sd
mu.lower[mu.lower<0]<-0
mu.upper<-est_lodi$mean+1.96*est_lodi$sd
mu.med<-xts(est_lodi$mean,order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
step.ahead<-Forecast
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Lodi (Credible Interval ",100*0.95,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

### 2020-03-17 BYM_model/INLA_def.R

@@ -1,0 +1,200 @@
+#Load libraries
+library(spdep)
+library(maptools)
+library(INLA)
+library(rgdal)
+library(foreign)
+library("dygraphs")
+library("xts")
+#define work directory
+################################
+rm(list=ls())
+wd<-"~/Dropbox/COVID19"
+setwd(wd)
+################################ IMPORT DATASETS
+# ISTAT Province dataset
+db_istat<-read.csv("Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
+pop_provincia<-data.frame(codice_provincia=as.numeric(names(table(db_istat[,3]))),pop=tapply(db_istat[,20],db_istat[,3],sum))
+###import international dataset
+dat_int<-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv",header=T)
+hubei<-as.numeric(dat_int[dat_int$Province.State=="Hubei",5:dim(dat_int)[2]])
+
+###import updated italian dataset  for each province
+dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv",header=T)
+dat_csv<-merge(dat_csv,pop_provincia,by=c("codice_provincia"),all.y=TRUE)
+dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
+#select a Region, in this example "Lombardia" Code region 3
+Region<-3
+dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
+dat_csv_n$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
+#### number of province
+nprov<-length(table(dat_csv$denominazione_provincia)[table(dat_csv$denominazione_provincia)>0])
+
+#order the obsercation for date and space, useful for INLA later....
+dat_csv<-dat_csv[order(dat_csv$data),]
+dat_csv<-dat_csv[order(dat_csv$codice_provincia),]
+###space ID
+dat_csv$ID<-as.factor(dat_csv$codice_provincia)
+levels(dat_csv$ID)<-1:length(levels(dat_csv$ID))
+dat_csv$ID<-as.numeric(dat_csv$ID)
+#calculate new_cases
[…]

### 2020-03-17 BYM_model/INLA_def.Rmd
Add home page, navigation and theme

@@
 ```{r}
-db_istat<-read.csv("Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
+db_istat<-read.csv("https://raw.githubusercontent.com/Paolin83/COVID19_SEIR_model/master/BYM_model/Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
 pop_provincia<-data.frame(codice_provincia=as.numeric(names(table(db_istat[,3]))),pop=tapply(db_istat[,20],db_istat[,3],sum))
@@
 dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
-dat_csv_n$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
+dat_csv$denominazione_provincia<-droplevels(dat_csv$denominazione_provincia)
 #### number of province
@@
 #import dataset shapefile Italian province
-nc.province <- readShapePoly(paste(wd,"/ProvCM01012019_g/ProvCM01012019_g_WGS84.shp",sep=""))
+
+nc.province <- readOGR("BYM_model/ProvCM01012019_g",layer ="ProvCM01012019_g_WGS84")
 nc.province<-nc.province[nc.province$COD_REG==Region,]

### 2020-03-17 BYM_model/INLA_def.Rmd
Add home page, navigation and theme

@@
 #import dataset shapefile Italian province
-
-nc.province <- readOGR("BYM_model/ProvCM01012019_g",layer ="ProvCM01012019_g_WGS84")
+nc.province <- readShapePoly("../BYM_model/ProvCM01012019_g/ProvCM01012019_g_WGS84.shp")
 nc.province<-nc.province[nc.province$COD_REG==Region,]
@@
 nc.nb <- poly2nb(nc.province,snap=0.01)
-nb2INLA(paste(wd,"nc.adj",sep=""), nc.nb)
-file.adj <- paste(wd,"nc.adj",sep="")
+nb2INLA("nc.adj", nc.nb)
+file.adj <- "nc.adj"
 ################################

### 2020-03-17 BYM_model/INLA_def_html.Rmd
Add home page, navigation and theme

---
title: "COVID19 - Forecast and predictions using a BYM model in Italy"
author: "Paolo Girardi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:html_document:  
  df_print: paged  
pdf_document:   
  toc: yes

always_allow_html: yes
---
```{r echo=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
knitr::opts_chunk$set(error = FALSE)
```
<a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.

# Disclaimer

* We want to investigate the evolution of the coronavirus pandemic in  Italy

* Our point of view is that of surveillance with the goal of detecting important changes in the underlying (random) process as soon as possible after it has occured.

* We use data provided by  Italian Civil Protection Department and the analysis was restricted to the Lombardy Region

* This document is in a draft mode, and it is continuously updated.

* The layout of the draft must definitely be improved.

NB: to compile this file please download the Italian Shapefile from the repository  
and set the working directory here  
```{r echo=TRUE}
setwd("/Users/Paolo/My_Website/public/COVID19_SEIR_model")  
```
## The COVID dataset

The present analysis started from the dataset on COVID19 updated in
<https://github.com/pcm-dpc/COVID-19>, database provided by the Italian
Civil Protection.

# Software

Install  packages `dygraphs`,  `xts` and `EpiDynamics` if not available

```{r echo=TRUE}
checkpackage <- function(package) {
  if (!package %in% installed.packages()) install.packages(package)
}
checkpackage("dygraphs")
checkpackage("xts")
checkpackage("ggplot2")
checkpackage("spdep")
checkpackage("maptools")
checkpackage("INLA")
checkpackage("rgdal")
```
and load them.
```{r, warning=FALSE}
#Load libraries
library("dygraphs")
library("xts")
library("ggplot2")
library("spdep")
library("maptools")
library("INLA")
library("rgdal")
```
# Datasets loading
```{r}
db_istat<-read.csv("https://raw.githubusercontent.com/Paolin83/COVID19_SEIR_model/master/BYM_model/Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
pop_provincia<-data.frame(codice_provincia=as.numeric(names(table(db_istat[,3]))),pop=tapply(db_istat[,20],db_istat[,3],sum))
###import international dataset
dat_int<-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv",header=T)
hubei<-as.numeric(dat_int[dat_int$Province.State=="Hubei",5:dim(dat_int)[2]])

###import updated italian dataset  for each province
dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv",header=T)
dat_csv<-merge(dat_csv,pop_provincia,by=c("codice_provincia"),all.y=TRUE)
dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
#select a Region, in this example "Lombardia" Code region 3
Region<-3
dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
dat_csv$denominazione_provincia<-droplevels(dat_csv$denominazione_provincia)
#### number of province
nprov<-length(table(dat_csv$denominazione_provincia)[table(dat_csv$denominazione_provincia)>0])

#order the obsercation for date and space, useful for INLA later....
dat_csv<-dat_csv[order(dat_csv$data),]
dat_csv<-dat_csv[order(dat_csv$codice_provincia),]
###space ID
dat_csv$ID<-as.factor(dat_csv$codice_provincia)
levels(dat_csv$ID)<-1:length(levels(dat_csv$ID))
dat_csv$ID<-as.numeric(dat_csv$ID)
#calculate new_cases Y ~ Poi(mu)
dat_csv$cumulative_cases<-dat_csv$totale_casi
dat_csv$totale_casi[dat_csv$data!="2020-02-24 18:00:00"]<-unlist(tapply(dat_csv$totale_casi,dat_csv$ID,diff))
dat_csv$totale_casi[dat_csv$totale_casi<0]<-0
#import dataset shapefile Italian province and adjancies matrix
#### IMPORTANT, SET THE Work Directory of the shapefile 
nc.province <- readShapePoly("BYM_model/ProvCM01012019_g/ProvCM01012019_g_WGS84.shp")
nc.province<-nc.province[nc.province$COD_REG==Region,]
nc.province<-nc.province[order(nc.province$COD_PROV),]
nc.province$ID<-as.factor(nc.province$COD_PROV)
levels(nc.province$ID)<-1:length(levels(nc.province$ID))
nc.province$ID<-as.numeric(nc.province$ID)
#Create adjacency matrix
nc.nb <- poly2nb(nc.province,snap=0.01)
nb2INLA("nc.adj", nc.nb)
file.adj <- "nc.adj"
################################
```

# Spatial analysis with R-INLA and BYM 
We modelled COVID 19 cases by means of a BYM (Besag, York and Molli\'e) model using an Integrated Nested Laplace Approximation (INLA).  
* Starting phase

For the $i$-th Nuts-3 Region (Italian province), the observed number of COVID19, $y_{i}$, was modelled as follows:

$$
  Y_i|\lambda_i\sim \mbox{Poisson}(\lambda_i), \qquad i=1,2,\ldots \\
$$

and $\lambda_i>0$. We modelled the incidence of COVID-19 by means of a BYM model including the population size as offset variable as:

$$
 log(\frac{\lambda_{i}}{N_{i}})=\alpha+\mu_i+\nu_i, \\
$$
where $\alpha$ is the intercept, $\mu_i$ and $\nu_i$ are two area specific effects with normal distribution modelled using an intrinsic conditional autoregressive structure (iCAR) and $N_i$ is the population size of each NUTS-3 Region.

The parameter were estimated by INLA (Integrated Nested Laplace Approximation) and R software.

```{r}
nomi_prov<-levels(dat_csv$denominazione_provincia)
barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum),names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number fo cases")
```
The number of observed cases $y_i$ is very different among  NUTS-3 regions.

```{r}
pop<-tapply(dat_csv$pop,dat_csv$denominazione_provincia,mean)
barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum)/pop*1000,names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number of cases x 1.000 inhabitants")
```
A correct comparison is the made by incidence cases $y_i/N_i$, here riported for x1000 inhabitants. 
#Spatial model 
We estimate a BYM model fixing prior distributions for iid and besag components as follows:
```{r}
prior.iid = c(1,0.01)
prior.besag = c(1,0.001)
initial.iid = 4
initial.besag = 3
```
We create new variables that are required by INLA procedure. 
```{r}
##########################indexing, for 
#### number of days
dat_csv$t<-as.numeric(dat_csv$data)
days<-max(dat_csv$t)
dat_csv$t3<-dat_csv$t2<-dat_csv$t
#### province
dat_csv$ID2<-dat_csv$ID
########################## formula BYM
formula.bym= totale_casi ~ 1+f(ID, model = "bym", graph = file.adj, param = c(prior.iid, prior.besag), initial = c(initial.iid, initial.besag))
fit_1<-inla(formula.bym, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_1)
##incidence rate ratio for each province
IRR_mean <- exp(fit_1$summary.random$ID$mean[1:nprov])
nc.province$IRR_mean<-IRR_mean
spplot(nc.province, c( "IRR_mean"),main="Incidence Rate Ratios")
nomi_prov
```
The image reported the IRR=exp($\mu_i$), the increase of the Incidence of Covid-19 in each NUTS-3 Region respect to the overall mean.

#Temporal model 
We consider the temporal aspect estimating a Bayesian RW2 model to model the temporal trend of COVID-19 in the reported temporal window. The observed number of COVID-19 a the time $t$, $y_{t}$, is modelled as follows  
$$
  Y_t|\lambda_t\sim \mbox{Poisson}(\lambda_t), \qquad t=i,2,\ldots \\
$$

with t=1,...,T, the quantity $\lambda_{t}$, is modelled by a Random Walk of order 2 (RW2) variable  
$$
  log{\lambda_{it}}=\alpha+\gamma_t+\phi_t,  \\
$$

where $\alpha$ is the intercept, $\gamma_t \sim RW2$  are the coefficients related to the random walk process while $\phi_t \sim N(0, \tau_t^{-1})$ are the temporal specific random errors.

```{r}
Date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days-1,1)
plot(Date,tapply(dat_csv$totale_casi,dat_csv$t,sum),xlab="Date",ylab="New cases",type="l")
points(Date,tapply(dat_csv$totale_casi,dat_csv$t,sum))
formula_t = totale_casi ~ 1+f(t,model="rw2", constr = FALSE)+f(t2,model="iid")
fit_2<-inla(formula_t, family="poisson", data=dat_csv,control.compute = list(dic=T))
summary(fit_2)
##Incidence rate ratios, increase od cases respect to time 1
plot(Date,exp(fit_2$summary.random$t$mean),ylab="temporal IRR",xlab="Date")

label<-Date

mean  <- exp(fit_2$summary.random$t$mean[1:days]+fit_2$summary.fixed$mean)*nprov
lower <- exp(fit_2$summary.random$t$mean[1:days]-1.96*fit_2$summary.random$t$sd[1:days]+fit_2$summary.fixed$mean)*nprov
upper <- exp(fit_2$summary.random$t$mean[1:days]+1.96*fit_2$summary.random$t$sd[1:days]+fit_2$summary.fixed$mean)*nprov

df <- data.frame(label, mean, lower, upper)

fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
  geom_pointrange() +
  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
  xlab("Date") + ylab("Estimated new cases") +
  theme_bw() 
print(fp)

```
We now consider a BYM model with spatial and temporal components and their complete interactions (Type IV). The observed number of COVID-19, $y_{it}$, was modelled as:  
$$
  Y_{it} \sim Poisson (\lambda_{it}),  \\
$$

with i=1,...,`r nprov`and t=1,...,`r days`.  
We modelled the ratio $\frac{\lambda_{it}}{N_{i}}$, the Incidence of COVID-19, with a BYM model including temporal covariates as follows
$$
  log(\frac{\lambda_{it}}{N_{i}})=\alpha+\mu_i+\nu_i+\gamma_t+\phi_t+\delta_{it},  \\
$$
  where $\alpha$ is the intercept and:

*	space: $\mu_i$ and $\nu_i$ are two area specific effects with normal distribution modelled using an intrinsic conditional autoregressive structure (iCAR);

*	time:$\gamma_t \sim RW2$  are the coefficients related to the random walk process, while $\phi_t \sim N(0, \tau_t^{-1} )$ are the temporal specific random errors.

*	space-time: $\delta_{it}$ are the coefficients related space-time interactions that in the its type 4 formulation (please see Spatial and Spatio-temporal Bayesian Models with R - INLA, Blangiardo and Cameletti) is made by the Kronecker product of time and space indexes.

```{r}
####specification
formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
  f(t,model="rw2",constr = FALSE) +
  f(t2,model="iid") +
  f(ID2,model="besag", graph=file.adj,
    group=t3,
    control.group=list(model="rw2"))
#estimation
fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
summary(fit_st4)
```

The time and space component mean estimates can be extracted here

```{r}
#overall trend
t_1<-fit_st4$summary.random$t$mean 
#iid trend
t_2<-fit_st4$summary.random$t2$mean 
 # spatial iCAR component
s_1<-fit_st4$summary.random$ID$mean[1:nprov]
# spatial idd
s_2<-fit_st4$summary.random$ID$mean[(nprov+1):(nprov*2)] 
#spatio-temporal model with car and ar1 specification
st<-fit_st4$summary.random$ID2$mean 
```

When can represent the IRR(=exp($\mu_i+\frac{1}{T} \sum_{i=t}^{T} \delta_{it}$)) for each NUTS-3 regions.

```{r}
nc.province$IRR_st<-exp(s_1+apply(t(matrix(st,nrow=nprov)),2,mean))
l1 = list("sp.text", coordinates(nc.province), nc.province$DEN_PROV,col="grey", cex=0.7,font=2)
spplot(nc.province, c( "IRR_st"),main="IRR", sp.layout=list(l1))
```
  Trend for each province is reported below.

```{r}
##overall trend
plot(Date,exp(t_1),ylab="IRR",main="Overall time trend")

## time trend for each province
db_plot<-data.frame(t_1+t(matrix(st,nrow=nprov)))
names(db_plot)<-nc.province$DEN_PROV
names(db_plot)[4]<-"Milano"
province_trend<-xts(db_plot, order.by = Date, frequency = 7)
p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
p

#space time variation for each province from overall trend
db_plot<-data.frame(t(matrix(st,nrow=nprov)))
names(db_plot)<-nc.province$DEN_PROV
names(db_plot)[4]<-"Milano"
province_trend<-xts(db_plot, order.by = Date, frequency = 7)
p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
p
```

#Predictions 3 days forward of new COVID-19 cases
We made a forecast of 3 days forward on the number of COVID19 by means of BYM model and considering as covariate the number of Hubei epidemic lagged of 8 days because the different moment of restrictive measures between Italy and Hubei.
```{r}
Forecast=3
dat_csv2<-dat_csv 
dat_csv2$t<-rep(1:days+days,nprov)
dat_csv2$totale_casi<-NA
dat_csv2<-dat_csv2[dat_csv2$t<=(days+Forecast),]
dat_csv_n<-rbind(dat_csv,dat_csv2)
dat_csv_n$t3<-dat_csv_n$t2<-dat_csv_n$t
#introducting hubei starting on 22/01, quarantine started on 25/01
# applying a smoothing , data has errors
hubei_new_cases<-filter(diff(hubei), rep(1, 5))

plot(hubei/58.5,type="l",ylab="COVID19 cases for milion of inhabitant",ylim=c(0,1530),xlab="Days since over 100 cases")
lines(c(rep(NA,2),cumsum(tapply(dat_csv_n$totale_casi,dat_csv_n$t,sum))/10),col=2)
abline(v=2,col=1,lty=2)
abline(v=4,col=2,lty=2)
abline(v=10,col=3,lty=3)
legend("bottomright",c("Hubei","Lombardy"),lty=1,col=1:2)
text(10,1500,"Hubei quarantine")
text(12,1350,"Lodi home lock",col=2)
text(20,1200,"Lombardy home lock",col=3)
#province of lodi restrictive measure started on 24/02 lag 28 days, at the beginning of the time series
#other provinces on 11/03 lagged started their restrictions
hubei_lodi<-hubei_new_cases[1:(days+Forecast)]
hubei_others<-c(rep(NA,8),hubei_new_cases[1:(days+Forecast-8)])
dat_csv_n<-dat_csv_n[order(dat_csv_n$codice_provincia,dat_csv_n$t),]
dat_csv_n$hubei<-rep(hubei_others,12)
dat_csv_n$hubei[dat_csv_n$denominazione_provincia=="Lodi"]<-hubei_lodi
#overdispersion
dat_csv_n$over<-1:dim(dat_csv_n)[1]
```
We model the number of cases.

In the rw2 I set constr = FALSE is set to FALSE and that, for this reason, the intercept is not included in the linear predictor. 
```{r}
## include overdispersion parameter
formula.intIVn<- totale_casi ~ log(hubei+1)+log(pop)+
  f(t,model="rw2", constr = FALSE) +
  f(t2,model="iid") +f(over,model="iid") +
f(ID,model="bym",graph=file.adj) +
  f(ID2,model="besag", graph=file.adj,
    group=t3,
    control.group=list(model="rw2"))+ f(over,model="iid")

fit_st4n<-inla(formula.intIVn, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = 1))
summary(fit_st4n)
```

```{r}
#### components
#overall trend
t_1<-fit_st4n$summary.random$t$mean 
#iid trend
t_2<-fit_st4n$summary.random$t2$mean 
 # spatial CAR componen
s_1<-fit_st4n$summary.random$ID$mean[1:nprov]
# spatial idd
s_2<-fit_st4n$summary.random$ID$mean[(nprov+1):(nprov*2)] 
# spatio-temporal coefficients 
st<-fit_st4n$summary.random$ID2$mean 
## Trends for province
Date_n<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days+Forecast-1,1)
## time trend for each province
db_plot<-data.frame(t_1+t(matrix(st,nrow=nprov)))
names(db_plot)<-nc.province$DEN_PROV
names(db_plot)[4]<-"Milano"
province_trend<-xts(db_plot, order.by = Date_n, frequency = 7)
p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
p

#space time variation for each province from overall trend
db_plot<-data.frame(t(matrix(st,nrow=nprov)))
names(db_plot)<-nc.province$DEN_PROV
names(db_plot)[4]<-"Milano"
province_trend<-xts(db_plot, order.by = Date_n, frequency = 7)
p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
p
#extract number of cases for lodi

#### FOR LODI as for example
est_lodi<-fit_st4n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Lodi",]

days.before<-Date_n[1:days]
days.ahead<-Date_n[(days+1):(days+Forecast)]
mu.lower<-est_lodi$mean-1.96*est_lodi$sd
mu.lower[mu.lower<0]<-0
mu.upper<-est_lodi$mean+1.96*est_lodi$sd
mu.med<-xts(est_lodi$mean,order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
step.ahead<-Forecast
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Lodi (Credible Interval ",100*0.95,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

### 2020-03-17 BYM_model/INLA_def_html_prova.Rmd
Add home page, navigation and theme

--- 2020-03-17 BYM_model/INLA_def_html.Rmd
@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:html_document:  
+output:
+  html_document:
   df_print: paged
@@
   toc: yes 
-
-always_allow_html: yes
 ---
@@
 $$
-
-and $\lambda_i>0$. We modelled the incidence of COVID-19 by means of a BYM model including the population size as offset variable as:
-
+with $i=1,\ldots,T$ and $\lambda_i>0$.
+We modelled the incidence of COVID-19 by means of a BYM model including the population size as offset variable as:  
 $$
@@
 where $\alpha$ is the intercept, $\mu_i$ and $\nu_i$ are two area specific effects with normal distribution modelled using an intrinsic conditional autoregressive structure (iCAR) and $N_i$ is the population size of each NUTS-3 Region.
-
-The parameter were estimated by INLA (Integrated Nested Laplace Approximation) and R software.
-
-```{r}
-nomi_prov<-levels(dat_csv$denominazione_provincia)
-barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum),names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number fo cases")
-```
-The number of observed cases $y_i$ is very different among  NUTS-3 regions.
-
-```{r}
-pop<-tapply(dat_csv$pop,dat_csv$denominazione_provincia,mean)
-barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum)/pop*1000,names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number of cases x 1.000 inhabitants")
-```
-A correct comparison is the made by incidence cases $y_i/N_i$, here riported for x1000 inhabitants. 
-#Spatial model 
-We estimate a BYM model fixing prior distributions for iid and besag components as follows:
-```{r}
-prior.iid = c(1,0.01)
-prior.besag = c(1,0.001)
-initial.iid = 4
-initial.besag = 3
-```
-We create new variables that are required by INLA procedure. 
-```{r}
-##########################indexing, for 
-#### number of days
-dat_csv$t<-as.numeric(dat_csv$data)
-days<-max(dat_csv$t)
-dat_csv$t3<-dat_csv$t2<-dat_csv$t
-#### province
-dat_csv$ID2<-dat_csv$ID
-########################## formula BYM
-formula.bym= totale_casi ~ 1+f(ID, model = "bym", graph = file.adj, param = c(prior.iid, prior.besag), initial = c(initial.iid, initial.besag))
-fit_1<-inla(formula.bym, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
-summary(fit_1)
-##incidence rate ratio for each province
-IRR_mean <- exp(fit_1$summary.random$ID$mean[1:nprov])
-nc.province$IRR_mean<-IRR_mean
-spplot(nc.province, c( "IRR_mean"),main="Incidence Rate Ratios")
-nomi_prov
-```
-The image reported the IRR=exp($\mu_i$), the increase of the Incidence of Covid-19 in each NUTS-3 Region respect to the overall mean.
-
-#Temporal model 
-We consider the temporal aspect estimating a Bayesian RW2 model to model the temporal trend of COVID-19 in the reported temporal window.
-The observed number of COVID-19 a the time $t$, $y_{t}$, is modelled as follows
-$$
-  Y_t|\lambda_t\sim \mbox{Poisson}(\lambda_t), \qquad t=i,2,\ldots \\
-$$
-
-with t=1,...,T, the quantity $\lambda_{t}$, is modelled by a Random Walk of order 2 (RW2) variable  
-$$
-  log{\lambda_{it}}=\alpha+\gamma_t+\phi_t,  \\
-$$
-
-where $\alpha$ is the intercept, $\gamma_t \sim RW2$  are the coefficients related to the random walk process while $\phi_t \sim N(0, \tau_t^{-1})$ are the temporal specific random errors.
-
-```{r}
-Date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days-1,1)
-plot(Date,tapply(dat_csv$totale_casi,dat_csv$t,sum),xlab="Date",ylab="New cases",type="l")
-points(Date,tapply(dat_csv$totale_casi,dat_csv$t,sum))
-formula_t = totale_casi ~ 1+f(t,model="rw2", constr = FALSE)+f(t2,model="iid")
-fit_2<-inla(formula_t, family="poisson", data=dat_csv,control.compute = list(dic=T))
-summary(fit_2)
-##Incidence rate ratios, increase od cases respect to time 1
-plot(Date,exp(fit_2$summary.random$t$mean),ylab="temporal IRR",xlab="Date")
-
-label<-Date
-
-mean  <- exp(fit_2$summary.random$t$mean[1:days]+fit_2$summary.fixed$mean)*nprov
-lower <- exp(fit_2$summary.random$t$mean[1:days]-1.96*fit_2$summary.random$t$sd[1:days]+fit_2$summary.fixed$mean)*nprov
-upper <- exp(fit_2$summary.random$t$mean[1:days]+1.96*fit_2$summary.random$t$sd[1:days]+fit_2$summary.fixed$mean)*nprov
-
-df <- data.frame(label, mean, lower, upper)
-
-fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
-  geom_pointrange() +
-  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
-  xlab("Date") + ylab("Estimated new cases") +
-  theme_bw() 
-print(fp)
-
-```
-We now consider a BYM model with spatial and temporal components and their complete interactions (Type IV). The observed number of COVID-19, $y_{it}$, was modelled as:  
-$$
-  Y_{it} \sim Poisson (\lambda_{it}),  \\
-$$
-
-with i=1,...,`r nprov`and t=1,...,`r days`.  
-We modelled the ratio $\frac{\lambda_{it}}{N_{i}}$, the Incidence of COVID-19, with a BYM model including temporal covariates as follows
-$$
-  log(\frac{\lambda_{it}}{N_{i}})=\alpha+\mu_i+\nu_i+\gamma_t+\phi_t+\delta_{it},  \\
-$$
-  where $\alpha$ is the intercept and:
-
-*	space: $\mu_i$ and $\nu_i$ are two area specific effects with normal distribution modelled using an intrinsic conditional autoregressive structure (iCAR);
-
-*	time:$\gamma_t \sim RW2$  are the coefficients related to the random walk process, while $\phi_t \sim N(0, \tau_t^{-1} )$ are the temporal specific random errors.
-
-*	space-time: $\delta_{it}$ are the coefficients related space-time interactions that in the its type 4 formulation (please see Spatial and Spatio-temporal Bayesian Models with R - INLA, Blangiardo and Cameletti) is made by the Kronecker product of time and space indexes.
-
-```{r}
-####specification
-formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
-  f(t,model="rw2",constr = FALSE) +
-  f(t2,model="iid") +
-  f(ID2,model="besag", graph=file.adj,
-    group=t3,
-    control.group=list(model="rw2"))
-#estimation
-fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
-summary(fit_st4)
-```
-
-The time and space component mean estimates can be extracted here
-
-```{r}
-#overall trend
-t_1<-fit_st4$summary.random$t$mean 
-#iid trend
-t_2<-fit_st4$summary.random$t2$mean 
- # spatial iCAR component
-s_1<-fit_st4$summary.random$ID$mean[1:nprov]
-# spatial idd
-s_2<-fit_st4$summary.random$ID$mean[(nprov+1):(nprov*2)] 
-#spatio-temporal model with car and ar1 specification
-st<-fit_st4$summary.random$ID2$mean 
-```
-
-When can represent the IRR(=exp($\mu_i+\frac{1}{T} \sum_{i=t}^{T} \delta_{it}$)) for each NUTS-3 regions.
-
-```{r}
-nc.province$IRR_st<-exp(s_1+apply(t(matrix(st,nrow=nprov)),2,mean))
-l1 = list("sp.text", coordinates(nc.province), nc.province$DEN_PROV,col="grey", cex=0.7,font=2)
-spplot(nc.province, c( "IRR_st"),main="IRR", sp.layout=list(l1))
-```
-  Trend for each province is reported below.
-
-```{r}
-##overall trend
-plot(Date,exp(t_1),ylab="IRR",main="Overall time trend")
-
-## time trend for each province
-db_plot<-data.frame(t_1+t(matrix(st,nrow=nprov)))
-names(db_plot)<-nc.province$DEN_PROV
-names(db_plot)[4]<-"Milano"
-province_trend<-xts(db_plot, order.by = Date, frequency = 7)
-p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
-p
-
-#space time variation for each province from overall trend
-db_plot<-data.frame(t(matrix(st,nrow=nprov)))
-names(db_plot)<-nc.province$DEN_PROV
-names(db_plot)[4]<-"Milano"
-province_trend<-xts(db_plot, order.by = Date, frequency = 7)
-p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
-p
-```
-
-#Predictions 3 days forward of new COVID-19 cases
-We made a forecast of 3 days forward on the number of COVID19 by means of BYM model and considering as covariate the number of Hubei epidemic lagged of 8 days because the different moment of restrictive measures between Italy and Hubei.
-```{r}
-Forecast=3
-dat_csv2<-dat_csv 
-dat_csv2$t<-rep(1:days+days,nprov)
-dat_csv2$totale_casi<-NA
-dat_csv2<-dat_csv2[dat_csv2$t<=(days+Forecast),]
-dat_csv_n<-rbind(dat_csv,dat_csv2)
-dat_csv_n$t3<-dat_csv_n$t2<-dat_csv_n$t
-#introducting hubei starting on 22/01, quarantine started on 25/01
-# applying a smoothing , data has errors
-hubei_new_cases<-filter(diff(hubei), rep(1, 5))
-
-plot(hubei/58.5,type="l",ylab="COVID19 cases for milion of inhabitant",ylim=c(0,1530),xlab="Days since over 100 cases")
-lines(c(rep(NA,2),cumsum(tapply(dat_csv_n$totale_casi,dat_csv_n$t,sum))/10),col=2)
-abline(v=2,col=1,lty=2)
-abline(v=4,col=2,lty=2)
-abline(v=10,col=3,lty=3)
-legend("bottomright",c("Hubei","Lombardy"),lty=1,col=1:2)
-text(10,1500,"Hubei quarantine")
-text(12,1350,"Lodi home lock",col=2)
-text(20,1200,"Lombardy home lock",col=3)
-#province of lodi restrictive measure started on 24/02 lag 28 days, at the beginning of the time series
-#other provinces on 11/03 lagged started their restrictions
-hubei_lodi<-hubei_new_cases[1:(days+Forecast)]
-hubei_others<-c(rep(NA,8),hubei_new_cases[1:(days+Forecast-8)])
-dat_csv_n<-dat_csv_n[order(dat_csv_n$codice_provincia,dat_csv_n$t),]
-dat_csv_n$hubei<-rep(hubei_others,12)
-dat_csv_n$hubei[dat_csv_n$denominazione_provincia=="Lodi"]<-hubei_lodi
-#overdispersion
-dat_csv_n$over<-1:dim(dat_csv_n)[1]
-```
-We model the number of cases.
-
-In the rw2 I set constr = FALSE is set to FALSE and that, for this reason, the intercept is not included in the linear predictor. 
-```{r}
-## include overdispersion parameter
-formula.intIVn<- totale_casi ~ log(hubei+1)+log(pop)+
-  f(t,model="rw2", constr = FALSE) +
-  f(t2,model="iid") +f(over,model="iid") +
-f(ID,model="bym",graph=file.adj) +
-  f(ID2,model="besag", graph=file.adj,
-    group=t3,
-    control.group=list(model="rw2"))+ f(over,model="iid")
-
-fit_st4n<-inla(formula.intIVn, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = 1))
-summary(fit_st4n)
-```
-
-```{r}
-#### components
-#overall trend
-t_1<-fit_st4n$summary.random$t$mean 
-#iid trend
-t_2<-fit_st4n$summary.random$t2$mean 
- # spatial CAR componen
-s_1<-fit_st4n$summary.random$ID$mean[1:nprov]
-# spatial idd
-s_2<-fit_st4n$summary.random$ID$mean[(nprov+1):(nprov*2)] 
-# spatio-temporal coefficients 
-st<-fit_st4n$summary.random$ID2$mean 
-## Trends for province
-Date_n<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days+Forecast-1,1)
-## time trend for each province
-db_plot<-data.frame(t_1+t(matrix(st,nrow=nprov)))
-names(db_plot)<-nc.province$DEN_PROV
-names(db_plot)[4]<-"Milano"
-province_trend<-xts(db_plot, order.by = Date_n, frequency = 7)
-p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
-p
-
-#space time variation for each province from overall trend
-db_plot<-data.frame(t(matrix(st,nrow=nprov)))
-names(db_plot)<-nc.province$DEN_PROV
-names(db_plot)[4]<-"Milano"
-province_trend<-xts(db_plot, order.by = Date_n, frequency = 7)
-p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
-p
-#extract number of cases for lodi
-
-#### FOR LODI as for example
-est_lodi<-fit_st4n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Lodi",]
-
-days.before<-Date_n[1:days]
-days.ahead<-Date_n[(days+1):(days+Forecast)]
-mu.lower<-est_lodi$mean-1.96*est_lodi$sd
-mu.lower[mu.lower<0]<-0
-mu.upper<-est_lodi$mean+1.96*est_lodi$sd
-mu.med<-xts(est_lodi$mean,order.by = c(days.before,days.ahead),frequency = 7)
-counts<-mu.med
-step.ahead<-Forecast
-mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Lodi (Credible Interval ",100*0.95,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
-p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
-p
-```

### 2020-03-17 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Veneto Region: Scenario 3  (Credible Interval ",100*50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 3  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")

### 2020-03-18 draft_analysis_Italy_html.Rmd
Add home page, navigation and theme

--- 2020-03-15 draft_analysis_Italy.Rmd
@@
 ```
-
 ---
@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:
-  html_document:
+output:html_document:
   df_print: paged
@@
 ---
-
 <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.
@@
 checkpackage("EpiDynamics")
+checkpackage("webshot")
+checkpackage("bsts")
+checkpackage("ggplot2")
 ```
@@
 library(EpiDynamics)
+library(webshot)
+library(bsts)
+library(EpiDynamics)
+library(ggplot2)
 ```
@@
 $$
-\log(I(t))= \alpha + \beta  t +e_t
+\log(I(t))= \beta_0 + \beta_1  t +e_t
 $$
 
-The estimated slope coefficient $\hat\beta$ is used to estimate $R_0$ as in the following formula:
+The estimated slope coefficient $\hat\beta_1$ is used to estimate $R_0$ as in the following formula:
 
-$$\widehat\beta=(\widehat{R_0})\,(\gamma+\mu)$$
+$$
+\widehat\beta_1=(\widehat{R_0}-1)\,(\gamma+\mu)
+$$
 The parameter $\mu$<<$\gamma$ and it can not be considered. As consequence, R0 can be estimated as follows
-$$\hat{R_0}=\frac{\hat{\beta}}{\gamma}
 $$
+\hat{R_0}=1+\frac{\hat{\beta_1}}{\gamma}
+$$
+Respect to the SIR model, $R_0$ can be estimated as follows:
+$$
+\hat{R_0}=\frac{\hat{\beta}}{\gamma}
+$$
+And this was we can retrive the value of $\beta$ in the SEIR model by means of 
+$$
+\hat{R_0}=\frac{\hat{\beta}}{\gamma}=1+\frac{\hat{\beta_1}}{\gamma}\\
+\hat{\beta}=(1+\frac{\hat{\beta_1}}{\gamma})*{\gamma}={\gamma}+\hat{\beta}_1
+$$
+where $\beta_1$ is the slope coefficient. \\
 
@@
 
-mean  <- (beta_vec*14)
-lower <- ((beta_vec-1.96*sd_vec)*14)
-upper <- ((beta_vec+1.96*sd_vec)*14)
+mean  <- 1+(beta_vec*14)
+lower <- 1+((beta_vec-1.96*sd_vec)*14)
+upper <- 1+((beta_vec+1.96*sd_vec)*14)
 
@@
 The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
-We predict beta (and R0) for the next 14 days assuming a Gamma distribution for the beta (the slope) forcing its value to be greater than 0.
-The trend was not monotonic, we use a simple splines to increase the fitting of the model to the data.
+We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Normal distribution for the beta (the slope).
 ```{r R0 forecast}
-library(splines)
+
 time<-3:(days-2)
-beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=Gamma)
+beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian)
 forecast=14
@@
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-beta.predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
-
+predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
+beta.predict<-predict 
 r0.predict<-beta.predict
-r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14
+r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14+1
 # plot the points (actual observations), regression line, and confidence interval
@@
 - S0=N, the size of Italian population   
-- E= f * I0 (with f a fixed factor of the previous scenario)  
-- I0: initial number of COVID-19 cases  
-- R0: initial number of recovered
+- E= f * I_start (with f a fixed factor of the previous scenario)  
+- I_start: initial number of COVID-19 cases  
+- R_start: initial number of recovered
 
 and parameters:  
-- beta: the quantity connected to R0 is considered to vary according the previous estimation  
+- beta: gamma+slope $\beta_1$
 - gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
@@
 # initial number of infectus
-I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I0
+I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
 # initial number of recovered
-R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
+R_start<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R_start
 # italian poulation
@@
 <img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>  
+Respect to the previous formulation, the compartment of exposed people (E) was inserted between the susceptible and infected compartments.  
+(https://en.wikipedia.org/wiki/Bayesian_structural_time_series)
 
-where the parameter beta here is time dependent, as estimated before by the gamma regression model.
-```{r  scenario plot, echo=TRUE}
-library(EpiDynamics)
+We perfom simulation on the trend of beta coefficient by means of a Bayesian Structural Time Series using the library bsts of R.    
+We estimate a BSTS model specifing a local linear trend ans 1000 simulation.
 
+We made a forecast forward of 14 days dropping the first 100 simulation as burn-in.
+
+```{r bsts model fit and prediction, echo=TRUE}
+# Bayesian Structural Time Series
+set.seed(123)
+ss <- AddLocalLinearTrend(list(), beta_vec)
+model1 <- bsts(beta_vec,
+               state.specification = ss,
+               niter = 1000)
+par(mfrow = c(1,1))
+plot(model1, "components", ylab="Beta coefficient", xlab="days")
+#previsioni
+pred1 <- predict(model1, horizon = 16, burn = 100)
+par(mfrow = c(1,1))
+plot(pred1 , ylab="Beta coefficient",main="Data and predictions")
+# matrix of beta coefficients
+coef<-(pred1$distribution[,3:16])
+par(mfrow = c(1,1))
+```
+For each vectors of simulated beta coefficients we perform a SEIR model (one for each scenario).  
+We save the results and plot a credible interval at 90% (5-95% percentile).
+
+```{r  scenario plot , echo=TRUE}
+
+seir1_sim<-seir2_sim<-seir3_sim<-NULL
+for(s in 1:dim(coef)[1]){
   # average number of single connections of an infected person
@@
   for(i in 1:forecast){
-parameters <- c(mu = mu0, beta = beta.predict$fit[days+i], sigma = sigma0, gamma = 1/duration)
+    parameters <- c(mu = mu0, beta = matrix(coef[s,i])+1/duration, sigma = sigma0, gamma = 1/duration)
     f1<-10
-if( i==1) initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
+    if( i==1) initials <- c(S = 0.95, E = (f1*I_start/N), I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
@@
     f2<-5
-if( i==1) initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
+    if( i==1) initials <- c(S = 0.95, E = (f2*I_start/N), I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
@@
     f3<-3
-if( i==1) initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
+    if( i==1) initials <- c(S = 0.95, E = (f3*I_start/N), I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])

### 2020-03-18 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

@@
 The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
-We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Log-normal distribution for the beta (the slope) and forcing its value to be greater than 0.
+We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Normal distribution for the beta (the slope).
 ```{r R0 forecast}
@@
 time<-3:(days-2)
-beta.model<-glm(log(beta_vec)~time,weights = 1/sd_vec,family=gaussian)
+beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian)
 forecast=14
@@
 beta.predict<-predict 
-beta.predict[,3:5]<-exp(predict[,3:5])
 r0.predict<-beta.predict
-r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14
+r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14+1
 # plot the points (actual observations), regression line, and confidence interval
@@
 We made a forecast by means of a SEIR model fixing a series of initial status:  
-- S0=N, the size of Italian population   
-- E= f * I0 (with f a fixed factor of the previous scenario)  
-- I0: initial number of COVID-19 cases  
-- R0: initial number of recovered
+- S_start=N, the size of Italian population   
+- E= f * I_start (with f a fixed factor of the previous scenario)  
+- I_start: initial number of COVID-19 cases  
+- R_start: initial number of recovered
 
 and parameters:  
-- beta: the quantity connected to R0 is considered to vary according the previous estimation  
+- beta: gamma+slope $\beta_1$
 - gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
@@
 set.seed(123)
-ss <- AddLocalLinearTrend(list(), log(beta_vec))
-model1 <- bsts(log(beta_vec),
+ss <- AddLocalLinearTrend(list(), beta_vec)
+model1 <- bsts(beta_vec,
                state.specification = ss,
@@
 par(mfrow = c(1,1))
-plot(model1, "components", ylab="Log beta coefficient", xlab="days")
+plot(model1, "components", ylab="Beta coefficient", xlab="days")
 #previsioni
@@
 par(mfrow = c(1,1))
-plot(pred1 , ylab="Log beta coefficient",main="Data and predictions")
+plot(pred1 , ylab="Beta coefficient",main="Data and predictions")
 # matrix of beta coefficients
-coef<-(exp(pred1$distribution)[,3:16])
+coef<-(pred1$distribution[,3:16])
 par(mfrow = c(1,1))
@@
   for(i in 1:forecast){
-    parameters <- c(mu = mu0, beta = matrix(coef[s,i]), sigma = sigma0, gamma = 1/duration)
+    parameters <- c(mu = mu0, beta = matrix(coef[s,i])+1/duration, sigma = sigma0, gamma = 1/duration)
     f1<-10
-    if( i==1) initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
+    if( i==1) initials <- c(S = 0.95, E = (f1*I_start/N), I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
@@
     f2<-5
-    if( i==1) initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
+    if( i==1) initials <- c(S = 0.95, E = (f2*I_start/N), I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
@@
     f3<-3
-    if( i==1) initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
+    if( i==1) initials <- c(S = 0.95, E = (f3*I_start/N), I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])

### 2020-03-18 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

@@
 The number of infected individuals $I$ at time $t$ is approximately 
-$$I(t)\;\approx\;I_0\,e^{R_0\,(\gamma+\mu)\,t}$$
+$$I(t)\;\approx\;I_0\,e^{(R_0-1)\,(\gamma+\mu)\,t}$$
 where $I_0$ is the (small) number of infectives at time $0$, $\frac{1}{\gamma}$ is the infectious period, and $\frac{1}{\mu}$ is the host lifespan.
@@
 
-$$\log{I}(t)\;\approx\;\log{I_0}+(R_0)\,(\gamma+\mu)\,t,$$
+$$\log{I}(t)\;\approx\;\log{I_0}+(R_0-1)\,(\gamma+\mu)\,t,$$
 which implies that a semi-log plot of $I$ vs $t$ should be approximately linear with a slope proportional to $R_0$ and the recovery rate.
@@
 $$
-\log(I(t))= \alpha + \beta  t +e_t
+\log(I(t))= \beta_0 + \beta_1  t +e_t.
 $$
 
-The estimated slope coefficient $\hat\beta$ is used to estimate $R_0$ as in the following formula:
+The estimated slope coefficient $\hat\beta_1$ is used to estimate $R_0$ as in the following formula:
 
-$$\widehat\beta=(\widehat{R_0})\,(\gamma+\mu)$$
+$$
+\widehat\beta_1=(\widehat{R_0}-1)\,(\gamma+\mu)
+$$
 The parameter $\mu$<<$\gamma$ and it can not be considered. As consequence, R0 can be estimated as follows
-$$\hat{R_0}=\frac{\hat{\beta}}{\gamma}
 $$
+\hat{R_0}=1+\frac{\hat{\beta_1}}{\gamma}
+$$
+Respect to the SIR model, $R_0$ can be estimated as follows:
+$$
+\hat{R_0}=\frac{\hat{\beta}}{\gamma}
+$$
+And this was we can retrive the value of $\beta$ in the SEIR model by means of 
+$$
+\hat{R_0}=\frac{\hat{\beta}}{\gamma}=1+\frac{\hat{\beta_1}}{\gamma}\\
+\hat{\beta}=(1+\frac{\hat{\beta_1}}{\gamma})*{\gamma}={\gamma}+\hat{\beta}_1
+$$
+where $\beta_1$ is the slope coefficient. \\
 
@@
 - S0=N, the size of Veneto Region population   
-- E= f * I0 (with f a fixed factor of the previous scenario)  
-- I0: initial number of COVID-19 cases  
-- R0: initial number of recovered
+- E= f * I_start (with f a fixed factor of the previous scenario)  
+- I_start: initial number of COVID-19 cases  
+- R_start: initial number of recovered
 
 and parameters:  
-- beta: the quantity connected to R0 is considered to vary according the previous estimation  
-- gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
+- beta: gamma + slope
+- gamma= 1/duration (rate of infection duration of COVID-19, duration=14 days)  
 - sigma0: the coronavirus transmission rate (half of flu epidemic)   
@@
 # initial number of infectus
-I0<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I0
+I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
 # initial number of recovered
-R0<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R0
+R_start<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R_start
 # Veneto Region population
@@
 # Bayesian Structural Time Series
-ss <- AddLocalLinearTrend(list(), log(beta_vec))
-model1 <- bsts(log(beta_vec),
+ss <- AddLocalLinearTrend(list(), beta_vec)
+model1 <- bsts(beta_vec,
                state.specification = ss,
@@
 par(mfrow = c(1,1))
-plot(model1, "components", ylab="Log beta coefficient", xlab="days")
+plot(model1, "components", ylab="Beta1 coefficient", xlab="days")
 #previsioni
@@
 par(mfrow = c(1,1))
-plot(pred1 , ylab="Log beta coefficient",main="Data and predictions")
+plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions")
 # matrix of beta coefficients
-coef<-(exp(pred1$distribution)[,3:16])
+coef<-(pred1$distribution[,3:16])
 par(mfrow = c(1,1))
@@
   for(i in 1:forecast){
-    parameters <- c(mu = mu0, beta = matrix(coef[s,i]), sigma = sigma0, gamma = 1/duration)
+    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma0, gamma = 1/duration)
     f1<-10
-    if( i==1) initials <- c(S = 0.95, E = (f1*I0/N), I = I0/N, R = R0/N)
+    if( i==1) initials <- c(S = 0.95, E = (f1*I_start/N), I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
@@
     f2<-5
-    if( i==1) initials <- c(S = 0.95, E = (f2*I0/N), I = I0/N, R = R0/N)
+    if( i==1) initials <- c(S = 0.95, E = (f2*I_start/N), I =  I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
@@
     f3<-3
-    if( i==1) initials <- c(S = 0.95, E = (f3*I0/N), I = I0/N, R = R0/N)
+    if( i==1) initials <- c(S = 0.95, E = (f3*I_start/N), I =  I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])

### 2020-03-18 BYM_model/INLA_Veneto.R

@@ -1,0 +1,202 @@
+#Load libraries
+library(spdep)
+library(maptools)
+library(INLA)
+library(rgdal)
+library(foreign)
+library("dygraphs")
+library("xts")
+#define work directory
+################################
+rm(list=ls())
+wd<-"~/Dropbox/COVID19"
+setwd(wd)
+################################ IMPORT DATASETS
+# ISTAT Province dataset
+db_istat<-read.csv("Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
+pop_provincia<-data.frame(codice_provincia=as.numeric(names(table(db_istat[,3]))),pop=tapply(db_istat[,20],db_istat[,3],sum))
+###import international dataset
+dat_int<-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv",header=T)
+hubei<-as.numeric(dat_int[dat_int$Province.State=="Hubei",5:dim(dat_int)[2]])
+
+###import updated italian dataset  for each province
+dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv",header=T)
+dat_csv<-merge(dat_csv,pop_provincia,by=c("codice_provincia"),all.y=TRUE)
+dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
+#select a Region, in this example "Lombardia" Code region 3
+Region<-5
+dat_csv<-dat_csv[dat_csv$codice_regione %in% Region,]
+dat_csv_n$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
+#### number of province
+nprov<-length(table(dat_csv$denominazione_provincia)[table(dat_csv$denominazione_provincia)>0])
+
+#order the obsercation for date and space, useful for INLA later....
+dat_csv<-dat_csv[order(dat_csv$data),]
+dat_csv<-dat_csv[order(dat_csv$codice_provincia),]
+###space ID
+dat_csv$ID<-as.factor(dat_csv$codice_provincia)
+levels(dat_csv$ID)<-1:length(levels(dat_csv$ID))
+dat_csv$ID<-as.numeric(dat_csv$ID)
+#calculate new_cases
[…]

### 2020-03-18 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

@@
 The 3 scenarios show different numbers. 
-If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N)[forecast]`).
-
-In the next plot the cumulative number of infected.  
-At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N+tapply(seir2_sim$R,seir2_sim$time,median)*N)[forecast]`).
+If we consider the second scenario at the end of the 2 weeks (`r date[length(date)]`):
+* the number of infected is  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N)[forecast]`).
+* the total number of COVID19 cases is expected to be  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N+tapply(seir2_sim$R,seir2_sim$time,median)*N)[forecast]`).

### 2020-03-18 draft_analysis_Veneto_html.Rmd
Add home page, navigation and theme

--- 2020-03-18 draft_analysis_Veneto_new.Rmd
@@
----
-title: "COVID19 - Forecast and predictions using a time dependent SEIR model - Veneto Region"
-author: "Paolo Girardi"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: rmarkdown::github_document
-always_allow_html: yes
----
 ```{r echo=FALSE}
@@
 ```
+---
+title: "COVID19 - Forecast and predictions using a time dependent SEIR model for the Veneto Region"
+author: "Paolo Girardi"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output:html_document:
+  df_print: paged
+pdf_document: 
+  toc: yes
+---
 <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.

### 2020-03-19 BYM_model/INLA_Veneto_html.Rmd
Add home page, navigation and theme

--- 2020-03-17 BYM_model/INLA_def_html.Rmd
@@
 dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
-#select a Region, in this example "Lombardia" Code region 3
-Region<-3
+#select a Region, in this example "Veneto" Code region 3
+Region<-5
 dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
@@
 nomi_prov<-levels(dat_csv$denominazione_provincia)
-barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum),names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number fo cases")
+barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum),names =1:nprov,args.legend = list(x = "topright"),legend.text =paste(1:nprov,nomi_prov),ylab="Number fo cases")
 ```
@@
 pop<-tapply(dat_csv$pop,dat_csv$denominazione_provincia,mean)
-barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum)/pop*1000,names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number of cases x 1.000 inhabitants")
+barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum)/pop*1000,names =1:nprov,args.legend = list(x = "topright"),legend.text =paste(1:nprov,nomi_prov),ylab="Number of cases x 1.000 inhabitants")
 ```
@@
 names(db_plot)<-nc.province$DEN_PROV
-names(db_plot)[4]<-"Milano"
+names(db_plot)[4]<-"Venezia"
 province_trend<-xts(db_plot, order.by = Date, frequency = 7)
-p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
+p <- dygraph(province_trend,main=paste("Veneto Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
 p
@@
 names(db_plot)<-nc.province$DEN_PROV
-names(db_plot)[4]<-"Milano"
+names(db_plot)[4]<-"Venezia"
 province_trend<-xts(db_plot, order.by = Date, frequency = 7)
-p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
+p <- dygraph(province_trend,main=paste("Veneto Region Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
 p
@@
 abline(v=10,col=3,lty=3)
-legend("bottomright",c("Hubei","Lombardy"),lty=1,col=1:2)
+legend("bottomright",c("Hubei","Veneto Region"),lty=1,col=1:2)
 text(10,1500,"Hubei quarantine")
 text(12,1350,"Lodi home lock",col=2)
-text(20,1200,"Lombardy home lock",col=3)
+text(20,1200,"Veneto home lock",col=3)
 #province of lodi restrictive measure started on 24/02 lag 28 days, at the beginning of the time series
@@
 dat_csv_n<-dat_csv_n[order(dat_csv_n$codice_provincia,dat_csv_n$t),]
-dat_csv_n$hubei<-rep(hubei_others,12)
-dat_csv_n$hubei[dat_csv_n$denominazione_provincia=="Lodi"]<-hubei_lodi
+dat_csv_n$hubei<-rep(hubei_others,nprov)
 #overdispersion
@@
 province_trend<-xts(db_plot, order.by = Date_n, frequency = 7)
-p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
+p <- dygraph(province_trend,main=paste("Veneto Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
 p
@@
 names(db_plot)<-nc.province$DEN_PROV
-names(db_plot)[4]<-"Milano"
+names(db_plot)[4]<-"Venezia"
 province_trend<-xts(db_plot, order.by = Date_n, frequency = 7)
-p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
+p <- dygraph(province_trend,main=paste("Veneto Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
 p
-#extract number of cases for lodi
+#extract number of cases for Verona
 
-#### FOR LODI as for example
-est_lodi<-fit_st4n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Lodi",]
+#### FOR Verona as for example
+est_verona<-fit_st4n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Verona",]
 
@@
 days.ahead<-Date_n[(days+1):(days+Forecast)]
-mu.lower<-est_lodi$mean-1.96*est_lodi$sd
+mu.lower<-est_verona$mean-1.96*est_verona$sd
 mu.lower[mu.lower<0]<-0
-mu.upper<-est_lodi$mean+1.96*est_lodi$sd
-mu.med<-xts(est_lodi$mean,order.by = c(days.before,days.ahead),frequency = 7)
+mu.upper<-est_verona$mean+1.96*est_verona$sd
+mu.med<-xts(est_verona$mean,order.by = c(days.before,days.ahead),frequency = 7)
 counts<-mu.med
@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Lodi (Credible Interval ",100*0.95,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Verona (Credible Interval ",100*0.95,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")

### 2020-03-19 draft_analysis_Veneto_html.Rmd
Add home page, navigation and theme

@@
     f1<-10
-    if( i==1) initials <- c(S = 0.95, E = (f1*I_start/N), I = I_start/N, R = R_start/N)
+    if( i==1) initials <- c(S = 1-((1+f1)*I_start/N)-R_start/N, E = (f1*I_start/N), I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
@@
     f2<-5
-    if( i==1) initials <- c(S = 0.95, E = (f2*I_start/N), I =  I_start/N, R = R_start/N)
+    if( i==1) initials <- c(S = 1-((1+f1)*I_start/N)-R_start/N, E = (f2*I_start/N), I =  I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
@@
     f3<-3
-    if( i==1) initials <- c(S = 0.95, E = (f3*I_start/N), I =  I_start/N, R = R_start/N)
+    if( i==1) initials <- c(S = 1-((1+f1)*I_start/N)-R_start/N, E = (f3*I_start/N), I =  I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])

### 2020-03-19 BYM_model/INLA_Veneto.R

@@ -23,7 +23,7 @@ hubei<-as.numeric(dat_int[dat_int$Province.State=="Hubei",5:dim(dat_int)[2]])
 dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-province/dpc-covid19-ita-province.csv",header=T)
 dat_csv<-merge(dat_csv,pop_provincia,by=c("codice_provincia"),all.y=TRUE)
 dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
-#select a Region, in this example "Lombardia" Code region 3
+#select a Region, in this example "Lombardia" Code region 3, Veneto 5
 Region<-5
 dat_csv<-dat_csv[dat_csv$codice_regione %in% Region,]
 dat_csv_n$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
@@ -104,6 +104,8 @@ formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
 #estimation
 fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
 summary(fit_st4)
+fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv,control.compute = list(dic=T))
+summary(fit_st4)
 
 #### components
 t_1<-fit_st4$summary.random$t$mean #overall trend
@@ -117,6 +119,14 @@ spplot(nc.province, c( "IRR_st"),main="IRR")
 ## Trends for province
 
 trends<-t_1+t(matrix(st,nrow=nprov))
+R0_vec<-NULL
+for ( i in 1:nprov){
+  R0_vec<-c(R0_vec,coef(lm(trends[10:days,i]~I(10:days)))[2]*14+1)
+}
+
+nc.province$R0<-R0_vec
+spplot(nc.province, c( "R0"),main="R0, since 5/03")
+
 # trends+ specific
 matplot(trends)
 #specific for each province
@@ -180,6 +190,7 @@ trends<-t_1+t(matrix(st,nrow=nprov))
 plot(Date_n,t_1,ylab="RW2 Time coefficient",xlab="Date")
 plot(Date_n,trends,ylab="RW2 Time coefficient",xlab="Date")
 matplot(trends,ylab="RW2 Time coefficient for Province",xlab="Date")
+
 
 matplot(t(matrix(st,nrow=nprov)),ylab="RW2 Time coefficient for Province",xlab="Date")
 #extract number of Verona

### 2020-03-21 BYM_model/INLA_Veneto_html.Rmd
Add home page, navigation and theme

@@
 formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
-  f(t,model="rw2",constr = FALSE) +
+  f(t,model="rw2") +
   f(t2,model="iid") +

### 2020-03-21 draft_analysis_Italy_html.Rmd
Add home page, navigation and theme

@@
 ```{r}
-
 rm(list=ls())
@@
 basis of the actual data, we used a SEIR model applied to the COVID19
-epidemic in Italy.
+epidemic to Italy
 
@@
 Here, $B$ is the crude birth rate (births per unit time), $\mu$ is the death rate and $\gamma$ is the recovery rate.
-We'll assume that the force of infection, $\lambda$, for a constant population $N$
+We'll assume that the force of infection, $\beta$, for a constant population $N$
 $$\lambda = \beta\,\frac{I}{N},$$
@@
 $$
-\log(I(t))= \beta_0 + \beta_1  t +e_t
+\log(I(t))= \beta_0 + \beta_1  t +e_t.
 $$
@@
 
-The incubation period $1/ \gamma$ for the coronavirus is in mean 5.1 days with a range from 2-14 days.
+The incubation period  for the coronavirus is in mean 5.1 days with a range from 2-14 days.
 Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
 However, the incubation period is used for epidemic diseases that causes the immediate home isolation of infected subjects.
+The duration of the diseases is about 2 weeks.
 
-In the calculation we considered an "incubation period" of 14 days for two reasons:  
-1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 
-2) 14 days is the worst scenario because in this period low impact COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).
+However, in the calculation of R0 we considered an infectious period of 21 days for two reasons:  
+1) the majority of cases is asymptomatic;
+2) 21 days is the worst scenario because in this period COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).
 
@@
 #vector for beta and standard deviation
+duration<-21
 beta_vec<-NULL
@@
 
-mean  <- 1+(beta_vec*14)
-lower <- 1+((beta_vec-1.96*sd_vec)*14)
-upper <- 1+((beta_vec+1.96*sd_vec)*14)
+mean  <- 1+(beta_vec*duration)
+lower <- 1+((beta_vec-1.96*sd_vec)*duration)
+upper <- 1+((beta_vec+1.96*sd_vec)*duration)
 
@@
 r0.predict<-beta.predict
-r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*14+1
+r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*21+1
 # plot the points (actual observations), regression line, and confidence interval
@@
 R0 passes from a value of `r round(r0.predict[1,1],2)`   in the initial phase to an  estimated value of `r round(r0.predict[days+forecast,3],2)` at the ending of the 14-days forecast.  
+We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:
 
-We want to make a short term forecast (14 days) with 3 scenario, based on the number of exposed people:
+$$\begin{aligned}
+\frac{d}{dt}S(t) &= \mu (N-S)-\beta \frac{SI}{N}-\nu S\\  
+\frac{d}{dt}E(t) &= \beta \frac{SI}{N}-(\mu+\sigma) E    \\
+\frac{d}{dt}I(t) &=\sigma E- (\mu+\gamma)I \\
+\frac{d}{dt}R(t) &= \gamma\,I(t)-\mu\,R(t)+\nu S.\\
+\end{aligned}$$
 
--Scenario 1: 10 exposed people for each COVID-19 case (no home restrictions made or even no effects)
+<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>
 
--Scenario 2: 5 exposed people for each COVID-19 case (-50% exposed people)
+Respect to the previous formulation, the compartment of exposed people (E) was inserted between the susceptible and infected compartments.  
+(https://en.wikipedia.org/wiki/Bayesian_structural_time_series).
+The parameter $\nu$ is the rate of vaccination.
 
--Scenario 3: 3 exposed people for each COVID-19 case  (-70% exposed people)
+We want to make a short term forecast (14 days).
 
 We made a forecast by means of a SEIR model fixing a series of initial status:
-- S0=N, the size of Italian population   
-- E= f * I_start (with f a fixed factor of the previous scenario)  
-- I_start: initial number of COVID-19 cases  
-- R_start: initial number of recovered
+*S:N, the size of Italian population  
+*E: The number of exposed people (unknown)  
+*I_start: initial number of COVID-19 cases     
+*R_start: initial number of recovered
 
 and parameters:  
-- beta: gamma+slope $\beta_1$
-- gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
-- sigma0: the coronavirus transmission rate (half of flu epidemic)   
+- beta: gamma + slope
+- gamma= 1/duration of diseases (duration=21 days)
+- sigma: the coronavirus transmission rate 
 - mu0: the overall mortality rate
 
-```{r initation plot, echo=TRUE}
+```{r ini parameters plot, echo=TRUE}
 # initial number of infectus
 I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
-# initial number of recovered
-R_start<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R_start
-# italian poulation
+# initial number of recovered, based on the proportion of discharged from the health services
+prop<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]/dat_csv$totale_ospedalizzati[dim(dat_csv)[1]]
+R_start<-prop*dat_csv$totale_casi[dim(dat_csv)[1]]; R_start
+# Italian population
 N=60480000
-# duration of COVID19 
-duration<-14
-#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
-sigma0<-0.05
+# duration of COVID19 as the sum of incubation (5-14 days) and the duration of diseases (about 20 days)
+duration<-21
 #mortality rate 
@@
 ```
-We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:
 
-<img src="http://www.public.asu.edu/~hnesse/classes/seireqn.png"/>  
-<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>  
-Respect to the previous formulation, the compartment of exposed people (E) was inserted between the susceptible and infected compartments.  
-(https://en.wikipedia.org/wiki/Bayesian_structural_time_series)
+We try to estimates sigma (from exposed to infected) and the number of exposed population on the basis of the last 5 days of observed number of total infected people.
 
-We perfom simulation on the trend of beta coefficient by means of a Bayesian Structural Time Series using the library bsts of R.    
+We suppose that the number of total infected $I(t)$ can be the realization of a Poisson random variable
+$$
+I(t) \sim Poisson (\theta_t)
+$$
+where $\theta_t$ is the unknown mean of the random process at the time $t$.
+
+The Poisson distribution has a fixed variance that is equal to the mean; in the next part we try to estimates the parameters considering a Negative Binomial distribution which allows for overdispersion.
+
+However, the parameter $\theta_t$ can be supposed to be a realization of a SEIR model with initial parameters $\theta=(\mu,\beta,\sigma, \gamma)$  and initial status $S=(S_{start},E_{start},I_{start},R_{start})$.
+
+The parameter $\sigma$ and the fraction of exposed people ($E_{start}$) are our parameters of interest; we estimate $\sigma$ and $E_{start}$ minimizing the log likelihood of the described Poisson distribution (or Negative Binomial one).
+
+```{r estimation of gamma and exposed, echo=TRUE}
+# number of the last considered days for calibration
+last<-5
+#I_start
+I_1<-dat_csv$totale_attualmente_positivi[(days-last)]
+#R_start
+R_1<-prop*dat_csv$totale_casi[(days-last)]
+#loglikelihood function
+LogLikfun <- function (initials,parameters,obs) {
+n<-length(obs)
+N=60480000
+seir_fit <- SEIR(pars = parameters, init = initials, time = 0:last)
+#Poisson
+#sum(dpois(x=obs[-1],lambda = seir_fit$results$I[-1]*N,log=TRUE))
+#Neg Binomial
+sum(dnbinom(x=obs[-1],mu=seir_fit$results$I[-1]*N,size=N,log=TRUE))
+#SSE
+#-sum((obs[-1]-seir_fit$results$I[-1]*N)^2)
+}
+### logit and its inverse for sigma 
+logit <- function (p) log(p/(1-p))    # the logit transform
+expit <- function (x) 1/(1+exp(-x))   # inverse logit
+
+f1<-function(par){
+f_exp<-expit(par[1])
+sigma<-expit(par[2])
+parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma =1/duration)
+initials <- c(S = (1-(f_exp+I_1/N-R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
+-LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+}
+est<-optim(fn=f1,par=c(logit(0.0005),logit(0.1)))
+#estimated par
+expit(est$par)
+f_exp<-expit(est$par[1])
+sigma<-expit(est$par[2])
+parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma = 1/duration)
+initials <- c(S = (1-(f_exp+I_1/N-R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
+pro<-SEIR(pars = parameters, init = initials, time = 0:last)
+#predicted vs observed
+cbind(pro$results$I*N,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+```
+
+The values of initial parameters are
+```{r ini parameters, echo=TRUE}
+#mu0
+mu0;
+#sigma
+sigma
+#gamma
+1/duration
+```
+
+```{r status, echo=TRUE}
+#mthe unknowrn fraction of exposed people is
+f_exp;
+#that is 
+f_exp/(I_1/N)
+#times the infected
+```
+For the beta parameter, we perfom a simulation on its trend by means of a Bayesian Structural Time Series using the library bsts of R.    
 We estimate a BSTS model specifing a local linear trend ans 1000 simulation.
@@
 # Bayesian Structural Time Series
-set.seed(123)
 ss <- AddLocalLinearTrend(list(), beta_vec)
@@
 par(mfrow = c(1,1))
-plot(model1, "components", ylab="Beta coefficient", xlab="days")
+plot(model1, "components", ylab="Beta1 coefficient", xlab="days")
 #previsioni
@@
 par(mfrow = c(1,1))
-plot(pred1 , ylab="Beta coefficient",main="Data and predictions")
+plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions")
 # matrix of beta coefficients
@@
 ```
-For each vectors of simulated beta coefficients we perform a SEIR model (one for each scenario).  
-We save the results and plot a credible interval at 90% (5-95% percentile).
+For each vector of simulated beta coefficients we perform a SEIR model.
+We save the results and plot a credible interval at 50% (25-75% percentile).
 
 ```{r  scenario plot , echo=TRUE}
-
-seir1_sim<-seir2_sim<-seir3_sim<-NULL
+seir1_sim<-NULL
 for(s in 1:dim(coef)[1]){
@@
   forecast<-14
-  seir1<-seir2<-seir3<-NULL
+  seir1<-NULL
   for(i in 1:forecast){
-    parameters <- c(mu = mu0, beta = matrix(coef[s,i])+1/duration, sigma = sigma0, gamma = 1/duration)
-    f1<-10
-    if( i==1) initials <- c(S = 0.95, E = (f1*I_start/N), I = I_start/N, R = R_start/N)
+    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
+    if( i==1) initials <- c(S = 1-(f_exp+I_start/N+R_start/N), E = f_exp, I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
@@
     seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
-    f2<-5
-    if( i==1) initials <- c(S = 0.95, E = (f2*I_start/N), I = I_start/N, R = R_start/N)
-    if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
-    seir2_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
-    seir2 <- rbind(seir2,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
-    f3<-3
-    if( i==1) initials <- c(S = 0.95, E = (f3*I_start/N), I = I_start/N, R = R_start/N)
-    if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])
-    seir3_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
-    seir3 <- rbind(seir3,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
   }
   seir1_sim<-rbind(seir1_sim,cbind(rep(s,forecast),seir1))
-  seir2_sim<-rbind(seir2_sim,cbind(rep(s,forecast),seir2))
-  seir3_sim<-rbind(seir3_sim,cbind(rep(s,forecast),seir3))
+
 }
-seir1_sim[,2]<-seir2_sim[,2]<-seir3_sim[,2]<-rep(1:forecast,dim(coef)[1])
-colnames(seir1_sim)[1]<-colnames(seir2_sim)[1]<-colnames(seir3_sim)[1]<-"sim"
+seir1_sim[,2]<-rep(1:forecast,dim(coef)[1])
+colnames(seir1_sim)[1]<-"sim"
 
@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Italy: Scenario 1 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
-p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
-p
-
-### confidence limits - scenario 2
-I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
-I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=limits[1])
-I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=limits[2])
-
-mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
-mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
-mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
-counts<-mu.med
-mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Italy: Scenario 2 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
-p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
-p
-
-### confidence limits - scenario 3
-I_seir_med<-tapply(seir3_sim$I,seir3_sim$time,median)
-I_seir_lwr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=limits[1])
-I_seir_upr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=limits[2])
-
-mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
-mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
-mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
-counts<-mu.med
-mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Italy: Scenario 3 (Credible Interval ",100*confidence,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Italy: Scenario  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
@@
 
-The 3 scenarios show different numbers. 
-If we consider the second scenario, at the end of the 2 weeks (`r date[length(date)]`) the number of infected is  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N)[forecast]`).
+At the end of the 2 weeks (`r date[length(date)]`):
+*the number of infected is  (`r (tapply(seir1_sim$I,seir1_sim$time,median)*N)[forecast]`).
 
-In the next plot the cumulative number of infected.  
-At the end of the 2 weeks (`r date[length(date)]`) the total number of COVID19 cases is expected to be  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N+tapply(seir2_sim$R,seir2_sim$time,median)*N)[forecast]`).
+*the total number of COVID19 cases is expected to be  (`r (tapply(seir1_sim$I,seir1_sim$time,median)*N+tapply(seir1_sim$R,seir1_sim$time,median)*N)[forecast]`).
+
+The estimated (median) numbers of the current scenario by date are:
+```{r  scenario numbers, echo=TRUE}
+S_seir_med<-tapply(seir1_sim$S,seir1_sim$time,median)
+E_seir_med<-tapply(seir1_sim$E,seir1_sim$time,median)
+R_seir_med<-tapply(seir1_sim$R,seir1_sim$time,median)
+forecast<-data.frame(S_seir_med,E_seir_med,I_seir_med,R_seir_med)*N
+colnames(forecast)<-c("Susceptible","Exposed","Infected","Removed")
+rownames(forecast)<-days.ahead
+kable(forecast)
+```

### 2020-03-21 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

@@
 ```
+---
+title: "COVID19 - Forecast and predictions using a time dependent SEIR model - Italy"
+author: "Paolo Girardi"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: rmarkdown::github_document
+always_allow_html: yes
+---
 <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.
@@
 R0 passes from a value of `r round(r0.predict[1,1],2)`   in the initial phase to an  estimated value of `r round(r0.predict[days+forecast,3],2)` at the ending of the 14-days forecast.  
+We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:
 
-We want to make a short term forecast (14 days) with 3 scenario, based on the number of exposed people:
+$$\begin{aligned}
+\frac{d}{dt}S(t) &= \mu (N-S)-\beta \frac{SI}{N}-\nu S\\  
+\frac{d}{dt}E(t) &= \beta \frac{SI}{N}-(\mu+\sigma) E    \\
+\frac{d}{dt}I(t) &=\sigma E- (\mu+\gamma)I \\
+\frac{d}{dt}R(t) &= \gamma\,I(t)-\mu\,R(t)+\nu S.\\
+\end{aligned}$$
 
--Scenario 1: 10 exposed people for each COVID-19 case (no home restrictions made or even no effects)
+<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>
 
--Scenario 2: 5 exposed people for each COVID-19 case (-50% exposed people)
+Respect to the previous formulation, the compartment of exposed people (E) was inserted between the susceptible and infected compartments.  
+(https://en.wikipedia.org/wiki/Bayesian_structural_time_series).
+The parameter $\nu$ is the rate of vaccination.
 
--Scenario 3: 3 exposed people for each COVID-19 case  (-70% exposed people)
+We want to make a short term forecast (14 days).
 
 We made a forecast by means of a SEIR model fixing a series of initial status:
-- S_start=N, the size of Italian population   
-- E= f * I_start (with f a fixed factor of the previous scenario)  
-- I_start: initial number of COVID-19 cases  
-- R_start: initial number of recovered
+*S:N, the size of Italian population  
+*E: The number of exposed people (unknown)  
+*I_start: initial number of COVID-19 cases     
+*R_start: initial number of recovered
 
 and parameters:  
-- beta: gamma+slope $\beta_1$
-- gamma= 1/duration (rate of infection duration of COVID-19, 14 days)  
-- sigma0: the coronavirus transmission rate (half of flu epidemic)   
+- beta: gamma + slope
+- gamma= 1/duration of diseases (duration=21 days)
+- sigma: the coronavirus transmission rate 
 - mu0: the overall mortality rate
 
-```{r initation plot, echo=TRUE}
+```{r ini parameters plot, echo=TRUE}
 # initial number of infectus
 I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
-# initial number of recovered
-R_start<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R_start
-# italian poulation
+# initial number of recovered, based on the proportion of discharged from the health services
+prop<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]/dat_csv$totale_ospedalizzati[dim(dat_csv)[1]]
+R_start<-prop*dat_csv$totale_casi[dim(dat_csv)[1]]; R_start
+# Italian population
 N=60480000
-# duration of COVID19 
-duration<-14
-#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
-sigma0<-0.05
+# duration of COVID19 as the sum of incubation (5-14 days) and the duration of diseases (about 20 days)
+duration<-21
 #mortality rate 
@@
 
-The 3 scenarios show different numbers. 
-If we consider the second scenario at the end of the 2 weeks (`r date[length(date)]`):
-* the number of infected is  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N)[forecast]`).
-* the total number of COVID19 cases is expected to be  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N+tapply(seir2_sim$R,seir2_sim$time,median)*N)[forecast]`).
+At the end of the 2 weeks (`r date[length(date)]`):
+*the number of infected is  (`r (tapply(seir1_sim$I,seir1_sim$time,median)*N)[forecast]`).
+
+*the total number of COVID19 cases is expected to be  (`r (tapply(seir1_sim$I,seir1_sim$time,median)*N+tapply(seir1_sim$R,seir1_sim$time,median)*N)[forecast]`).
+
+The estimated (median) numbers of the current scenario by date are:
+```{r  scenario numbers, echo=TRUE}
+S_seir_med<-tapply(seir1_sim$S,seir1_sim$time,median)
+E_seir_med<-tapply(seir1_sim$E,seir1_sim$time,median)
+R_seir_med<-tapply(seir1_sim$R,seir1_sim$time,median)
+forecast<-data.frame(S_seir_med,E_seir_med,I_seir_med,R_seir_med)*N
+colnames(forecast)<-c("Susceptible","Exposed","Infected","Removed")
+rownames(forecast)<-days.ahead
+kable(forecast)
+```

### 2020-03-21 draft_analysis_Veneto_html.Rmd
Add home page, navigation and theme

--- 2020-03-18 draft_analysis_Veneto_html.Rmd
@@
 
-* We use data provided by  Italian Civil Protection Department and the analysis was restricted to the Veneto Rego
+* We use data provided by  Italian Civil Protection Department and the analysis was restricted to the Veneto Region
 
@@
 basis of the actual data, we used a SEIR model applied to the COVID19
-epidemic to the Veneto Region in Italy
+epidemic to the Veneto Region
 
@@
 R0 passes from a value of `r round(r0.predict[1,1],2)`   in the initial phase to an  estimated value of `r round(r0.predict[days+forecast,3],2)` at the ending of the 14-days forecast.  
+We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:
 
-We want to make a short term forecast (14 days) with 3 scenario, based on the number of exposed people:
+$$\begin{aligned}
+\frac{d}{dt}S(t) &= \mu (N-S)-\beta \frac{SI}{N}-\nu S\\  
+\frac{d}{dt}E(t) &= \beta \frac{SI}{N}-(\mu+\sigma) E    \\
+\frac{d}{dt}I(t) &=\sigma E- (\mu+\gamma)I \\
+\frac{d}{dt}R(t) &= \gamma\,I(t)-\mu\,R(t)+\nu S.\\
+\end{aligned}$$
 
--Scenario 1: 10 exposed people for each COVID-19 case (no home restrictions made or even no effects)
+<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>
 
--Scenario 2: 5 exposed people for each COVID-19 case (-50% exposed people)
+Respect to the previous formulation, the compartment of exposed people (E) was inserted between the susceptible and infected compartments.  
+(https://en.wikipedia.org/wiki/Bayesian_structural_time_series).
+The parameter $\nu$ is the rate of vaccination.
 
--Scenario 3: 3 exposed people for each COVID-19 case  (-70% exposed people)
+We want to make a short term forecast (14 days).
 
 We made a forecast by means of a SEIR model fixing a series of initial status:
-- S0=N, the size of Veneto Region population   
-- E= f * I_start (with f a fixed factor of the previous scenario)  
-- I_start: initial number of COVID-19 cases  
-- R_start: initial number of recovered
+*S:N, the size of the Veneto Region population  
+*E: The number of exposed people (unknown)  
+*I_start: initial number of COVID-19 cases     
+*R_start: initial number of recovered
 
@@
 - beta: gamma + slope
-- gamma= 1/duration (rate of infection duration of COVID-19, duration=14 days)  
-- sigma0: the coronavirus transmission rate (half of flu epidemic)   
+- gamma= 1/duration of diseases (duration=21 days)
+- sigma: the coronavirus transmission rate 
 - mu0: the overall mortality rate
 
-```{r initation plot, echo=TRUE}
+```{r ini parameters plot, echo=TRUE}
 # initial number of infectus
 I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
-# initial number of recovered
-R_start<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]; R_start
-# Veneto Region population
-N=4905854
-# duration of COVID19 
-duration<-14
-#sigma0 is the coronavirus transmission rate fixed to 5%  (half of flu epidemic)
-sigma0<-0.05
+# initial number of recovered, based on the proportion of discharged from the health services
+prop<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]/dat_csv$totale_ospedalizzati[dim(dat_csv)[1]]
+R_start<-prop*dat_csv$totale_casi[dim(dat_csv)[1]]; R_start
+# Veneto population
+N=4905864
+# duration of COVID19 as the sum of incubation (5-14 days) and the duration of diseases (about 20 days)
+duration<-21
 #mortality rate 
@@
 ```
-We use the library(EpiDynamics) and the function SEIR() to implement a SEIR model:
 
-<img src="http://www.public.asu.edu/~hnesse/classes/seireqn.png"/>  
-<img src="https://upload.wikimedia.org/wikipedia/commons/3/3d/SEIR.PNG"/>  
-Respect to the previous formulation, the compartment of exposed people (E) was inserted between the susceptible and infected compartments.  
-(https://en.wikipedia.org/wiki/Bayesian_structural_time_series)
+We try to estimates sigma (from exposed to infected) and the number of exposed population on the basis of the last 5 days of observed number of total infected people.
 
-We perfom simulation on the trend of beta coefficient by means of a Bayesian Structural Time Series using the library bsts of R.    
+We suppose that the number of total infected $I(t)$ can be the realization of a Poisson random variable
+$$
+I(t) \sim Poisson (\theta_t)
+$$
+where $\theta_t$ is the unknown mean of the random process at the time $t$.
+
+The Poisson distribution has a fixed variance that is equal to the mean; in the next part we try to estimates the parameters considering a Negative Binomial distribution which allows for overdispersion.
+
+However, the parameter $\theta_t$ can be supposed to be a realization of a SEIR model with initial parameters $\theta=(\mu,\beta,\sigma, \gamma)$  and initial status $S=(S_{start},E_{start},I_{start},R_{start})$.
+
+The parameter $\sigma$ and the fraction of exposed people ($E_{start}$) are our parameters of interest; we estimate $\sigma$ and $E_{start}$ minimizing the log likelihood of the described Poisson distribution (or Negative Binomial one).
+
+```{r estimation of gamma and exposed, echo=TRUE}
+# number of the last considered days for calibration
+last<-5
+#I_start
+I_1<-dat_csv$totale_attualmente_positivi[(days-last)]
+#R_start
+R_1<-prop*dat_csv$totale_casi[(days-last)]
+#loglikelihood function
+LogLikfun <- function (initials,parameters,obs) {
+n<-length(obs)
+N=4905864
+seir_fit <- SEIR(pars = parameters, init = initials, time = 0:last)
+#Poisson
+#sum(dpois(x=obs[-1],lambda = seir_fit$results$I[-1]*N,log=TRUE))
+#Neg Binomial
+sum(dnbinom(x=obs[-1],mu=seir_fit$results$I[-1]*N,size=N,log=TRUE))
+#SSE
+#-sum((obs[-1]-seir_fit$results$I[-1]*N)^2)
+}
+### logit and its inverse for sigma 
+logit <- function (p) log(p/(1-p))    # the logit transform
+expit <- function (x) 1/(1+exp(-x))   # inverse logit
+
+f1<-function(par){
+f_exp<-expit(par[1])
+sigma<-expit(par[2])
+parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma =1/duration)
+initials <- c(S = (1-(f_exp+I_1/N-R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
+-LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+}
+est<-optim(fn=f1,par=c(logit(0.0005),logit(0.1)))
+#estimated par
+expit(est$par)
+f_exp<-expit(est$par[1])
+sigma<-expit(est$par[2])
+parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma = 1/duration)
+initials <- c(S = (1-(f_exp+I_1/N-R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
+pro<-SEIR(pars = parameters, init = initials, time = 0:last)
+#predicted vs observed
+cbind(pro$results$I*N,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+```
+
+The values of initial parameters are
+```{r ini parameters, echo=TRUE}
+#mu0
+mu0;
+#sigma
+sigma
+#gamma
+1/duration
+```
+
+```{r status, echo=TRUE}
+#mthe unknowrn fraction of exposed people is
+f_exp;
+#that is 
+f_exp/(I_1/N)
+#times the infected
+```
+For the beta parameter, we perfom a simulation on its trend by means of a Bayesian Structural Time Series using the library bsts of R.    
 We estimate a BSTS model specifing a local linear trend ans 1000 simulation.
@@
 ```
-For each vectors of simulated beta coefficients we perform a SEIR model (one for each scenario).  
+For each vector of simulated beta coefficients we perform a SEIR model.
 We save the results and plot a credible interval at 50% (25-75% percentile).
@@
 ```{r  scenario plot , echo=TRUE}
-
-seir1_sim<-seir2_sim<-seir3_sim<-NULL
+seir1_sim<-NULL
 for(s in 1:dim(coef)[1]){
@@
   forecast<-14
-  seir1<-seir2<-seir3<-NULL
+  seir1<-NULL
   for(i in 1:forecast){
-    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma0, gamma = 1/duration)
-    f1<-10
-    if( i==1) initials <- c(S = 0.95, E = (f1*I_start/N), I = I_start/N, R = R_start/N)
+    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
+    if( i==1) initials <- c(S = 1-(f_exp+I_start/N+R_start/N), E = f_exp, I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
@@
     seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
-    f2<-5
-    if( i==1) initials <- c(S = 0.95, E = (f2*I_start/N), I =  I_start/N, R = R_start/N)
-    if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
-    seir2_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
-    seir2 <- rbind(seir2,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
-    f3<-3
-    if( i==1) initials <- c(S = 0.95, E = (f3*I_start/N), I =  I_start/N, R = R_start/N)
-    if( i>1) initials <- c(S = seir3_temp$results$S[2], E = seir3_temp$results$E[2], I =seir3_temp$results$I[2], R = seir3_temp$results$R[2])
-    seir3_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
-    seir3 <- rbind(seir3,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
   }
   seir1_sim<-rbind(seir1_sim,cbind(rep(s,forecast),seir1))
-  seir2_sim<-rbind(seir2_sim,cbind(rep(s,forecast),seir2))
-  seir3_sim<-rbind(seir3_sim,cbind(rep(s,forecast),seir3))
+
 }
-seir1_sim[,2]<-seir2_sim[,2]<-seir3_sim[,2]<-rep(1:forecast,dim(coef)[1])
-colnames(seir1_sim)[1]<-colnames(seir2_sim)[1]<-colnames(seir3_sim)[1]<-"sim"
+seir1_sim[,2]<-rep(1:forecast,dim(coef)[1])
+colnames(seir1_sim)[1]<-"sim"
 
@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
-p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
-p
-
-### confidence limits - scenario 2
-I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
-I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.25)
-I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.75)
-
-mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
-mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
-mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
-counts<-mu.med
-mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Veneto Region: Scenario 2  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
-p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
-p
-
-### confidence limits - scenario 3
-I_seir_med<-tapply(seir3_sim$I,seir3_sim$time,median)
-I_seir_lwr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=0.25)
-I_seir_upr<-tapply(seir3_sim$I,seir3_sim$time,quantile,p=0.75)
-
-mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
-mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
-mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
-counts<-mu.med
-mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Veneto Region: Scenario 3  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Veneto Region: Scenario  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")

### 2020-03-21 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

@@
 ```
+---
+title: "COVID19 - Forecast and predictions using a time dependent SEIR model for the Veneto Region"
+author: "Paolo Girardi"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: rmarkdown::github_document
+always_allow_html: yes
+---
 <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.

### 2020-03-21 BYM_model/INLA_Veneto.R

@@ -26,7 +26,7 @@ dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
 #select a Region, in this example "Lombardia" Code region 3, Veneto 5
 Region<-5
 dat_csv<-dat_csv[dat_csv$codice_regione %in% Region,]
-dat_csv_n$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
+dat_csv$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
 #### number of province
 nprov<-length(table(dat_csv$denominazione_provincia)[table(dat_csv$denominazione_provincia)>0])
 
@@ -96,7 +96,7 @@ plot(Date,exp(fit_2$summary.random$t$mean),ylab="Indicence of cases",xlab="Date"
 ####interaction model type 4 (complete time trend interaction to estimate specific curves for each province)
 ####specification
 formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
-  f(t,model="rw2",constr = FALSE) +
+  f(t,model="rw2",constr = TRUE) +
   f(t2,model="iid") +
   f(ID2,model="besag", graph=file.adj,
     group=t3,
@@ -104,8 +104,9 @@ formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
 #estimation
 fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
 summary(fit_st4)
-fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv,control.compute = list(dic=T))
-summary(fit_st4)
+# BYM sul numero di casi
+#fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv,control.compute = list(dic=T))
+#summary(fit_st4)
 
 #### components
 t_1<-fit_st4$summary.random$t$mean #overall trend
@@ -125,16 +126,21 @@ for ( i in 1:nprov){
 }
 
 nc.province$R0<-R0_vec
-spplot(nc.province, c( "R0"),main="R0, since 5/03")
+spplot(nc.province, c( "R0"),main="R0, period 5 - 19/03")
+levels(nc.province$DEN_PROV)<-c(levels(nc.province$DEN_PROV),"Venezia")
[…]

### 2020-03-21 Previsioni_COVID_2019.R

@@ -1,96 +1,0 @@
-rm(list=ls())
-#base R code for
-###import italian dataset updated 11 March 2020
-dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv",header=T)
-days<-dim(dat_csv)[1]
-dat_csv$t<-1:days
-
-##### SEIR model
-
-#estimate r0
-#see https://kingaa.github.io/clim-dis/parest/parest.html
-#calculate r0 based in the last XX observation 
-beta_vec<-NULL
-sd_vec<-NULL
-for (i in 1:(days-6)){
-fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[i:days,])
-beta_vec<-c(beta_vec,coef(fit)[2])
-sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
-}
-
-label <-days:7
-mean  <- (beta_vec*14+1)
-lower <- ((beta_vec-1.96*sd_vec)*14+1)
-upper <- ((beta_vec+1.96*sd_vec)*14+1)
-
-df <- data.frame(label, mean, lower, upper)
-
-# reverses the factor level ordering for labels after coord_flip()
-df$label <- factor(df$label, levels=rev(df$label))
-
-library(ggplot2)
-fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
-  geom_pointrange() + 
-  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
-  coord_flip() +  # flip coordinates (puts labels on y axis)
-  xlab("Last days used for the calculation") + ylab("R0 Mean (95% CI)") +
-  theme_bw()  # use a white background
-print(fp)
-
-
-
-
-fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv)
-
-# stimo 
-
-plot(dat_csv$t,log(dat_csv$totale_attualmente_positivi))
-abline(coef(summary(fit1))[,1])
-summary(fit1)
-# R0
-slope <-coef(summary(fit1))[2,1]; slope
-slope.se <- coef(summary(fit1))[2,2]; slope.se
-###R0 incubation time of 14 days
-R_0=slope*14+1;R_0
-(slope+c(-1,1)*1.96*slope.se)*14+1
-
[…]

### 2020-03-21 BYM_model/INLA_Veneto.R
Add home page, navigation and theme

@@ -26,7 +26,7 @@ dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
 #select a Region, in this example "Lombardia" Code region 3, Veneto 5
 Region<-5
 dat_csv<-dat_csv[dat_csv$codice_regione %in% Region,]
-dat_csv$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
+dat_csv$denominazione_provincia<-droplevels(dat_csv$denominazione_provincia)
 #### number of province
 nprov<-length(table(dat_csv$denominazione_provincia)[table(dat_csv$denominazione_provincia)>0])
 
@@ -81,7 +81,7 @@ spplot(nc.province, c( "IRR_mean"))
 ########Pure temporal model (RW2)
 #RW2 temporal model
 Date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days-1,1)
-formula_t = totale_casi ~ 1+f(t,model="rw2", constr = FALSE)+f(t2,model="iid")
+formula_t = totale_casi ~ 1+f(t,model="rw2")+f(t2,model="iid")
 fit_2<-inla(formula_t, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
 summary(fit_2)
 plot(Date,fit_2$summary.random$t$mean,ylab="Temp Coef",xlab="Date")
@@ -90,8 +90,8 @@ lm_fit<-lm(fit_2$summary.random$t$mean~I(1:days),subset=-1)
 coef(lm_fit)[2]
 plot(1:days,fit_2$summary.random$t$mean,ylab="Temp Coef",xlab="Date")
 abline(lm_fit)
-##Incidence rate ratios, the rate of increase od cases respect to time 1
-plot(Date,exp(fit_2$summary.random$t$mean),ylab="Indicence of cases",xlab="Date")
+##Incidence rate ratios, the rate of increase od cases respect to average.
+plot(Date,exp(fit_2$summary.random$t$mean),ylab="IRR",xlab="Date")
 
 ####interaction model type 4 (complete time trend interaction to estimate specific curves for each province)
 ####specification
@@ -140,7 +140,7 @@ legend("topleft",c("Veneto",paste(1:nprov,nc.province$DEN_PROV)),col=1:(nprov+1)
 
 
 ##################predictions 3 days forward
-Forecast=14
+Forecast=3
 dat_csv2<-dat_csv 
 dat_csv2$t<-rep(1:days+days,nprov)
 dat_csv2$totale_casi<-NA

### 2020-03-21 BYM_model/INLA_def.R
Add home page, navigation and theme

@@ -24,9 +24,9 @@ dat_csv<-read.csv("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dat
 dat_csv<-merge(dat_csv,pop_provincia,by=c("codice_provincia"),all.y=TRUE)
 dat_csv<-dat_csv[dat_csv$codice_provincia<112,]
 #select a Region, in this example "Lombardia" Code region 3
-Region<-3
-dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
-dat_csv_n$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
+#Region<-3
+#dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
+dat_csv_n$denominazione_provincia<-droplevels(dat_csv$denominazione_provincia)
 #### number of province
 nprov<-length(table(dat_csv$denominazione_provincia)[table(dat_csv$denominazione_provincia)>0])
 

### 2020-03-21 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

@@
-```{r echo=FALSE}
-library(knitr)
-knitr::opts_chunk$set(echo = FALSE)
-knitr::opts_knit$set(root.dir= normalizePath('..'))
-knitr::opts_chunk$set(error = FALSE)
-```
 ---
@@
 
-*NB: set the file output format to
-
-#output:html_document:  
-  df_print: paged  
-pdf_document:   
-  toc: yes
-
-which  performs the same analysis enabling Javascript Pictures.
-
 ## The COVID dataset
@@
 checkpackage("ggplot2")
+checkpackage("knitr")
 ```
@@
 library(ggplot2)
+library(knitr)
 ```

### 2020-03-23 draft_analysis_Italy_html.Rmd
Add home page, navigation and theme

@@
 ---
-title: "COVID19 - Forecast and predictions using a time dependent SEIR model"
+title: "COVID19 - Forecast and predictions using a time dependent SEIR model - Italy"
 author: "Paolo Girardi"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:html_document:
-  df_print: paged
-pdf_document: 
-  toc: yes
+output: rmarkdown::github_document
+always_allow_html: yes
 ---
@@
 checkpackage("ggplot2")
+checkpackage("knitr")
+checkpackage("splines")
 ```
@@
 library(ggplot2)
+library(knitr)
+library(splines)
 ```
@@
 
-However, in the calculation of R0 we considered an infectious period of 21 days for two reasons:  
-1) the majority of cases is asymptomatic;
-2) 21 days is the worst scenario because in this period COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).
+However, in the calculation of R0 we considered an infectious period of 18 days (https://www.nejm.org/doi/10.1056/NEJMoa2001316).
 
@@
 #vector for beta and standard deviation
-duration<-21
+duration<-18
 beta_vec<-NULL
@@
 ```{r R0 forecast}
-
 time<-3:(days-2)
-beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian)
+weekend<-rep(c(1,2,3,4,5,6,7),ceiling(days/7))[3:(days-2)]
+data=data.frame(time,weekend)
+beta.model<-glm(beta_vec~time+bs(weekend),weights = 1/sd_vec,family=gaussian,data=data)
+summary(beta.model)
+anova(beta.model,test="Chisq")
+#there is an effect of the week, as supposed 
 forecast=14
 # add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
-pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
+weekend_pre<-rep(c(1,2,3,4,5,6,7),ceiling((days+forecast)/7))[1:(days+forecast)]
+datanew<-data.frame(time=1:(days+forecast),weekend=weekend_pre)
+pre<-predict(beta.model,type='response',newdata=datanew,se.fit=TRUE)
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
+predict <- data.frame(beta_vec=c(rep(NA,2),beta_vec,rep(NA,forecast+2)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
 beta.predict<-predict 
 r0.predict<-beta.predict
-r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*21+1
+r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*duration+1
 # plot the points (actual observations), regression line, and confidence interval
@@
 N=60480000
-# duration of COVID19 as the sum of incubation (5-14 days) and the duration of diseases (about 20 days)
-duration<-21
+# duration of COVID19 as the sum of incubation (5-14 days) and the duration of diseases (about 18 days)
+duration<-18
 #mortality rate 
@@
 # number of the last considered days for calibration
-last<-5
+last<-10
 #I_start
@@
 sigma<-expit(par[2])
-parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma =1/duration)
+parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-2)])+1/duration), sigma = sigma, gamma =1/duration)
 initials <- c(S = (1-(f_exp+I_1/N-R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
@@
 sigma<-expit(est$par[2])
-parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma = 1/duration)
+parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-2)])+1/duration), sigma = sigma, gamma = 1/duration)
 initials <- c(S = (1-(f_exp+I_1/N-R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
@@
 #mthe unknowrn fraction of exposed people is
-f_exp;
+pro$results$E[last+1];
 #that is 
-f_exp/(I_1/N)
+pro$results$E[last+1]/(I_1/N)
 #times the infected
@@
 ss <- AddLocalLinearTrend(list(), beta_vec)
+ss <- AddSeasonal(ss, beta_vec, nseasons = 7)
 model1 <- bsts(beta_vec,
                state.specification = ss,
-               niter = 1000)
+               niter = 1000,seed=123)
 par(mfrow = c(1,1))
-plot(model1, "components", ylab="Beta1 coefficient", xlab="days")
+plot(model1, "components")
 #previsioni
@@
 par(mfrow = c(1,1))
-plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions")
+plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions",ylim=c(-1,1)*IQR(pred1$distribution)+median(pred1$distribution))
 # matrix of beta coefficients
@@
     parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
-    if( i==1) initials <- c(S = 1-(f_exp+I_start/N+R_start/N), E = f_exp, I = I_start/N, R = R_start/N)
+    if( i==1) initials <- c(S = 1-(pro$results$E[last+1]+I_start/N+R_start/N), E =pro$results$E[last+1] , I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])

### 2020-03-23 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

@@
 checkpackage("knitr")
+checkpackage("splines")
 ```
@@
 library(knitr)
+library(splines)
 ```

### 2020-03-23 draft_analysis_Veneto_html.Rmd
Add home page, navigation and theme

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:html_document:
-  df_print: paged
-pdf_document: 
-  toc: yes
+output: rmarkdown::github_document
+always_allow_html: yes
 ---
@@
 * The layout of the draft must definitely be improved.
-
-*NB: set the file output format to
-
-#output:html_document:  
-  df_print: paged  
-pdf_document:   
-  toc: yes
-
-which  performs the same analysis enabling Javascript Pictures.
 
@@
 ```
+The time series of COVID-19 in the Veneto Region was influenced by the presence of a hotspot in the small village of Vo Euganeo.
 
+```{r}
+plot(days_dy[-1],diff(dat_csv$totale_casi), ylab="New cases",xlab="Date",type="l")
+```
 ### The S(E)IR model (to be revised)
@@
 $$
-\hat{R_0}=\frac{\hat{\beta}}{\gamma}
+\hat{R_0}=\frac{\beta}{\gamma}
 $$
@@
 
-However, in the calculation of R0 we considered an infectious period of 21 days for two reasons:  
-1) the majority of cases is asymptomatic;
-2) 21 days is the worst scenario because in this period COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).
+However, in the calculation of R0 we considered an infectious period of 18 days on the basis of recent publications (https://www.nature.com/articles/s41421-020-0148-0).
 
@@
 ```{r R0 forecast}
-
+#start from the day 7, since vo hotspot can influence the trend
 time<-3:(days-2)
-beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian)
+weekend<-rep(c(1,2,3,4,5,6,7),ceiling(days/7))[3:(days-2)]
+data<-data.frame(time)
+beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian,data)
+summary(beta.model)
+anova(beta.model,test="Chisq")
+#there is an effect of the week, as supposed 
+#we can use
+#summary(gamlss(formula = beta_vec ~ time + bs(weekend, 3),      weights = 1/sd_vec,Lp=1) )
 forecast=14
@@
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
+predict <- data.frame(beta_vec=c(rep(NA,2),beta_vec,rep(NA,forecast+2)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
 beta.predict<-predict 
 r0.predict<-beta.predict
-r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*21+1
+r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*duration+1
 # plot the points (actual observations), regression line, and confidence interval
@@
 N=4905864
-# duration of COVID19 as the sum of incubation (5-14 days) and the duration of diseases (about 20 days)
-duration<-21
+#  duration of COVID19 diseases 
+duration<-18
 #mortality rate 
@@
 #Poisson
-#sum(dpois(x=obs[-1],lambda = seir_fit$results$I[-1]*N,log=TRUE))
+sum(dpois(x=obs[-1],lambda = seir_fit$results$I[-1]*N,log=TRUE))
 #Neg Binomial
-sum(dnbinom(x=obs[-1],mu=seir_fit$results$I[-1]*N,size=N,log=TRUE))
+#sum(dnbinom(x=obs[-1],mu=seir_fit$results$I[-1]*N,size=N,log=TRUE))
 #SSE
@@
 f_exp<-expit(par[1])
-sigma<-expit(par[2])
-parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma =1/duration)
-initials <- c(S = (1-(f_exp+I_1/N-R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
+#max 0.33 according to https://www.nejm.org/doi/10.1056/NEJMoa2001316 is 1/5.2
+sigma<-expit(par[2])/3
+parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[days-last-2])+1/duration), sigma = sigma, gamma =1/duration)
+initials <- c(S = (1-(f_exp+I_1/N+R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
 -LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
 }
-est<-optim(fn=f1,par=c(logit(0.0005),logit(0.1)))
+est<-optim(fn=f1,par=c(logit(0.02),logit(0.2)))
 #estimated par
 expit(est$par)
+
 f_exp<-expit(est$par[1])
-sigma<-expit(est$par[2])
-parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma = 1/duration)
-initials <- c(S = (1-(f_exp+I_1/N-R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
+sigma<-expit(est$par[2])/3
+parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[days-last-2])+1/duration), sigma = sigma, gamma = 1/duration)
+initials <- c(S = (1-(f_exp+I_1/N+R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
 pro<-SEIR(pars = parameters, init = initials, time = 0:last)
@@
 #mthe unknowrn fraction of exposed people is
-f_exp;
+pro$results$E[last+1]
 #that is 
-f_exp/(I_1/N)
+pro$results$E[last+1]/(I_1/N)
 #times the infected
 ```
-For the beta parameter, we perfom a simulation on its trend by means of a Bayesian Structural Time Series using the library bsts of R.    
-We estimate a BSTS model specifing a local linear trend ans 1000 simulation.
+
+# Forecast 1 - Results based on beta coefficient trend
+
+For the beta parameter, we perfomed a simulation on its trend by means of a Bayesian Structural Time Series using the library bsts of R.    
+We estimated a BSTS model specifing a local linear trend and a seasonality component. The model predicted 1000 trajectories forward of the beta coefficent.
 
@@
 # Bayesian Structural Time Series
-ss <- AddLocalLinearTrend(list(), beta_vec)
-model1 <- bsts(beta_vec,
+#we drop the first 6 beta_vec
+ss <- AddLocalLinearTrend(list(), beta_vec[6:(days-4)])
+ss <- AddSeasonal(ss, beta_vec[6:(days-4)], nseasons = 7)
+model1 <- bsts(beta_vec[6:(days-4)],
                state.specification = ss,
-               niter = 1000)
+               niter = 1000,seed=123)
 par(mfrow = c(1,1))
-plot(model1, "components", ylab="Beta1 coefficient", xlab="days")
+plot(model1, "components")
 #previsioni
@@
 ```
+
 For each vector of simulated beta coefficients we perform a SEIR model.
@@
     parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
-    if( i==1) initials <- c(S = 1-(f_exp+I_start/N+R_start/N), E = f_exp, I = I_start/N, R = R_start/N)
+    if( i==1) initials <- c(S = 1-(pro$results$E[last+1]+I_start/N+R_start/N), E = pro$results$E[last+1], I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Veneto Region: Scenario  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
@@
 The estimated (median) numbers of the current scenario by date are:
+
 ```{r  scenario numbers, echo=TRUE}
@@
 ```
+
+# Scenario 2 - Forecast based on the last beta coefficient
+
+We estimate a second scenario using a Random Walk on the basis of the last value of the beta.
+
+```{r  scenario2 sim and plot , echo=TRUE}
+ss<-NULL
+set.seed(123)
+ss <- AddLocalLinearTrend(list(), rnorm(100,beta_vec[days-4],sd_vec[days-4]))
+model1 <- bsts(beta_vec[6:(days-4)],
+               state.specification = ss,
+               niter = 1000,seed=123)
+par(mfrow = c(1,1))
+plot(model1, "components")
+#previsioni
+pred1 <- predict(model1, horizon = 16, burn = 100)
+par(mfrow = c(1,1))
+plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions",ylim=c(-1,1)*IQR(pred1$distribution)+median(pred1$distribution))
+# matrix of beta coefficients
+coef<-(pred1$distribution[,3:16])
+par(mfrow = c(1,1))
+
+seir2_sim<-NULL
+for(s in 1:dim(coef)[1]){
+  # average number of single connections of an infected person
+  # less contacts, less probability of new infections
+  # we keep constant the other parameters
+  forecast<-14
+  seir2<-NULL
+  for(i in 1:forecast){
+    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
+    if( i==1) initials <- c(S = 1-(pro$results$E[last+1]+I_start/N+R_start/N), E = pro$results$E[last+1], I = I_start/N, R = R_start/N)
+    if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
+    seir2_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
+    seir2 <- rbind(seir2,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
+  }
+  seir2_sim<-rbind(seir2_sim,cbind(rep(s,forecast),seir2))
+
+}
+seir2_sim[,2]<-rep(1:forecast,dim(coef)[1])
+colnames(seir2_sim)[1]<-"sim"
+
+### confidence limits
+I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
+I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.25)
+I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.75)
+
+days.before<-date[1:days]
+days.ahead<-date[(days+1):(days+forecast)]
+step.ahead<-forecast+1
+mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+counts<-mu.med
+mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 2 (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
+p
+```
+
+At the end of the 2 weeks (`r date[length(date)]`):
+*the number of infected is  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N)[forecast]`).
+
+*the total number of COVID19 cases is expected to be  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N+tapply(seir2_sim$R,seir2_sim$time,median)*N)[forecast]`).
+
+The estimated (median) numbers of the second scenario by date are:
+```{r  scenario2 numbers, echo=TRUE}
+S_seir_med<-tapply(seir2_sim$S,seir2_sim$time,median)
+E_seir_med<-tapply(seir2_sim$E,seir2_sim$time,median)
+R_seir_med<-tapply(seir2_sim$R,seir2_sim$time,median)
+forecast<-data.frame(S_seir_med,E_seir_med,I_seir_med,R_seir_med)*N
+colnames(forecast)<-c("Susceptible","Exposed","Infected","Removed")
+rownames(forecast)<-days.ahead
+kable(forecast)
+```
+
+# Number of admitted in intensive care
+
+The number of person admitted in intensive care repart is a fraction of the COVID 19 case.
+The trend can be estimated by means a binomial regression model.  
+We report the number of intensive care patients on the basis of the estimated COVID19 cases of  the previous scenario 1 and 2.
+
+```{r  intensive care numbers, echo=TRUE}
+fit_int_care<-glm(cbind(terapia_intensiva,totale_attualmente_positivi-terapia_intensiva)~t,family=binomial,data=dat_csv)
+
+forecast=14
+# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
+pre<-predict(fit_int_care,type='response',newdata=data.frame(t=1:(days+forecast)),se.fit=TRUE)
+date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
+plot(date,pre$fit,type="l",main="Proportion of admitted COVID 19 in intensive care",ylim=range(dat_csv$terapia_intensiva/dat_csv$totale_attualmente_positivi))
+points(date[1:days],dat_csv$terapia_intensiva/dat_csv$totale_attualmente_positivi)
+lines(date,pre$fit+1.96*pre$se.fit,col=3,lty=2)
+lines(date,pre$fit-1.96*pre$se.fit,col=3,lty=2)
+legend("topright",c("fitted","95% Confidence Interval"),lty=1:2,col=c(1,3))
+
+days.before<-date[1:days]
+days.ahead<-date[(days+1):(days+forecast)]
+step.ahead<-forecast+1
+
+#Scenario 1
+
+I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
+I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.25)
+I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.75)
+
+mu.lower<-c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit-1.645*pre$se.fit)[(days+1):(days+forecast)]))
+mu.upper<-c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit+1.645*pre$se.fit)[(days+1):(days+forecast)]))
+mu.med<-xts(c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit)[(days+1):(days+forecast)])),order.by = c(days.before,days.ahead),frequency = 7)
+counts<-mu.med
+mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Confidence Interval ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
+p
+#Numbers
+intensive_care<-data.frame(mu.med,mu.lower,mu.upper)[(days+1):(days+forecast),]
+kable(intensive_care)
+
+#Scenario 2
+
+I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
+I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.25)
+I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.75)
+
+mu.lower<-c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit-1.645*pre$se.fit)[(days+1):(days+forecast)]))
+mu.upper<-c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit+1.645*pre$se.fit)[(days+1):(days+forecast)]))
+mu.med<-xts(c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit)[(days+1):(days+forecast)])),order.by = c(days.before,days.ahead),frequency = 7)
+counts<-mu.med
+mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 2 (Confidence Interval ",100*0.9,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
+p
+#Numbers
+intensive_care<-data.frame(mu.med,mu.lower,mu.upper)[(days+1):(days+forecast),]
+kable(intensive_care)
+```

### 2020-03-23 functions.R

@@ -1,0 +1,18 @@
+
+SEIR_betavar<-function(theta, B, status,fw){
+#pars=parameters
+  #beta=expit(est$par[3])-0.5
+  #status=initials
+  #forecast=14
+seir1<-NULL
+if(length(B)==1) B=rep(B,fw)
+for(i in 1:fw){
+parameters <- c(mu=theta[1], beta=B[i], sigma=theta[2], gamma=theta[3])
+if( i==1) initials <- c(S=status[1], E=status[2], I=status[3], R=status[4])
+if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
+seir1_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
+seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
+}
+seir1$time<-1:fw
+seir1
+}

### 2020-03-24 draft_analysis_Italy_html.Rmd
Add home page, navigation and theme

@@
 data=data.frame(time,weekend)
-beta.model<-glm(beta_vec~time+bs(weekend),weights = 1/sd_vec,family=gaussian,data=data)
+beta.model<-glm(beta_vec~time+I(cos(time*2*pi/7))+I(sin(time*2*pi/7)),weights = 1/sd_vec,family=gaussian,data=data)
 summary(beta.model)
@@
 *S:N, the size of Italian population  
-*E: The number of exposed people (unknown)  
+*E: The number of exposed people, but supposed to be at least 3 times the infected (3 x I_start)  
 *I_start: initial number of COVID-19 cases       
@@
 
-The parameter $\sigma$ and the fraction of exposed people ($E_{start}$) are our parameters of interest; we estimate $\sigma$ and $E_{start}$ minimizing the log likelihood of the described Poisson distribution (or Negative Binomial one).
+The parameter $\sigma$ is our parameter of interest; we estimate $\sigma$ fixing $E_{start}= 3*I_{start}$ minimizing the log likelihood of the described Poisson distribution (or Negative Binomial one).
 
@@
 f1<-function(par){
-f_exp<-expit(par[1])
-sigma<-expit(par[2])
+sigma<-expit(par)
+f_exp<-3*I_1/N
 parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-2)])+1/duration), sigma = sigma, gamma =1/duration)
@@
 }
-est<-optim(fn=f1,par=c(logit(0.0005),logit(0.1)))
+est<-optim(fn=f1,par=logit(0.1))
 #estimated par
 expit(est$par)
-f_exp<-expit(est$par[1])
-sigma<-expit(est$par[2])
+sigma<-expit(est$par)
+f_exp<-3*I_1/N
 parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-2)])+1/duration), sigma = sigma, gamma = 1/duration)
@@
 ```{r status, echo=TRUE}
-#mthe unknowrn fraction of exposed people is
+#the unknowrn fraction of exposed people is
 pro$results$E[last+1];
 #that is 
-pro$results$E[last+1]/(I_1/N)
+pro$results$E[last+1]/pro$results$I[last+1]
 #times the infected

### 2020-03-24 draft_analysis_Veneto_html.Rmd
Add home page, navigation and theme

@@
 *S:N, the size of the Veneto Region population  
-*E: The number of exposed people (unknown)  
+*E: The number of exposed people, but supposed to be at least 3 times the infected (3 x I_start)  
 *I_start: initial number of COVID-19 cases     
@@
 and parameters:   
-- beta: gamma + slope
-- gamma= 1/duration of diseases (duration=21 days)
-- sigma: the coronavirus transmission rate 
-- mu0: the overall mortality rate
+-beta: gamma + slope   
+-gamma= 1/duration of diseases (duration=21 days)   
+-sigma: the coronavirus transmission rate   
+-mu0: the overall mortality rate
 
@@
 f1<-function(par){
-f_exp<-expit(par[1])
-#max 0.33 according to https://www.nejm.org/doi/10.1056/NEJMoa2001316 is 1/5.2
-sigma<-expit(par[2])/3
-parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[days-last-2])+1/duration), sigma = sigma, gamma =1/duration)
+#max 0.20 according to https://www.nejm.org/doi/10.1056/NEJMoa2001316 is 1/5.2
+sigma<-expit(par)
+f_exp<-3*I_1/N
+parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[(days-last-2)])+1/duration), sigma = sigma, gamma =1/duration)
 initials <- c(S = (1-(f_exp+I_1/N+R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
@@
 }
-est<-optim(fn=f1,par=c(logit(0.02),logit(0.2)))
-#estimated par
+est<-optim(fn=f1,par=logit(0.1))
+est
 expit(est$par)
 
-f_exp<-expit(est$par[1])
-sigma<-expit(est$par[2])/3
-parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[days-last-2])+1/duration), sigma = sigma, gamma = 1/duration)
+f_exp<-3*I_1/N
+sigma<-expit(est$par)
+
+parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[(days-last-2)])+1/duration), sigma = sigma, gamma = 1/duration)
 initials <- c(S = (1-(f_exp+I_1/N+R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
@@
 #that is 
-pro$results$E[last+1]/(I_1/N)
+pro$results$E[last+1]/pro$results$I[last+1]
 #times the infected
@@
 #we drop the first 6 beta_vec
-ss <- AddLocalLinearTrend(list(), beta_vec[6:(days-4)])
+ss <- AddStudentLocalLinearTrend(list(), beta_vec[6:(days-4)])
 ss <- AddSeasonal(ss, beta_vec[6:(days-4)], nseasons = 7)
@@
 set.seed(123)
-ss <- AddLocalLinearTrend(list(), rnorm(100,beta_vec[days-4],sd_vec[days-4]))
-model1 <- bsts(beta_vec[6:(days-4)],
+ss <-   AddLocalLevel(list(), beta_vec[(days-9):(days-4)],sigma=SdPrior(sd_vec[(days-4)]))
+model1 <- bsts(beta_vec[(days-9):(days-4)],
                state.specification = ss,
@@
 I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
-I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.25)
-I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.75)
+I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.05)
+I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.95)
 
@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Veneto Region: Scenario 2 (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 2 (Credible Interval ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
@@
 I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
-I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.25)
+I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.05)
 I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.75)

### 2020-03-24 BYM_model/INLA_def.Rmd
Add home page, navigation and theme

--- 2020-03-17 BYM_model/INLA_def.Rmd
@@
 ---
-title: "COVID19 - Forecast and predictions using a BYM model - Veneto Region"
+title: "COVID19 - Forecast and predictions using a BYM model in Italy"
 author: "Paolo Girardi"
@@
 ```{r echo=FALSE}
+library(knitr)
+knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_knit$set(root.dir= normalizePath('..'))
+knitr::opts_chunk$set(error = FALSE)
+```
+<a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.
+
+# Disclaimer
+
+* We want to investigate the evolution of the coronavirus pandemic in  Italy
+
+* Our point of view is that of surveillance with the goal of detecting important changes in the underlying (random) process as soon as possible after it has occured.
+
+* We use data provided by  Italian Civil Protection Department and the analysis was restricted to the Lombardy Region
+
+* This document is in a draft mode, and it is continuously updated.
+
+* The layout of the draft must definitely be improved.
+
+*NB: set the file output format to
+
+#output:html_document:  
+  df_print: paged  
+pdf_document:   
+  toc: yes
+
+which  performs the same analysis enabling Javascript Pictures.
+
+## The COVID dataset
+
+The present analysis started from the dataset on COVID19 updated in
+<https://github.com/pcm-dpc/COVID-19>, database provided by the Italian
+Civil Protection.
+
+# Software
+
+Install  packages `dygraphs`,  `xts` and `EpiDynamics` if not available
+
+```{r echo=TRUE}
+checkpackage <- function(package) {
+  if (!package %in% installed.packages()) install.packages(package)
+}
+checkpackage("dygraphs")
+checkpackage("xts")
+checkpackage("ggplot2")
+checkpackage("spdep")
+checkpackage("maptools")
+checkpackage("INLA")
+checkpackage("rgdal")
+```
+and load them.
+```{r, warning=FALSE}
 #Load libraries
-library(spdep)
-library(maptools)
-library(INLA)
-library(rgdal)
-library(foreign)
-library(dygraphs)
-library(xts)
+library("dygraphs")
+library("xts")
+library("ggplot2")
+library("spdep")
+library("maptools")
+library("INLA")
+library("rgdal")
 ```
-################################ IMPORT DATASETS
-# ISTAT Province dataset
+# Datasets loading
 ```{r}
-db_istat<-read.csv("Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
+db_istat<-read.csv("https://raw.githubusercontent.com/Paolin83/COVID19_SEIR_model/master/BYM_model/Elenco-comuni-italiani2.csv",sep=";",header=T,dec=",")
 pop_provincia<-data.frame(codice_provincia=as.numeric(names(table(db_istat[,3]))),pop=tapply(db_istat[,20],db_istat[,3],sum))
@@
 #select a Region, in this example "Lombardia" Code region 3
-Region<-3
-dat_csv<-dat_csv[dat_csv$codice_regione==Region,]
-dat_csv_n$denominazione_provincia<-droplevels(dat_csv_n$denominazione_provincia)
+Region<-c(3)
+dat_csv<-dat_csv[dat_csv$codice_regione %in%Region,]
+dat_csv$denominazione_provincia<-droplevels(dat_csv$denominazione_provincia)
 #### number of province
@@
 dat_csv$ID<-as.numeric(dat_csv$ID)
-#calculate new_cases
+#calculate new_cases Y ~ Poi(mu)
 dat_csv$cumulative_cases<-dat_csv$totale_casi
@@
 dat_csv$totale_casi[dat_csv$totale_casi<0]<-0
-
-#import dataset shapefile Italian province
-nc.province <- readShapePoly(paste(wd,"/ProvCM01012019_g/ProvCM01012019_g_WGS84.shp",sep=""))
-nc.province<-nc.province[nc.province$COD_REG==Region,]
+#import dataset shapefile Italian province and adjancies matrix
+#### IMPORTANT, SET THE Work Directory of the shapefile 
+setwd("/Users/Paolo/My_Website/public/COVID19_SEIR_model")
+nc.province <- readShapePoly("BYM_model/ProvCM01012019_g/ProvCM01012019_g_WGS84.shp")
+nc.province<-nc.province[nc.province$COD_REG %in% Region,]
 nc.province<-nc.province[order(nc.province$COD_PROV),]
@@
 
-################################
-#prior distribution for iid and besag components
+# Spatial analysis with R-INLA and BYM   
+We modelled COVID 19 cases by means of a BYM (Besag, York and Molli\'e) model using an Integrated Nested Laplace Approximation (INLA).   
+\textbf{Starting phase}
+
+For the $i$-th Nuts-3 Region (Italian province), the observed number of COVID19, \textbf{$y_{i}$}, was modelled as follows:  
+$$
+y_{i} \sim Poisson (\lambda_{i})\\
+$$
+with i=1,..., `r nprov`. We modelled the incidence of COVID-19 by means of a BYM model including the population size as offset variable as:  
+$$
+	log(\frac{\lambda_{i}}{N_{i}})=\alpha+\mu_i+\nu_i,\\
+$$
+where $\alpha$ is the intercept, $\mu_i$ and $\nu_i$ are two area specific effects with normal distribution modelled using an intrinsic conditional autoregressive structure (iCAR) and $N_i$ is the population size of each NUTS-3 Region.
+
+The parameter were estimated by INLA (Integrated Nested Laplace Approximation) and R software.
+
+```{r}
+nomi_prov<-levels(dat_csv$denominazione_provincia)
+barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum),names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number fo cases")
+```
+The number of observed cases $y_i$ is very different among  NUTS-3 regions.
+
+```{r}
+pop<-tapply(dat_csv$pop,dat_csv$denominazione_provincia,mean)
+barplot(tapply(dat_csv$totale_casi,dat_csv$denominazione_provincia,sum)/pop*1000,names =1:12,args.legend = list(x = "topright"),legend.text =paste(1:12,nomi_prov),ylab="Number of cases x 1.000 inhabitants")
+```
+A correct comparison is the made by incidence cases $y_i/N_i$, here riported for x1000 inhabitants. 
+#Spatial model  
+We estimate a BYM model fixing prior distributions for iid and besag components as follows:  
 ```{r}
@@
 ```
-####COVID19 MAPS BYM with INLA
+We create new variables that are required by INLA procedure.  
+```{r}
 ##########################indexing, for 
 #### number of days
-```{r}
 dat_csv$t<-as.numeric(dat_csv$data)
@@
 dat_csv$ID2<-dat_csv$ID
-##########################
+#### province
+dat_csv$IDt<-1:dim(dat_csv)[1]
+########################## formula BYM
 formula.bym= totale_casi ~ 1+f(ID, model = "bym", graph = file.adj, param = c(prior.iid, prior.besag), initial = c(initial.iid, initial.besag))
@@
 nc.province$IRR_mean<-IRR_mean
-spplot(nc.province, c( "IRR_mean"))
+nc.province$DEN_PROV<-droplevels(nc.province$DEN_PROV)
+
+spplot(nc.province, c( "IRR_mean"),main="Incidence Rate Ratios", sp.layout=
+         list("sp.text", coordinates(nc.province), nc.province$DEN_PROV, cex=0.7, col="gray"))
+
 ```
-#Pure temporal model (RW2)
-RW2 temporal model
+The image reported the IRR=exp($\mu_i$), the increase of the Incidence of Covid-19 in each NUTS-3 Region respect to the overall mean.
+
+#Temporal model   
+We consider the temporal aspect estimating a Bayesian RW2 model to model the temporal trend of COVID-19 in the reported temporal window.
+The observed number of COVID-19 a the time $t$, \textbf{$y_{t}$}, is modelled as follows
+$$
+y_{t} \sim Poisson (\lambda_{t})
+$$
+with t=1,...,`r days`.
+\end{center}
+where the quantity $\frac{\lambda_{t}}$, is modelled by a Random Walk of order 2 variable
+$$
+log({\lambda_{it}})=\alpha+\gamma_t+\phi_t, 
+$$
+where $\alpha$ is the intercept and $\gamma_t \sim RW2$  are the coefficients related to the random walk process while $\phi_t \sim N(0, \tau_t^{-1} )$ are the temporal specific random errors.
 ```{r}
 Date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days-1,1)
+plot(Date,tapply(dat_csv$totale_casi,dat_csv$t,sum),xlab="Date",ylab="New cases",type="l")
+points(Date,tapply(dat_csv$totale_casi,dat_csv$t,sum))
 formula_t = totale_casi ~ 1+f(t,model="rw2", constr = FALSE)+f(t2,model="iid")
-fit_2<-inla(formula_t, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
+fit_2<-inla(formula_t, family="poisson", data=dat_csv,control.compute = list(dic=T))
 summary(fit_2)
-plot(Date,fit_2$summary.random$t$mean,ylab="Temp Coef",xlab="Date")
-#estimating a linear model without the first obs
-lm_fit<-lm(fit_2$summary.random$t$mean~I(1:days),subset=-1)
-coef(lm_fit)[2]
-plot(1:days,fit_2$summary.random$t$mean,ylab="Temp Coef",xlab="Date")
-abline(lm_fit)
-##Incidence rate ratios, the rate of increase od cases respect to time 1
-plot(Date,exp(fit_2$summary.random$t$mean),ylab="Indicence of cases",xlab="Date")
+
+##Incidence rate ratios, increase od cases respect to time 1
+plot(Date,exp(fit_2$summary.random$t$mean),ylab="temporal IRR",xlab="Date")
+
+label<-Date
+
+mean  <- exp(fit_2$summary.random$t$mean[1:days]+fit_2$summary.fixed$mean)*nprov
+lower <- exp(fit_2$summary.random$t$mean[1:days]-1.96*fit_2$summary.random$t$sd[1:days]+fit_2$summary.fixed$mean)*nprov
+upper <- exp(fit_2$summary.random$t$mean[1:days]+1.96*fit_2$summary.random$t$sd[1:days]+fit_2$summary.fixed$mean)*nprov
+
+df <- data.frame(label, mean, lower, upper)
+
+fp <- ggplot(data=df, aes(x=label, y=mean, ymin=lower, ymax=upper)) +
+  geom_pointrange() +
+  geom_hline(yintercept=1, lty=2) +  # add a dotted line at x=1 after flip
+  xlab("Date") + ylab("Estimated new cases") +
+  theme_bw() 
+print(fp)
+
+```
+We now consider a BYM model with spatial and temporal components and their complete interactions (Type IV). The observed number of COVID-19, \textbf{$y_{it}$}, was modelled as:
+$$
+y_{it} \sim Poisson (\lambda_{it})\\
+$$
+with i=1,..,`r nprov`, and t=1,...,`r days`. We modelled the ratio $\frac{\lambda_{it}}{N_{i}}$, the Incidence of COVID-19, with a BYM model including temporal covariates as follows  
+$$
+log(\frac{\lambda_{it}}{N_{i}})=\alpha+\mu_i+\nu_i+\gamma_t+\phi_t+\delta_{it},\\
+$$
+
+where $\alpha$ is the intercept and:  
+*space: $\mu_i$ and $\nu_i$ are two area specific effects with normal distribution modelled using an intrinsic conditional autoregressive structure (iCAR);
+
+*time: $\gamma_t$ are the coefficients related to the time process, while $\phi_t \sim N(0, \tau_t^{-1} )$ are the temporal specific random errors.
+Time process can be autoregressive or a random walk process (order 1 or 2);
+
+*space-time: $\delta_{it}$ are the coefficients related space-time interactions.
+
+There are many type of spatio-temporal interactions (please see Spatial and Spatio-temporal Bayesian Models with R - INLA, Blangiardo and Cameletti).
+
+We explored four spatio-temporal interactions:
+-Type I:  coefficients $\delta_{it}$ do not have a spatial or/and temporal structure ($\delta_{it}\sim N(0, 1/\tau_{\delta})$).
+-TypeII: coefficients $\delta_{it}$ have a structured temporal main effect and the unstructured spatial effect.  
+-TypeIII: coefficients $\delta_{it}$ have a spatial structure (as Besag), but are temporally unstructured.
+-TypeVI: coefficients $\delta_{it}$ have both a spatial  and temporal structure. It is the most complex type of space-time interaction.
+
+```{r}
+####specification
+formula.intI<-  totale_casi ~ 1+ f(ID,model="bym",graph=file.adj) +
+f(t,model="rw2") +f(t2,model="iid") +
+f(IDt,model="iid")
+
+formula.intII<-  totale_casi ~ 1+f(ID,model="bym",graph=file.adj) +
+f(t,model="rw2") + f(t2,model="iid") +
+f(ID2,model="iid", group=t3,control.group=list(model="rw2"))
+
+formula.intIII<-   totale_casi ~ 1+f(ID,model="bym",graph=file.adj) +
+f(t,model="rw2") + f(t2,model="iid") +
+f(t3,model="iid", group=ID2, control.group=list(model="besag", graph=file.adj))
+
+formula.intIV<- totale_casi ~ 1+f(ID,model="bym",graph=file.adj) +
+f(t,model="rw2") + f(t2,model="iid") +
+f(ID2,model="besag", graph=file.adj, group=t3,control.group=list(model="rw2"))
+
+#estimation
+fit_st1<-inla(formula.intI, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
+fit_st2<-inla(formula.intII, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
+fit_st3<-inla(formula.intIII, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
+fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
+
+# best DIC
+fit_st1$dic$dic
+fit_st2$dic$dic
+fit_st3$dic$dic
+fit_st4$dic$dic
+### the first model is the best, but we choice the third... because we want to force the spatial structure in the delta_ij.
 ```
 
-#interaction model type 4 (complete time trend interaction to estimate specific curves for each province)
+  The time and space component mean estimates can be extracted here
+
 ```{r}
-####specification
-formula.intIV<- totale_casi ~f(ID,model="bym",graph=file.adj) +
-  f(t,model="rw2",constr = FALSE) +
-  f(t2,model="iid") +
-  f(ID2,model="besag", graph=file.adj,
-    group=t3,
-    control.group=list(model="rw2"))
-#estimation
-fit_st4<-inla(formula.intIV, family="poisson", data=dat_csv, E=pop,control.compute = list(dic=T))
-summary(fit_st4)
+#overall trend
+t_1<-fit_st3$summary.random$t$mean 
+#iid trend
+t_2<-fit_st3$summary.random$t2$mean 
+ # spatial iCAR component
+s_1<-fit_st3$summary.random$ID$mean[1:nprov]
+# spatial idd
+s_2<-fit_st3$summary.random$ID$mean[(nprov+1):(nprov*2)] 
+#spatio-temporal model with car and ar1 specification
+st<-fit_st3$summary.random$t3$mean
+
 ```
-#### components
+
+When can represent the IRR(=exp($\mu_i+\frac{1}{T} \sum_{i=t}^{T} \delta_{it}$)) for each NUTS-3 regions.
+
 ```{r}
-t_1<-fit_st4$summary.random$t$mean #overall trend
-t_2<-fit_st4$summary.random$t2$mean #iid trend
-s_1<-fit_st4$summary.random$ID$mean[1:nprov] # spatial CAR component
-s_2<-fit_st4$summary.random$ID$mean[(nprov+1):(nprov*2)] # spatial idd
-st<-fit_st4$summary.random$ID2$mean # spatio-temporal model with car and ar1 specification ```
-## Mean IRR for each province
+nc.province$IRR_st<-exp(s_1+apply(t(matrix(st,nrow=days)),1,mean))
+l1 = list("sp.text", coordinates(nc.province), nc.province$DEN_PROV,col="grey", cex=0.7,font=2)
+spplot(nc.province, c( "IRR_st"),main="Incidence ", sp.layout=list(l1))
+```
+
+Trend for each province.
+
 ```{r}
-nc.province$IRR_st<-exp(s_1+apply(t(matrix(st,nrow=nprov)),2,mean))
-spplot(nc.province, c( "IRR_st"),main="IRR")
+##overall trend
+plot(Date,exp(t_1),ylab="IRR",main="Overall time trend")
+
+## time trend for each province
+db_plot<-data.frame(t_1+matrix(st,nrow=days))
+names(db_plot)<-nc.province$DEN_PROV
+names(db_plot)[4]<-"Milano"
+province_trend<-xts(db_plot, order.by = Date, frequency = 7)
+p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Province trends") 
+p
+
+#space time variation for each province from overall trend
+db_plot<-data.frame(matrix(st,nrow=days))
+names(db_plot)<-nc.province$DEN_PROV
+names(db_plot)[4]<-"Milano"
+province_trend<-xts(db_plot, order.by = Date, frequency = 7)
+p <- dygraph(province_trend,main=paste("Lombardy Region",sep =""),xlab="Day",height=400,width=800,ylab= "Delta coefficient") 
+p
 ```
-## Trends for province
+
 ```{r}
-trends<-t_1+t(matrix(st,nrow=nprov))
-# trends+ specific
-matplot(trends)
-#specific for each province
-matplot(t(matrix(st,nrow=nprov)))
+q4<-round(quantile(1:days,0:4/4),0)
+
+#Incidence divided by four
+
+Incidence_1q<-tapply(fit_st3$summary.fitted.values$mean[dat_csv$t %in%q4[1]:(q4[2]-1)], dat_csv$ID[dat_csv$t %in%q4[1]:(q4[2]-1)],sum)
+Incidence_2q<-tapply(fit_st3$summary.fitted.values$mean[dat_csv$t %in%q4[2]:(q4[3]-1)], dat_csv$ID[dat_csv$t %in%q4[2]:(q4[3]-1)],sum)
+Incidence_3q<-tapply(fit_st3$summary.fitted.values$mean[dat_csv$t %in%q4[3]:(q4[4]-1)], dat_csv$ID[dat_csv$t %in%q4[3]:(q4[4]-1)],sum)
+Incidence_4q<-tapply(fit_st3$summary.fitted.values$mean[dat_csv$t %in%q4[4]:(q4[5])], dat_csv$ID[dat_csv$t %in%q4[4]:(q4[5])],sum)
+nc.province$Incidence1<-Incidence_1q
+nc.province$Incidence2<-Incidence_2q
+nc.province$Incidence3<-Incidence_3q
+nc.province$Incidence4<-Incidence_4q
+cols = bpy.colors(100)
+col_lim<-range(c(Incidence_1q,Incidence_2q,Incidence_3q,Incidence_4q))
+c1i<-c(round((range(Incidence_1q)[1]/col_lim[2])*100))+1
+c1f<-c(round((range(Incidence_1q)[2]/col_lim[2])*100))
+c2i<-c(round((range(Incidence_2q)[1]/col_lim[2])*100))+1
+c2f<-c(round((range(Incidence_2q)[2]/col_lim[2])*100))
+c3i<-c(round((range(Incidence_3q)[1]/col_lim[2])*100))+1
+c3f<-c(round((range(Incidence_3q)[2]/col_lim[2])*100))
+c4i<-c(round((range(Incidence_4q)[1]/col_lim[2])*100))+1
+c4f<-c(round((range(Incidence_4q)[2]/col_lim[2])*100))
+
+plot<-list()
+i<-1
+p1<-spplot(nc.province, c(paste("Incidence",i,sep="")),main=paste("Incidence ",q4[i],"-",(q4[i+1]-1)," days", sep=""),col.regions =cols[c1i:c1f]) 
+i<-2
+p2<-spplot(nc.province, c(paste("Incidence",i,sep="")),main=paste("Incidence ",q4[i],"-",(q4[i+1]-1)," days", sep=""),col.regions =cols[c2i:c2f])
+i<-3
+p3<-spplot(nc.province, c(paste("Incidence",i,sep="")),main=paste("Incidence ",q4[i],"-",(q4[i+1]-1)," days", sep=""),col.regions =cols[c3i:c3f])
+i<-4
+p4<-spplot(nc.province, c(paste("Incidence",i,sep="")),main=paste("Incidence ",q4[i],"-",(q4[i+1])," days", sep=""),col.regions =cols[c4i:c4f])
+
+gc()
+library(gridExtra)
+grid.arrange(
+p1,p2,p3,p4,nrow = 2)
 ```
@@
 ##################predictions 3 days forward of new CODID 19 cases
+
 ```{r}
@@
 ```
-We model the number of cases.
+
+We modelled the number of cases.
 
 In the rw2 I set constr = FALSE is set to FALSE and that, for this reason, the intercept is not included in the linear predictor.
+
 ```{r}
-## include overdispersion parameter
-formula.intIVn<- totale_casi ~ log(hubei+1)+log(pop)+
-  f(t,model="rw2", constr = FALSE) +
-  f(t2,model="iid") +f(over,model="iid") +
-f(ID,model="bym",graph=file.adj) +
-  f(ID2,model="besag", graph=file.adj,
-    group=t3,
-    control.group=list(model="rw2"))+ f(over,model="iid")
+## include pop e hubei
+formula.int_n<-  totale_casi ~ 1+log(hubei+1)+log(pop)+ f(ID,model="bym",graph=file.adj) +
+f(t,model="rw2") + f(t2,model="iid") +
+f(t3,model="iid", group=ID2, control.group=list(model="besag", graph=file.adj))
 
-fit_st4n<-inla(formula.intIVn, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = 1))
-summary(fit_st4n)
+fit_n<-inla(formula.int_n, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = 1))
+summary(fit_n)
 ```
@@
 #### components
-t_1<-fit_st4n$summary.random$t$mean #overall trend
-t_2<-fit_st4n$summary.random$t2$mean #iid trend
-s_1<-fit_st4n$summary.random$ID$mean[1:nprov] # spatial CAR component
-s_2<-fit_st4n$summary.random$ID$mean[(nprov+1):(nprov*2)] # spatial idd
-st<-fit_st4n$summary.random$ID2$mean # spatio-temporal coefficients 
+#overall trend
+t_1<-fit_n$summary.random$t$mean 
+#iid trend
+t_2<-fit_n$summary.random$t2$mean 
+ # spatial CAR componen
+s_1<-fit_n$summary.random$ID$mean[1:nprov]
+# spatial idd
+s_2<-fit_n$summary.random$ID$mean[(nprov+1):(nprov*2)] 
+# spatio-temporal coefficients 
+st<-fit_n$summary.random$t3$mean
 ## Trends for province
@@
 #extract number of cases for lodi
-est_lodi<-fit_st4n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Lodi",]
+est_lodi<-fit_n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Lodi",]
 

### 2020-03-25 BYM_model/INLA_def.Rmd
Add home page, navigation and theme

@@
 nc.province$DEN_PROV<-droplevels(nc.province$DEN_PROV)
-
 spplot(nc.province, c( "IRR_mean"),main="Incidence Rate Ratios", sp.layout=
          list("sp.text", coordinates(nc.province), nc.province$DEN_PROV, cex=0.7, col="gray"))
-
 ```
@@
 
-##################predictions 3 days forward of new CODID 19 cases
+##################predictions 7 days forward of new CODID 19 cases
 
 ```{r}
-Forecast=3
+Forecast=7
 dat_csv2<-dat_csv 
@@
 
-plot(hubei/58.5,type="l",ylab="COVID19 cases for milion of inhabitant",ylim=c(0,1530),xlab="Days since over 100 cases")
-lines(c(rep(NA,2),cumsum(tapply(dat_csv_n$totale_casi,dat_csv_n$t,sum))/10),col=2)
-abline(v=2,col=1,lty=2)
-abline(v=4,col=2,lty=2)
-abline(v=10,col=3,lty=3)
-legend("bottomright",c("Hubei","Lombardy"),lty=1,col=1:2)
-text(10,1500,"Hubei quarantine")
-text(12,1350,"Lodi home lock",col=2)
-text(20,1200,"Lombardy home lock",col=3)
+#plot(hubei/58.5,type="l",ylab="COVID19 cases for milion of inhabitant",ylim=c(0,1530),xlab="Days since over 100 cases")
+#lines(c(rep(NA,2),cumsum(tapply(dat_csv_n$totale_casi,dat_csv_n$t,sum))/10),col=2)
+#abline(v=2,col=1,lty=2)
+#abline(v=4,col=2,lty=2)
+#abline(v=10,col=3,lty=3)
+#legend("bottomright",c("Hubei","Lombardy"),lty=1,col=1:2)
+#text(10,1500,"Hubei quarantine")
+#text(12,1350,"Lodi home lock",col=2)
+#text(20,1200,"Lombardy home lock",col=3)
 #province of lodi restrictive measure started on 24/02 lag 28 days, at the beginning of the time series
@@
 ## include pop e hubei
-formula.int_n<-  totale_casi ~ 1+log(hubei+1)+log(pop)+ f(ID,model="bym",graph=file.adj) +
-f(t,model="rw2") + f(t2,model="iid") +
-f(t3,model="iid", group=ID2, control.group=list(model="besag", graph=file.adj))
+formula.int_n<-  totale_casi ~ log(pop)+hubei+f(ID,model="bym",graph=file.adj) +
+f(t,model="rw2",constr = FALSE) + f(t2,model="iid") +
+f(over,model="iid")
 
-fit_n<-inla(formula.int_n, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = 1))
+link = rep(NA, dim(dat_csv_n)[1])
+link[which(is.na(dat_csv_n$totale_casi))] = 1
+
+fit_n<-inla(formula.int_n, family="poisson", data=dat_csv_n,control.compute = list(dic=T),control.predictor = list(link = link))
 summary(fit_n)
@@
 # spatio-temporal coefficients 
-st<-fit_n$summary.random$t3$mean
+st<-fit_n$summary.random$over$mean
 ## Trends for province
 Date_n<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+days+Forecast-1,1)
-trends<-t_1+t(matrix(st,nrow=nprov))
-plot(Date_n,t_1,ylab="RW2 Time coefficient",xlab="Date")
-matplot(t(matrix(st,nrow=nprov)),ylab="RW2 Time coefficient for Province",xlab="Date")
-#extract number of cases for lodi
+trends<-t_1+t(matrix(st,ncol=(days+Forecast)))
+plot(Date_n,t_1,ylab="RW2 Time coefficient",xlab="Date",ylim=range(trends),type="l",lwd=2)
+matlines(Date_n,trends,type="l")
+abline(v=Date_n[days],lty=2)
+```
+
+#Forecast for Lodi province
+
+```{r}
 est_lodi<-fit_n$summary.fitted.values[dat_csv_n$denominazione_provincia=="Lodi",]
@@
 days.ahead<-Date_n[(days+1):(days+Forecast)]
-mu.lower<-est_lodi$mean-1.96*est_lodi$sd
-mu.lower[mu.lower<0]<-0
-mu.upper<-est_lodi$mean+1.96*est_lodi$sd
-mu.med<-xts(est_lodi$mean,order.by = c(days.before,days.ahead),frequency = 7)
+mu.lower<-est_lodi$`0.025quant`
+mu.upper<-est_lodi$`0.975quant`
+mu.med<-xts(est_lodi$`0.5quant`,order.by = c(days.before,days.ahead),frequency = 7)
 counts<-mu.med

### 2020-03-25 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

@@
 ```{r}
-names(dat_csv[,-c(1:2,13)])
+names(dat_csv[,-c(1:2,17:19)])
 
@@
 ```{r}
-myDateTimeStr1 <- dat_csv$data
+myDateTimeStr1 <- paste(substr(dat_csv$data,1,10),substr(dat_csv$data,12,19))
 myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
 days_dy<-as.Date(myPOSIXct1)
-dat_csv_dy<-xts(dat_csv[,-c(1:2,12:13)], order.by = days_dy, frequency = 7)
+dat_csv_dy<-xts(dat_csv[,-c(1:2,13:15)], order.by = days_dy, frequency = 7)
 ```

### 2020-03-25 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

@@
 ```{r}
-myDateTimeStr1 <- dat_csv$data
+myDateTimeStr1 <- paste(substr(dat_csv$data,1,10),substr(dat_csv$data,12,19))
 myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
 days_dy<-as.Date(myPOSIXct1)
-dat_csv_dy<-xts(dat_csv[,-c(1:7,16:17)], order.by = days_dy, frequency = 7)
+dat_csv_dy<-xts(dat_csv[,-c(1:7,17:19)], order.by = days_dy, frequency = 7)
 ```

### 2020-03-27 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

@@
 
-We suppose that the number of total infected $I(t)$ can be the realization of a Poisson random variable
-$$
-I(t) \sim Poisson (\theta_t)
-$$
-where $\theta_t$ is the unknown mean of the random process at the time $t$.
-
-The Poisson distribution has a fixed variance that is equal to the mean; in the next part we try to estimates the parameters considering a Negative Binomial distribution which allows for overdispersion.
-
-However, the parameter $\theta_t$ can be supposed to be a realization of a SEIR model with initial parameters $\theta=(\mu,\beta,\sigma, \gamma)$  and initial status $S=(S_{start},E_{start},I_{start},R_{start})$.
-
-The parameter $\sigma$ is our parameter of interest; we estimate $\sigma$ fixing $E_{start}= 3*I_{start}$ minimizing the log likelihood of the described Poisson distribution (or Negative Binomial one).
+s
 

### 2020-03-30 BYM_model/INLA_def.Rmd
Add home page, navigation and theme

@@
 ###import international dataset
-dat_int<-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv",header=T)
+dat_int<-read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv",header=T)
 hubei<-as.numeric(dat_int[dat_int$Province.State=="Hubei",5:dim(dat_int)[2]])

### 2020-04-02 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

@@
 ```{r}
-dat_csv_dy$log_totale_attualmente_positivi<-log(dat_csv_dy$totale_attualmente_positivi)
-p <- dygraph(dat_csv_dy$log_totale_attualmente_positivi,main=paste("Italy"),ylab="Log Infected case",xlab="Day",height=400,width=800) 
+dat_csv_dy$log_totale_positivi<-log(dat_csv_dy$totale_positivi)
+p <- dygraph(dat_csv_dy$log_totale_positivi,main=paste("Italy"),ylab="Log Infected case",xlab="Day",height=400,width=800) 
 p
@@
 for (i in 3:(days-2)){
-fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(i-2):(i+2),])
+fit <- lm(log(totale_positivi)~t,data=dat_csv[(i-2):(i+2),])
 beta_vec<-c(beta_vec,coef(fit)[2])
@@
 # initial number of infectus
-I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
+I_start<-dat_csv$totale_positivi[dim(dat_csv)[1]]; I_start
 # initial number of recovered, based on the proportion of discharged from the health services
@@
 #I_start
-I_1<-dat_csv$totale_attualmente_positivi[(days-last)]
+I_1<-dat_csv$totale_positivi[(days-last)]
 #R_start
@@
 initials <- c(S = (1-(f_exp+I_1/N-R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
--LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+-LogLikfun(initials,parameters,dat_csv$totale_positivi[c((days-last):days)])
 }
@@
 #predicted vs observed
-cbind(pro$results$I*N,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+cbind(pro$results$I*N,dat_csv$totale_positivi[c((days-last):days)])
 ```
@@
 step.ahead<-forecast+1
-mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
-mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
-mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+mu.lower<-c(dat_csv$totale_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
 counts<-mu.med

### 2020-04-02 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

--- 2020-03-24 draft_analysis_Veneto_new.Rmd
@@
 ```{r}
-names(dat_csv[,-c(1:2,13)])
+names(dat_csv[,-c(1:6,18:20)])
 
@@
 ```{r}
-myDateTimeStr1 <- dat_csv$data
+myDateTimeStr1 <- paste(substr(dat_csv$data,1,10),substr(dat_csv$data,12,19))
 myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
 days_dy<-as.Date(myPOSIXct1)
-dat_csv_dy<-xts(dat_csv[,-c(1:7,16:17)], order.by = days_dy, frequency = 7)
+dat_csv_dy<-xts(dat_csv[,-c(1:6,18:20)], order.by = days_dy, frequency = 7)
 ```
@@
 ```{r}
-dat_csv_dy$log_totale_attualmente_positivi<-log(dat_csv_dy$totale_attualmente_positivi)
-p <- dygraph(dat_csv_dy$log_totale_attualmente_positivi,main=paste("Veneto Region"),ylab="Log Infected case",xlab="Day",height=400,width=800) 
+dat_csv_dy$log_totale_positivi<-log(dat_csv_dy$totale_positivi)
+p <- dygraph(dat_csv_dy$log_totale_positivi,main=paste("Veneto Region"),ylab="Log Infected case",xlab="Day",height=400,width=800) 
 p
@@
 initials <- c(S = (1-(f_exp+I_1/N+R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
--LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+-LogLikfun(initials,parameters,dat_csv$totale_positivi[c((days-last):days)])
 }
@@
 ```{r  intensive care numbers, echo=TRUE}
-fit_int_care<-glm(cbind(terapia_intensiva,totale_attualmente_positivi-terapia_intensiva)~t,family=binomial,data=dat_csv)
+fit_int_care<-glm(cbind(terapia_intensiva,totale_positivi-terapia_intensiva)~t,family=binomial,data=dat_csv)
 
@@
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-plot(date,pre$fit,type="l",main="Proportion of admitted COVID 19 in intensive care",ylim=range(dat_csv$terapia_intensiva/dat_csv$totale_attualmente_positivi))
-points(date[1:days],dat_csv$terapia_intensiva/dat_csv$totale_attualmente_positivi)
+plot(date,pre$fit,type="l",main="Proportion of admitted COVID 19 in intensive care",ylim=range(dat_csv$terapia_intensiva/dat_csv$totale_positivi))
+points(date[1:days],dat_csv$terapia_intensiva/dat_csv$totale_positivi)
 lines(date,pre$fit+1.96*pre$se.fit,col=3,lty=2)

### 2020-04-03 draft_analysis_Veneto_html.Rmd
Add home page, navigation and theme

--- 2020-03-24 draft_analysis_Veneto_html.Rmd
@@
 duration<-18
-beta_vec<-NULL
-sd_vec<-NULL
+beta1<-NULL
+sd_beta1<-NULL
 #for cycle for R0 estimates from days-2 to days+2
 for (i in 3:(days-2)){
-fit <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(i-2):(i+2),])
-beta_vec<-c(beta_vec,coef(fit)[2])
-sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
+fit <- lm(log(totale_positivi)~t,data=dat_csv[(i-2):(i+2),])
+beta1<-c(beta1,coef(fit)[2])
+sd_beta1<-c(sd_beta1,coef(summary(fit))[2,2])
 }
@@
 
-mean  <- 1+(beta_vec*duration)
-lower <- 1+((beta_vec-1.96*sd_vec)*duration)
-upper <- 1+((beta_vec+1.96*sd_vec)*duration)
+mean  <- 1+(beta1*duration)
+lower <- 1+((beta1-1.96*sd_beta1)*duration)
+upper <- 1+((beta1+1.96*sd_beta1)*duration)
 
@@
 The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
-We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Normal distribution for the beta (the slope).
+We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Log-Normal distribution for the the $\beta1$ and as consequence for the $\beta$ for the regression, rescaled of $1/\gamma$.
+We implement function to deal with log-normal distribution, in order to recover the parameter $\mu$ and $\sigma$ belonging normal distribution starting from this one:
+$$\beta_{LogN}(t) \sim Log N (\beta_1(t)+\frac{1}{\gamma}, \sigma_{\beta_1(t)}^2)$$
+In order to deal with estimations of $\hat\beta_1(t)$ that we have estimated before, we adopted an inverse transformation as reported here https://ocw.mit.edu/courses/civil-and-environmental-engineering/1-151-probability-and-statistics-in-engineering-spring-2005/lecture-notes/briefnts8_relnl.pdf.
+The following formulas provide us the mean and variance of the normal distribution
+
+$$\beta_{N}(t) \sim N (\mu_{\beta_N},\sigma_{\beta_N}^2)$$
+derived from the log-normal distribution $\beta_{LogN}(t)$. 
+$$\mu_{\beta_N} =e^{\mu_{\beta_{LogN}}+0.5*\sigma_{\beta_{LogN}}^2}$$
+
+$$\sigma_{\beta_N}^2 =-2\log(\mu_{\beta_{LogN}})+\log(\mu_{\beta_{LogN}}^2+\sigma_{\beta_{LogN}}^2)$$
+This transformation permits us to perform an inference based on the classical regression model for the predict forward of the future values of $\beta_{LogN}(t)$ with a direct transformation.
+
+```{r log normal formula}
+# define direct transf function
+M_from_Norm<-function(mu,sigma) exp(mu+0.5*sigma^2)
+SD_from_Norm<-function(mu,sigma)  sqrt(exp(2*mu+sigma^2)*(exp(sigma^2)-1))
+# define inverse transf function
+M_from_LNorm<-function(mu,sigma) 2*log(mu)-0.5*log(mu^2+sigma^2)
+SD_from_LNorm<-function(mu,sigma) sqrt( -2*log(mu)+log(mu^2+sigma^2))
+```
+
 ```{r R0 forecast}
@@
 data<-data.frame(time)
-beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian,data)
+beta_norm<-M_from_LNorm(beta1+1/duration,sd_beta1)
+sd_beta_norm<-SD_from_LNorm(beta1+1/duration,sd_beta1)
+sd_norm_mean<-sqrt(mean(SD_from_LNorm(beta1,sd_beta1)^2))
+
+beta.model<-glm(beta_norm~time,weights = 1/(sd_beta_norm)^2,family=gaussian,data)
 summary(beta.model)
@@
 #we can use
-#summary(gamlss(formula = beta_vec ~ time + bs(weekend, 3),      weights = 1/sd_vec,Lp=1) )
+#summary(gamlss(formula = beta_norm ~ time + bs(weekend, 3),Lp=1) )
 forecast=14
@@
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-predict <- data.frame(beta_vec=c(rep(NA,2),beta_vec,rep(NA,forecast+2)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
+predict <- data.frame(beta=c(rep(NA,2),beta_norm,rep(NA,forecast+2)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
 beta.predict<-predict 
 r0.predict<-beta.predict
-r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*duration+1
+r0.predict[,c(1,3:5)]<-M_from_Norm(r0.predict[,c(1,3:5)],sd_norm_mean)*duration
 # plot the points (actual observations), regression line, and confidence interval
-p <- ggplot(r0.predict, aes(date,beta_vec))
+p <- ggplot(r0.predict, aes(date,beta))
 p <- p + geom_point() +labs(x="Date",y="R0 value") 
@@
 f_exp<-3*I_1/N
-parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[(days-last-2)])+1/duration), sigma = sigma, gamma =1/duration)
+parameters <- c(mu = mu0, beta =(as.numeric(beta1[(days-last-2)])+1/duration), sigma = sigma, gamma =1/duration)
 initials <- c(S = (1-(f_exp+I_1/N+R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
--LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+-LogLikfun(initials,parameters,dat_csv$totale_positivi[c((days-last):days)])
 }
@@
 
-parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[(days-last-2)])+1/duration), sigma = sigma, gamma = 1/duration)
+parameters <- c(mu = mu0, beta =(as.numeric(beta1[(days-last-2)])+1/duration), sigma = sigma, gamma = 1/duration)
 initials <- c(S = (1-(f_exp+I_1/N+R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
@@
 # Bayesian Structural Time Series
-#we drop the first 6 beta_vec
-ss <- AddStudentLocalLinearTrend(list(), beta_vec[6:(days-4)])
-ss <- AddSeasonal(ss, beta_vec[6:(days-4)], nseasons = 7)
-model1 <- bsts(beta_vec[6:(days-4)],
+#we assume beta ~ LN (mu_b,sigma_b)
+
+#we drop the first 6 beta_norm
+ss <- AddStudentLocalLinearTrend(list(), beta_norm[6:(days-4)])
+ss <- AddSeasonal(ss, beta_norm[6:(days-4)], nseasons = 7)
+model1 <- bsts(beta_norm[6:(days-4)],
                state.specification = ss,
@@
 pred1 <- predict(model1, horizon = 16, burn = 100)
-par(mfrow = c(1,1))
-plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions",ylim=c(-1,1)*IQR(pred1$distribution)+median(pred1$distribution))
+plot(pred1 , ylab="Log Beta1 coefficient",main="Data and predictions",ylim=c(-3,3)*IQR(pred1$distribution)+median(pred1$distribution))
 # matrix of beta coefficients
-coef<-(pred1$distribution[,3:16])
+coef<-M_from_Norm(pred1$distribution[,3:16],sd_norm_mean)
 par(mfrow = c(1,1))
@@
 For each vector of simulated beta coefficients we perform a SEIR model.
-We save the results and plot a credible interval at 50% (25-75% percentile).
+We save the results and plot a credible interval at 90% (5-95% percentile).
 
@@
   for(i in 1:forecast){
-    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
+    parameters <- c(mu = mu0, beta = matrix(coef[s,i]), sigma = sigma, gamma = 1/duration)
     if( i==1) initials <- c(S = 1-(pro$results$E[last+1]+I_start/N+R_start/N), E = pro$results$E[last+1], I = I_start/N, R = R_start/N)
@@
 I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
-I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.25)
-I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.75)
+I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.05)
+I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.95)
 
@@
 step.ahead<-forecast+1
-mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
-mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
-mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+mu.lower<-c(dat_csv$totale_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
 counts<-mu.med
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Credible Interval ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
@@
 set.seed(123)
-ss <-   AddLocalLevel(list(), beta_vec[(days-9):(days-4)],sigma=SdPrior(sd_vec[(days-4)]))
-model1 <- bsts(beta_vec[(days-9):(days-4)],
+#we assum log_beta ~ N()
+ss <-   AddLocalLevel(, beta_norm[(days-9):(days-4)],sdy=sd_norm_mean)
+model1 <- bsts(beta_norm[(days-9):(days-4)],
                state.specification = ss,
@@
 # matrix of beta coefficients
-coef<-(pred1$distribution[,3:16])
+coef<-M_from_Norm(pred1$distribution[,3:16],sd_norm_mean)
 par(mfrow = c(1,1))

### 2020-04-03 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

@@
 duration<-18
-beta_vec<-NULL
-sd_vec<-NULL
+beta1<-NULL
+sd_beta1<-NULL
 #for cycle for R0 estimates from days-2 to days+2
@@
 fit <- lm(log(totale_positivi)~t,data=dat_csv[(i-2):(i+2),])
-beta_vec<-c(beta_vec,coef(fit)[2])
-sd_vec<-c(sd_vec,coef(summary(fit))[2,2])
+beta1<-c(beta1,coef(fit)[2])
+sd_beta1<-c(sd_beta1,coef(summary(fit))[2,2])
 }
@@
 The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
-We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Normal distribution for the beta (the slope).
+We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Log-Normal distribution for the the $\beta1$ and as consequence for the $\beta$ for the regression, rescaled of $1/\gamma$.
+We implement function to deal with log-normal distribution, in order to recover the parameter $\mu$ and $\sigma$ belonging normal distribution starting from this one:
+$$\beta_{LogN}(t) \sim Log N (\beta_1(t)+\frac{1}{\gamma}, \sigma_{\beta_1(t)}^2)$$
+In order to deal with estimations of $\hat\beta_1(t)$ that we have estimated before, we adopted an inverse transformation as reported here https://ocw.mit.edu/courses/civil-and-environmental-engineering/1-151-probability-and-statistics-in-engineering-spring-2005/lecture-notes/briefnts8_relnl.pdf.
+The following formulas provide us the mean and variance of the normal distribution
+
+$$\beta_{N}(t) \sim N (\mu_{\beta_N},\sigma_{\beta_N}^2)$$
+derived from the log-normal distribution $\beta_{LogN}(t)$. 
+$$\mu_{\beta_N} =e^{\mu_{\beta_{LogN}}+0.5*\sigma_{\beta_{LogN}}^2}$$
+
+$$\sigma_{\beta_N}^2 =-2\log(\mu_{\beta_{LogN}})+\log(\mu_{\beta_{LogN}}^2+\sigma_{\beta_{LogN}}^2)$$
+This transformation permits us to perform an inference based on the classical regression model for the predict forward of the future values of $ \beta_{LogN}(t) $ with a direct transformation.
+
+```{r log normal formula}
+# define direct transf function
+M_from_Norm<-function(mu,sigma) exp(mu+0.5*sigma^2)
+SD_from_Norm<-function(mu,sigma)  sqrt(exp(2*mu+sigma^2)*(exp(sigma^2)-1))
+# define inverse transf function
+M_from_LNorm<-function(mu,sigma) 2*log(mu)-0.5*log(mu^2+sigma^2)
+SD_from_LNorm<-function(mu,sigma) sqrt( -2*log(mu)+log(mu^2+sigma^2))
+```
+
 ```{r R0 forecast}
@@
 f_exp<-3*I_1/N
-parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[(days-last-2)])+1/duration), sigma = sigma, gamma =1/duration)
+parameters <- c(mu = mu0, beta =(as.numeric(beta1[(days-last-2)])+1/duration), sigma = sigma, gamma =1/duration)
 initials <- c(S = (1-(f_exp+I_1/N+R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
@@
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Credible Interval ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")

### 2020-04-07 draft_analysis_Italy_new.Rmd
Add home page, navigation and theme

--- 2020-03-25 draft_analysis_Italy_new.Rmd
@@
 ```{r}
-names(dat_csv[,-c(1:2,17:19)])
+names(dat_csv[,-c(1:2,14:16)])
 
@@
 days_dy<-as.Date(myPOSIXct1)
-dat_csv_dy<-xts(dat_csv[,-c(1:2,13:15)], order.by = days_dy, frequency = 7)
+dat_csv_dy<-xts(dat_csv[,-c(1:2,14:16)], order.by = days_dy, frequency = 7)
 ```
@@
 ```
+The time series of COVID-19 in the Italy.
 
+```{r}
+plot(days_dy[-1],diff(dat_csv$totale_casi), ylab="New cases",xlab="Date",type="l")
+```
 ### The S(E)IR model (to be revised)
@@
 basis of the actual data, we used a SEIR model applied to the COVID19
-epidemic to Italy
+epidemic to the Italy.
 
@@
 
-However, in the calculation of R0 we considered an infectious period of 18 days (https://www.nejm.org/doi/10.1056/NEJMoa2001316).
+However, in the calculation of R0 we considered an infectious period of 18 days on the basis of recent publications (https://www.nature.com/articles/s41421-020-0148-0).
 
@@
 The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
-We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Normal distribution for the beta (the slope).
+We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Log-Normal distribution for the the $\beta1$ and as consequence for the $\beta$ for the regression, rescaled of $1/\gamma$.
+We implement function to deal with log-normal distribution, in order to recover the parameter $\mu$ and $\sigma$ belonging normal distribution starting from this one:
+$$\beta_{LogN}(t) \sim Log N (\beta_1(t)+\frac{1}{\gamma}, \sigma_{\beta_1(t)}^2)$$
+In order to deal with estimations of $\hat\beta_1(t)$ that we have estimated before, we adopted an inverse transformation as reported here https://ocw.mit.edu/courses/civil-and-environmental-engineering/1-151-probability-and-statistics-in-engineering-spring-2005/lecture-notes/briefnts8_relnl.pdf.
+The following formulas provide us the mean and variance of the normal distribution
+
+$$\beta_{N}(t) \sim N (\mu_{\beta_N},\sigma_{\beta_N}^2)$$
+derived from the log-normal distribution $\beta_{LogN}(t)$. 
+$$\mu_{\beta_N} =e^{\mu_{\beta_{LogN}}+0.5*\sigma_{\beta_{LogN}}^2}$$
+
+$$\sigma_{\beta_N}^2 =-2\log(\mu_{\beta_{LogN}})+\log(\mu_{\beta_{LogN}}^2+\sigma_{\beta_{LogN}}^2)$$
+This transformation permits us to perform an inference based on the classical regression model for the predict forward of the future values of $\beta_{LogN}(t)$ with a direct transformation.
+
+```{r log normal formula}
+# define direct transf function
+M_from_Norm<-function(mu,sigma) exp(mu+0.5*sigma^2)
+SD_from_Norm<-function(mu,sigma)  sqrt(exp(2*mu+sigma^2)*(exp(sigma^2)-1))
+# define inverse transf function
+M_from_LNorm<-function(mu,sigma) 2*log(mu)-0.5*log(mu^2+sigma^2)
+SD_from_LNorm<-function(mu,sigma) sqrt( -2*log(mu)+log(mu^2+sigma^2))
+```
+
 ```{r R0 forecast}
+#start from the day 7, since vo hotspot can influence the trend
 time<-3:(days-2)
 weekend<-rep(c(1,2,3,4,5,6,7),ceiling(days/7))[3:(days-2)]
-data=data.frame(time,weekend)
-beta.model<-glm(beta_vec~time+I(cos(time*2*pi/7))+I(sin(time*2*pi/7)),weights = 1/sd_vec,family=gaussian,data=data)
+data<-data.frame(time)
+beta_norm<-M_from_LNorm(beta1+1/duration,sd_beta1)
+sd_beta_norm<-SD_from_LNorm(beta1+1/duration,sd_beta1)
+sd_norm_mean<-sqrt(mean(SD_from_LNorm(beta1,sd_beta1)^2))
+
+beta.model<-glm(beta_norm~time,weights = 1/(sd_beta_norm)^2,family=gaussian,data)
 summary(beta.model)
@@
 #there is an effect of the week, as supposed 
+#we can use
+#summary(gamlss(formula = beta_norm ~ time + bs(weekend, 3),Lp=1) )
 forecast=14
 # add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
-weekend_pre<-rep(c(1,2,3,4,5,6,7),ceiling((days+forecast)/7))[1:(days+forecast)]
-datanew<-data.frame(time=1:(days+forecast),weekend=weekend_pre)
-pre<-predict(beta.model,type='response',newdata=datanew,se.fit=TRUE)
+pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-predict <- data.frame(beta_vec=c(rep(NA,2),beta_vec,rep(NA,forecast+2)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
+predict <- data.frame(beta=c(rep(NA,2),beta_norm,rep(NA,forecast+2)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
 beta.predict<-predict 
 r0.predict<-beta.predict
-r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*duration+1
+r0.predict[,c(1,3:5)]<-M_from_Norm(r0.predict[,c(1,3:5)],sd_norm_mean)*duration
 # plot the points (actual observations), regression line, and confidence interval
-p <- ggplot(r0.predict, aes(date,beta_vec))
+p <- ggplot(r0.predict, aes(date,beta))
 p <- p + geom_point() +labs(x="Date",y="R0 value") 
@@
 We made a forecast by means of a SEIR model fixing a series of initial status:  
-*S:N, the size of Italian population  
+*S:N, the size of the Italian population  
 *E: The number of exposed people, but supposed to be at least 3 times the infected (3 x I_start)  
@@
 and parameters:   
-- beta: gamma + slope  
-- gamma= 1/duration of diseases (duration=21 days)   
-- sigma: the coronavirus transmission rate   
-- mu0: the overall mortality rate
+-beta: gamma + slope   
+-gamma= 1/duration of diseases (duration=21 days)   
+-sigma: the coronavirus transmission rate   
+-mu0: the overall mortality rate
 
@@
 R_start<-prop*dat_csv$totale_casi[dim(dat_csv)[1]]; R_start
-# Italian population
-N=60480000
-# duration of COVID19 as the sum of incubation (5-14 days) and the duration of diseases (about 18 days)
+# Italy
+N=60317000
+#  duration of COVID19 diseases 
 duration<-18
@@
 n<-length(obs)
-N=60480000
+N=60317000
 seir_fit <- SEIR(pars = parameters, init = initials, time = 0:last)
 #Poisson
-#sum(dpois(x=obs[-1],lambda = seir_fit$results$I[-1]*N,log=TRUE))
+sum(dpois(x=obs[-1],lambda = seir_fit$results$I[-1]*N,log=TRUE))
 #Neg Binomial
-sum(dnbinom(x=obs[-1],mu=seir_fit$results$I[-1]*N,size=N,log=TRUE))
+#sum(dnbinom(x=obs[-1],mu=seir_fit$results$I[-1]*N,size=N,log=TRUE))
 #SSE
@@
 f1<-function(par){
+#max 0.20 according to https://www.nejm.org/doi/10.1056/NEJMoa2001316 is 1/5.2
 sigma<-expit(par)
 f_exp<-3*I_1/N
-parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-2)])+1/duration), sigma = sigma, gamma =1/duration)
-initials <- c(S = (1-(f_exp+I_1/N-R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
--LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+parameters <- c(mu = mu0, beta =(as.numeric(beta1[(days-last-2)])+1/duration), sigma = sigma, gamma =1/duration)
+initials <- c(S = (1-(f_exp+I_1/N+R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
+-LogLikfun(initials,parameters,dat_csv$totale_positivi[c((days-last):days)])
 }
 est<-optim(fn=f1,par=logit(0.1))
-#estimated par
+est
 expit(est$par)
+
+f_exp<-3*I_1/N
 sigma<-expit(est$par)
-f_exp<-3*I_1/N
-parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-2)])+1/duration), sigma = sigma, gamma = 1/duration)
-initials <- c(S = (1-(f_exp+I_1/N-R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
+
+parameters <- c(mu = mu0, beta =(as.numeric(beta1[(days-last-2)])+1/duration), sigma = sigma, gamma = 1/duration)
+initials <- c(S = (1-(f_exp+I_1/N+R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
 pro<-SEIR(pars = parameters, init = initials, time = 0:last)
 #predicted vs observed
-cbind(pro$results$I*N,dat_csv$totale_attualmente_positivi[c((days-last):days)])
+cbind(pro$results$I*N,dat_csv$totale_positivi[c((days-last):days)])
 ```
@@
 ```{r status, echo=TRUE}
-#the unknowrn fraction of exposed people is
-pro$results$E[last+1];
+#mthe unknowrn fraction of exposed people is
+pro$results$E[last+1]
 #that is 
@@
 ```
-For the beta parameter, we perfom a simulation on its trend by means of a Bayesian Structural Time Series using the library bsts of R.    
-We estimate a BSTS model specifing a local linear trend ans 1000 simulation.
+
+# Forecast 1 - Results based on beta coefficient trend
+
+For the beta parameter, we perfomed a simulation on its trend by means of a Bayesian Structural Time Series using the library bsts of R.    
+We estimated a BSTS model specifing a local linear trend and a seasonality component. The model predicted 1000 trajectories forward of the beta coefficent.
 
@@
 # Bayesian Structural Time Series
-ss <- AddLocalLinearTrend(list(), beta_vec)
-ss <- AddSeasonal(ss, beta_vec, nseasons = 7)
-model1 <- bsts(beta_vec,
+#we assume beta ~ LN (mu_b,sigma_b)
+
+#we drop the first 6 beta_norm
+ss <- AddStudentLocalLinearTrend(list(), beta_norm[6:(days-4)])
+#ss <- AddSeasonal(ss, beta_norm[6:(days-4)], nseasons = 7)
+model1 <- bsts(beta_norm[6:(days-4)],
                state.specification = ss,
@@
 pred1 <- predict(model1, horizon = 16, burn = 100)
-par(mfrow = c(1,1))
-plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions",ylim=c(-1,1)*IQR(pred1$distribution)+median(pred1$distribution))
+plot(pred1 , ylab="Log Beta1 coefficient",main="Data and predictions",ylim=c(-3,3)*IQR(pred1$distribution)+median(pred1$distribution))
 # matrix of beta coefficients
-coef<-(pred1$distribution[,3:16])
+coef<-M_from_Norm(pred1$distribution[,3:16],sd_norm_mean)
 par(mfrow = c(1,1))
 ```
+
 For each vector of simulated beta coefficients we perform a SEIR model.
-We save the results and plot a credible interval at 50% (25-75% percentile).
+We save the results and plot a credible interval at 90% (5-95% percentile).
 
@@
   for(i in 1:forecast){
-    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
-    if( i==1) initials <- c(S = 1-(pro$results$E[last+1]+I_start/N+R_start/N), E =pro$results$E[last+1] , I = I_start/N, R = R_start/N)
+    parameters <- c(mu = mu0, beta = matrix(coef[s,i]), sigma = sigma, gamma = 1/duration)
+    if( i==1) initials <- c(S = 1-(pro$results$E[last+1]+I_start/N+R_start/N), E = pro$results$E[last+1], I = I_start/N, R = R_start/N)
     if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
@@
 step.ahead<-forecast+1
-mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
-mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
-mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
+mu.lower<-c(dat_csv$totale_positivi,I_seir_lwr*N)
+mu.upper<-c(dat_csv$totale_positivi,I_seir_upr*N)
+mu.med<-xts(c(dat_csv$totale_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
 counts<-mu.med
 mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
-p <- dygraph(mu,main=paste("Italy: Scenario  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
+p <- dygraph(mu,main=paste("Italy: Scenario (Credible Interval ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
 p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")

### 2020-04-07 draft_analysis_Italy_html.Rmd
Add home page, navigation and theme

@@
 ```{r}
-names(dat_csv[,-c(1:2,13)])
+names(dat_csv[,-c(1:2,14:16)])
 
@@
 ```{r}
-myDateTimeStr1 <- dat_csv$data
+myDateTimeStr1 <- paste(substr(dat_csv$data,1,10),substr(dat_csv$data,12,19))
 myPOSIXct1 <- as.POSIXct(myDateTimeStr1, format="%Y-%m-%d %H:%M:%S")
 days_dy<-as.Date(myPOSIXct1)
-dat_csv_dy<-xts(dat_csv[,-c(1:2,12:13)], order.by = days_dy, frequency = 7)
+dat_csv_dy<-xts(dat_csv[,-c(1:2,14:16)], order.by = days_dy, frequency = 7)
 ```

### 2020-04-07 draft_analysis_Veneto_new.Rmd
Add home page, navigation and theme

@@
 $$\sigma_{\beta_N}^2 =-2\log(\mu_{\beta_{LogN}})+\log(\mu_{\beta_{LogN}}^2+\sigma_{\beta_{LogN}}^2)$$
-This transformation permits us to perform an inference based on the classical regression model for the predict forward of the future values of $ \beta_{LogN}(t) $ with a direct transformation.
+This transformation permits us to perform an inference based on the classical regression model for the predict forward of the future values of $\beta_{LogN}(t)$ with a direct transformation.
 
@@
 ```{r  intensive care numbers, echo=TRUE}
-fit_int_care<-glm(cbind(terapia_intensiva,totale_positivi-terapia_intensiva)~t,family=binomial,data=dat_csv)
+fit_int_care<-glm(cbind(terapia_intensiva,totale_positivi-terapia_intensiva)~t+I(t^2),family=binomial,data=dat_csv[])
 
@@
 date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
-plot(date,pre$fit,type="l",main="Proportion of admitted COVID 19 in intensive care",ylim=range(dat_csv$terapia_intensiva/dat_csv$totale_positivi))
+plot(date,pre$fit,type="l",main="Proportion of admitted COVID 19 in intensive care",ylim=range(c(dat_csv$terapia_intensiva/dat_csv$totale_positivi,pre$fit)))
 points(date[1:days],dat_csv$terapia_intensiva/dat_csv$totale_positivi)
